### 2020-03-26 Markdown Documents/Analysis Progress Meeting 3_24.Rmd
x

---
title: "Analysis Progress Meeting 3/24"
author: "Matt Clark"
date: "3/24/2020"
output: html_document
---

```{r setup, include=FALSE}
library(rstan)
library(tidyverse)
dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
x<-dat%>%group_by(UnitCode,Season)%>%
  summarise(mv=max(RecreationVisits))%>%ungroup()%>%
  group_by(UnitCode)%>%filter(mv==max(mv))%>%mutate(cc=paste0(UnitCode,Season))

x2<-dat%>%group_by(UnitCode,Season)%>%
  summarise(mv=min(RecreationVisits))%>%ungroup()%>%
  group_by(UnitCode)%>%filter(mv==min(mv))%>%mutate(cc=paste0(UnitCode,Season))

dat$SeasType<-ifelse(dat$CatColS %in% x$cc,
                       "High",ifelse(dat$CatColS %in% x2$cc,"Low","Shoulder"))
#load(fit)
#load(fit2)
```

# Overall data

Looking at the overall, linear trends of smoke on vis didn't yeild any immediatly apparent insights
```{r echo=FALSE}
ggplot(dat,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=F,alpha=0.6,method="lm")+
  theme_classic()+ggtitle("Smoke values standardized accross all parks")

ggplot(dat,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=F,alpha=0.6,method="lm")+
  theme_classic()+ggtitle("Smoke values standardized for each park individually")

```

So I also looked at the polynomial curves
```{r echo=FALSE}
ggplot(dat,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=F,alpha=0.6)+
  theme_classic()+ggtitle("Smoke values standardized accross all parks")

ggplot(dat,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=F,alpha=0.6)+
  theme_classic()+ggtitle("Smoke values standardized for each park individually")
```

These apprear to have a breakpoint, i.e. vis increases with smoke (due to season) up to a certain point, but then rapidly drops off. We can model this and get breakpoint estimates at least for a subset of parks

I don't actually think that this is a good idea though. I think these plots have an overwhelming effect of season. Many parks when there is no smoke, it actually makes no sense to go there e.g. Rocky Mountain. The problem is that not all parks have the same visitation or smoke trends by season.

```{r echo=FALSE}
ggplot(dat,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(aes(color=Season))+
  theme_classic()+ggtitle("Smoke values standardized accross all parks")
```

It might be better to just compare between high, low, and shoulder seasons for each park.

```{r echo = FALSE}
ggplot(dat,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(aes(color=SeasType))+
  theme_classic()+ggtitle("Smoke values standardized for each park individually")
```

One obvious starting place here is to just look at the high season for each park.

We can look at the by park standardization and the overall standardiztion.

```{r echo= F}
dat%>%filter(SeasType =="High")%>%
ggplot(.,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=F,alpha=0.6,method="lm")+
  theme_classic()+ggtitle("Smoke values standardized for each park individually")

dat%>%filter(SeasType =="High")%>%
ggplot(.,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=F,alpha=0.6,method="lm")+
  theme_classic()+ggtitle("Smoke values standardized for each park individually")
```

Let's look at some simple model outputs for these

Standardized overall

__Didn't converge for some reason, I'll work on it today__

Standardized by park

__Didn't converge for some reason, I'll work on it today__

We can see from this that the effect of smoke on visitation is stronger when you consider the change from a baseline for each park.

Let's look at this for the other seasons now

```{r}
ggplot(dat,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(aes(color=SeasType),se=F,alpha=0.6,method="lm")+
  theme_classic()+ggtitle("Smoke values standardized for each park individually")
```

Might be interesting to look at the polynomials for those too

```{r}
ggplot(dat,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(aes(color=SeasType),se=F,alpha=0.6)+
  theme_classic()+ggtitle("Smoke values standardized for each park individually")
```

I'm not really sure what to make of that weird upswing that we see. I ran a breakpoint model on just the high season for these data and it picked up the upswing at the primary breakpoint. Basically the opposite of our hypothesis. I think that there's just a few high leverage points causing this though, I think this because of how wide the std error gets over there.

```{r echo = F}
dat%>%filter(SeasType =="High")%>%
  ggplot(.,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(aes(color=SeasType),alpha=0.6)+
  theme_classic()+ggtitle("Smoke values standardized for each park individually")
```

# Overall insights

I think overall what we're seeing is a ton of noise intoduced by the low seasons at parks. When we look at just the high season we see a negative impact of smoke to a very slight degree. We see that this effect is stronger when we scale the smoke by each park individually. We can say something about relative baselines with this as well as give projections about visitation lost for each park.

### 2020-03-26 Markdown Documents/Hypotheses_Rationale.Rmd
x

---
title: "Rationale and Hypotheses"
author: "Matt Clark & Alex Killion"
date: "11/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Testing - What is the best way to replicate a Word comment bubble when editing specific text in document?

## Project Rationale
Wildfires have been increasing in frequency and severity across the U.S. primarily in the West. While the direct impacts of worsening wildfires have been extensively measured and reported both in the academic literature and the popular media, the indirect, or secondary effects are not as well understood. One of these secondary effects is the production of widespread wildfire smoke. Smoke, which is a mixture of gases and particulate matter, poses a considerable risk to human health. Because of these health risks, people who live in areas affected by smoke post-wildfire are frequently cautioned against recreating outdoors following fire events. It is anecdotally understood that individuals abide by these warnings and limit time spent outdoors when large amounts of ambient smoke are present. We do not understand however, if individuals decide to reduce time spent outdoors as a result of top-down warnings like the Air Quality Index, or if they make these decisions through individual observation of atmospheric smoke, which we can measure with Dust Surface Mass Concentration. Furthermore, the exact relationship between wildfire smoke and people's decision making is unclear. We do not, for example, know if the effects are linear or if there is a threshold after which the majority of individuals will decide to stay indoors, but not before.

To get an estimate of how populations behave as a result of wildfire smoke, we will use the visitation records collected by the National Park Service. We will couple these data with spatial data on the Dust Surface Mass Concentration at or below 2.5, the primary health concern generated by wildfire smoke. These spatial data were obtained from NASA's Modern-Era Retrospective analysis for Research and Applications version 2 (MERRA-2) satellite. We will also attempt to disentangle these actual atmospheric smoke estimates from the Air Quality Indexes produced by the National Weather Service. This will allow us to estimate to what degree populations base their recreation decisions on top-down or bottom-up influences.

## Hypotheses

### H0
We observe no difference in national park visitation as a result of actual wildfire smoke recorded by the MERRA-2 satellite or Air Quality Indexes reported by the National Weather Service.

**Inference:** Individuals do not change their outdoor recreation behavior as a result of wildfire smoke.

### H1
We observe significantly lower reported park visitation during months with high levels of wildfire smoke recorded by the MERRA-2 satellite, but not during months with dangerous Air Quality Indexes reported by the National Weather Service.

**Inference:** Individuals significantly change their outdoor recreation behavior as a result of wildfire smoke based on personal observations and bottom-up decision making.

### H2
We observe significantly lower reported park visitation during months with dangerous Air Quality Indexes reported by the National Weather Service but not during months with high levels of wildfire smoke recorded by the MERRA-2 satellite.

**Inference:** Individuals significantly change their outdoor recreation behavior as a result of wildfire smoke based on public warnings and top-down influences.

### 2020-03-26 Markdown Documents/Markdown_Questions.Rmd
x

---
title: "Markdown/Github Questions for Matt Williamson"
author: "Matt Clark and Alex Killion"
date: "11/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Questions
1. Standard process for organizing citations using markdown.
2. Highlighting and commenting on text (analog to Word comment bubble)
3. How to work at the same time without creating merg conflicts
4. Should we be using .md files for writing, and only .rmd for anaylsis? Is there an advantage to .md?

### Answers
1.
2. I think you can do the comment stuff from the webpage using the 'inline comment' comment bit on the webpage
3. As far as writing at the same time, I'd use local branches (i.e., forks) to avoid that. So you create a fork called "IntroMC" and write that while Alex works on a fork called "mehtodsAK" when you're ready to pass it back, issue a pull request and merge te changes back to the master branch

### 2020-03-26 Markdown Documents/smokebkpteeb607.Rmd
x

---
title: "SmokeBreakPointEEB607"
author: "Matt Clark"
date: "3/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rstan)
dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
dat<-dat%>% filter(UnitCode !=c("GRSM"))
x<-dat%>%group_by(UnitCode)%>%summarize(max=max(RecreationVisits))%>%arrange(desc(max))
x<-x[1:8,]$UnitCode
dat<-dat%>% filter(UnitCode !=c("GRCA"))
dat<-dat%>% filter(UnitCode !=c("JEFF"))
dat<-dat%>% filter(UnitCode !=c("ACAD"))
dat<-dat%>% filter(UnitCode !=c("OLYM"))
dat<-dat%>% filter(UnitCode !=c("GLAC"))
dat<-dat%>%filter(UnitCode %in% x)
load("l.rda")
```

## Look at the data

These data have monthly visitation to 60 National Parks from 1980 - 2018. We have also gathered the maximum smoke (pm 2.5) values for each of the park boundaries for each of those months. We want to use these data to assess the impact of wildfire smoke on people's recreation behavior.

For this exercise let's just use 3 national parks for this example.

Here's what the data look like:

```{r }
head(dat)
```

## What we think is going on

We suspect that smoke doesn't actually keep people away from parks until it gets pretty bad, i.e. the effect is non-linear. We want to know what actually causes the drop off in visitation. Is it that people stop coming once it gets smokey enough? Smokey for long enough? When it's publicized enough? When the Air quality indexes get to a certain level? We think we may be able to use a break point model to figure this out.

One way we've started exploring this is by looking at the smoke variable standardized accross all parks vs standardized for each park individually. This may answe the question of whether people change their behavior based on some absolute level of smokiness or based on a deviation from a norm for that area. Here's a visual representaion of both of those.

```{r pressure, echo=FALSE}
ggplot(data=dat,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=FALSE,alpha=0.1)+
  theme_classic()+ggtitle("Standardized Overall")

ggplot(data=dat,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=FALSE,alpha=0.1)+
  theme_classic()+ggtitle("Standardized by Park")

```

## Modeling

It looks to me like the "standardized by park" version might land on one specific break point more so than the overall standardization. We can see if that's true using a break point model.

```{stan output.var='mod', eval = FALSE, tidy = FALSE}

  data {
  int N;
  real smoke[N];
  int count[N];
  int Nprk;
  int pcode[N];
}
parameters {
  real intercept;
  real slope1; //the regression parameters
  real slope2; //the regression parameters
  real<lower=0> phi; //the overdispersion parameters
  real<lower=0> sigma_pr;
  real ran_intercept[Nprk];
  real<lower=-1,upper=2> bkpoint;

}
model {  
  intercept ~ normal(0,1); //prior for the intercept following Gelman 2008
  slope1 ~ cauchy(0,2.5); //prior for the slopes following Gelman 2008
  slope2 ~ cauchy(0,2.5); //prior for the slopes following Gelman 2008
  phi ~ cauchy(0, 3);
  sigma_pr ~normal(0,1);
  bkpoint~normal(0,1);

for(i in 1:N){
if(smoke[i]<bkpoint){
  count[i] ~ neg_binomial_2(exp(intercept+slope1*smoke[i]+ran_intercept[pcode[i]]) ,phi);
}
else{
  count[i] ~ neg_binomial_2(exp(intercept+slope1*bkpoint+(smoke[i]-bkpoint)*slope2+ran_intercept[pcode[i]]) ,phi);
}
}
for(j in 1:Nprk){
  ran_intercept[j]~normal(0,sigma_pr);
}
}

```

```{r}
data_list <- list(
  N = nrow(dat),
  Nprk = length(unique(dat$UnitCode)),
  count = dat$RecreationVisits,
  smoke = dat$stdsmokepark,
  pcode = as.numeric(as.factor(dat$UnitCode )))

options(mc.cores=3)
```
```{r,eval=FALSE}
l<- stan( file="SmokePointBreak.stan" , data=data_list,chains=3 )

```

## Output
```{r}
print( l , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
```

We can see that the first slope is positive and the second is negative. This is what we might expect based on the visualization. It also tells us that the bkpoint is at -0.35 standard deviations below the mean which is weird. Not sure how to interpret that.

## Next Steps
I want to apply a version of this to all of the parks (or at least most of them).
That data looks a lot messier.

```{r include=FALSE}
datt<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
```

```{r echo=FALSE}
ggplot(data=datt,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=FALSE,alpha=0.1)+
  theme_classic()+ggtitle("Standardized Overall")

ggplot(data=datt,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
  geom_point(alpha=0.2)+
  geom_smooth(color="red",se=FALSE,alpha=0.1)+
  theme_classic()+ggtitle("Standardized by Park")

```

I think what we want to do is link the breakpoints to external events. Totally open to what those might be. We may want to even put time on the X axis and have the breakpoints be wildfires or something? Not sure about the math there.

We'll take any/all help and insights!

Thanks!

### 2020-03-26 Rcode/ClarkEx.R
x

@@ -1,0 +1,28 @@
+library(tidyverse)
+library(rstan)
+library(rethinking)
+raw.dat <- read_csv("~/Smoke_Proj/Data/ParkDataSubset.csv") 
+
+raw.dat<-raw.dat[,-c(1:2)]
+raw.dat%>% 
+  
+  mutate(season = rethinking::coerce_index(.$Season),
+         unit = rethinking::coerce_index(.$UnitCode)) 
+
+
+
+
+stan.dat <- list(N = nrow(raw.dat),
+                units = length(unique(raw.dat$unit)),
+                seasons = length(unique(raw.dat$season)),
+                unitid = raw.dat$unit,
+                season = raw.dat$season,
+                visits = raw.dat$RecreationVisits,
+                smoke = raw.dat$stdsmoke)
+
+
+
+
+mod <- stan_model("./Downloads/clarkmodel.stan")
+
+fit <- sampling(mod, data=stan.dat, chains = 3)

### 2020-03-26 Rcode/EcoModClass.R

@@ -1,0 +1,36 @@
+library(tidyverse)
+library(rstan)
+dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+dat<-dat%>% filter(UnitCode !=c("GRSM"))
+x<-dat%>%group_by(UnitCode)%>%summarize(max=max(RecreationVisits))%>%arrange(desc(max))
+x<-x[1:8,]$UnitCode
+dat<-dat%>% filter(UnitCode !=c("GRCA"))
+dat<-dat%>% filter(UnitCode !=c("JEFF"))
+dat<-dat%>% filter(UnitCode !=c("ACAD"))
+dat<-dat%>% filter(UnitCode !=c("OLYM"))
+dat<-dat%>% filter(UnitCode !=c("GLAC"))
+dat<-dat%>%filter(UnitCode %in% x)
+
+ggplot(data=dat,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
+  geom_point(alpha=0.2)+
+  geom_smooth(color="red",se=FALSE,alpha=0.1)+
+  theme_classic()+ggtitle("Standardized Overall")
+
+ggplot(data=dat,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
+  geom_point(alpha=0.2)+
+  geom_smooth(color="red",se=FALSE,alpha=0.1)+
+  theme_classic()+ggtitle("Standardized by Park")
+
+
+data_list <- list(
+  N = nrow(dat),
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$RecreationVisits,
+  smoke = dat$stdsmokepark,
+  pcode = as.numeric(as.factor(dat$UnitCode )))
+
+options(mc.cores=3)
+
+l<- stan( file="SmokePointBreak.stan" , data=data_list,chains=3 )
+
+print( l , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )

### 2020-03-26 Rcode/MeanSmokePlots.R

@@ -1,0 +1,68 @@
+library(tidyverse)
+dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+
+
+ggplot(data=dat,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
+  geom_point(alpha=0.2)+
+  geom_smooth(method="lm",color="red",se=FALSE)+
+  theme_classic()+ggtitle("max")
+
+ggplot(data=dat,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
+  geom_point(alpha=0.2)+
+  geom_smooth(color="red",se=FALSE,alpha=0.1)+
+  theme_classic()+ggtitle("max")
+
+
+
+
+ggplot(data=dat,aes(x=Year,y=stdsmoke,by=UnitCode))+
+  geom_jitter(alpha=0.2)+
+  geom_smooth(method="lm",color="red",se=FALSE,alpha=0.2)+
+  theme_classic()
+
+ggplot(data=dat,aes(x=Year,y=RecreationVisits,by=UnitCode))+
+  geom_jitter(alpha=0.2)+
+  geom_smooth(method="lm",color="red",se=FALSE,alpha=0.2)+
+  theme_classic()
+
+
+
+dat$LE<- ifelse(datmax$stdsmoke >=2,"True","False")
+
+ggplot(data=dat,aes(x=LE,y=RecreationVisits))+
+  geom_boxplot()+
+  theme_classic()
+
+
+stdize<-function(x) {return((x-mean(x)/(2*sd(x))))}
+
+dat<-dat%>%group_by(UnitCode)%>%
+  mutate(stdsmokepark=stdize(stdsmoke))
+
+ggplot(data=datmax,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
+  geom_point(alpha=0.2)+
+  geom_smooth(method="lm",color="red",se=FALSE,alpha=0.2)+
+  theme_classic()+ggtitle("max")
+
+ggplot(data=datmax,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
+  geom_point(alpha=0.2)+
+  geom_smooth(color="red",se=FALSE,alpha=0.2)+
+  theme_classic()+ggtitle("max")
+
+
+#stop here
+
+library(zoo)
+dat$Date<-zoo::as.yearmon(paste(as.numeric(dat$Year),
[…]

### 2020-03-26 Rcode/MergeSmokeVisData.R

@@ -1,0 +1,64 @@
+library(tidyverse)
+SmokeDat<-read_csv("~/Smoke_Proj/interm_data/final_max.csv")
+View(SmokeDat)
+VisDat<-read_csv("~/Smoke_Proj/Data/VisitationDataClean.csv")
+View(VisDat)
+
+library(tidyr)
+
+#Make a "long" version of the smoke data
+sml<-gather(SmokeDat, Date, Smoke, X198001:X201910, factor_key = F)
+View(sml)
+
+#Change the format of the dates to match the vis data
+sml$Year<-substr(sml$Date, 2,5)
+sml$Month<-substr(sml$Date, 6,7)
+
+#create a Jan-Sept vector
+x<-1:9
+#Make it a character so it's a btit easier
+VisDat$Month<-as.character(VisDat$Month)
+#add a 0 to the front of the single diget months to match the smoke data
+VisDat$Month<-ifelse(VisDat$Month %in% x,
+       paste0("0", VisDat$Month),paste0("", VisDat$Month)) 
+VisDat$Year<-as.character(VisDat$Year)
+
+#Match the names of the parks
+names(sml)[1]<-"UnitCode"
+
+#merge the datasets by park, and month
+mdat<-merge(x=VisDat,y=sml,by=c("UnitCode", "Year","Month"))
+#remove columns we dont need            
+mdat<-mdat[,-c(4,7)]
+
+#Create a standardized smoke variable
+
+##standardization function doing some weird rounding thing
+## Doing it in 2 steps doesn't yeild the same problem
+zz<-mdat$Smoke-mean(mdat$Smoke)
+mdat$stdsmoke<-zz/(2*sd(mdat$Smoke))
+
+#make month categorical
+mdat$Month<-as.character(mdat$Month)
+#Create a season variable
+mdat$Season<-ifelse(mdat$Month %in% c("03","04","05"),"Spring",
+                   ifelse(mdat$Month %in% c("06","07","08"),"Summer",
+                          ifelse(mdat$Month %in% c("09","10","11"),"Fall","Winter")))
+
+#make combined variables for park/season and park/month
+mdat$CatColS<-paste0(mdat$UnitCode,mdat$Season)
+
+mdat$CatColM<-paste0(mdat$UnitCode,mdat$Month)
+
+#create a variable for stdized smoke by park
+sdd1<-function(x) {return(x-mean(x))}
[…]

### 2020-03-26 Rcode/Merged_data_Mean_smoke.R

@@ -1,0 +1,53 @@
+library(tidyverse)
+SmokeDat<-read_csv("~/Smoke_Proj/interm_data/final_means.csv")
+View(SmokeDat)
+VisDat<-read_csv("~/Smoke_Proj/Data/VisitationDataClean.csv")
+View(VisDat)
+
+library(tidyr)
+
+#Make a "long" version of the smoke data
+sml<-gather(SmokeDat, Date, Smoke, X198001:X201910, factor_key = F)
+View(sml)
+
+#Change the format of the dates to match the vis data
+sml$Year<-substr(sml$Date, 2,5)
+sml$Month<-substr(sml$Date, 6,7)
+
+#create a Jan-Sept vector
+x<-1:9
+#Make it a character so it's a btit easier
+VisDat$Month<-as.character(VisDat$Month)
+#add a 0 to the front of the single diget months to match the smoke data
+VisDat$Month<-ifelse(VisDat$Month %in% x,
+                     paste0("0", VisDat$Month),paste0("", VisDat$Month)) 
+VisDat$Year<-as.character(VisDat$Year)
+
+#Match the names of the parks
+names(sml)[1]<-"UnitCode"
+
+#merge the datasets by park, and month
+mdat<-merge(x=VisDat,y=sml,by=c("UnitCode", "Year","Month"))
+#remove columns we dont need            
+mdat<-mdat[,-c(4,7)]
+
+#Create a standardized smoke variable
+
+##standardization function doing some weird rounding thing
+## Doing it in 2 steps doesn't yeild the same problem
+zz<-dat$Smoke-mean(mdat$Smoke)
+mdat$stdsmoke<-zz/(2*sd(mdat$Smoke))
+
+#make month categorical
+mdat$Month<-as.character(mdat$Month)
+#Create a season variable
+mdat$Season<-ifelse(mdat$Month %in% c("03","04","05"),"Spring",
+                    ifelse(mdat$Month %in% c("06","07","08"),"Summer",
+                           ifelse(mdat$Month %in% c("09","10","11"),"Fall","Winter")))
+
+#make combined variables for park/season and park/month
+mdat$CatColS<-paste0(mdat$UnitCode,mdat$Season)
+
+mdat$CatColM<-paste0(mdat$UnitCode,mdat$Month)
+#create the csv
+write.csv(mdat, file="MergedDataCompleteMeanSmoke.csv")

### 2020-03-26 Rcode/NegBinomRanInterceptRanSlopeModelsDone.R

@@ -1,0 +1,60 @@
+dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+x<-dat%>%group_by(UnitCode,Season)%>%
+  summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(mv==max(mv))%>%mutate(cc=paste0(UnitCode,Season))
+
+x2<-dat%>%group_by(UnitCode,Season)%>%
+  summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(mv==min(mv))%>%mutate(cc=paste0(UnitCode,Season))
+
+dat$SeasType<-ifelse(dat$CatColS %in% x$cc,
+                     "High",ifelse(dat$CatColS %in% x2$cc,"Low","Shoulder"))
+
+dat<-dat%>%filter(SeasType =="High")
+
+data_list <- list(
+  N = nrow(dat),
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$RecreationVisits,
+  smoke = dat$stdsmokepark,
+  pcode = as.numeric(as.factor(dat$UnitCode )))
+
+options(mc.cores=3)
+
+l<- stan( file="SmokeNegBinom.stan" , data=data_list,chains=3 ,warmup = 1000 ,iter = 10000)
+
+print( l , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
+
+save(l, file="stdsmokeparkNbinom.rda")
+
+data_list2 <- list(
+  N = nrow(dat),
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$RecreationVisits,
+  smoke = dat$stdsmoke,
+  pcode = as.numeric(as.factor(dat$UnitCode )))
+
+l2<- stan( file="SmokeNegBinom.stan" , data=data_list2,chains=3 ,warmup=1000 ,iter = 6000)
+print( l2 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
+
+save(l2, file="stdsmokeNbinom.rda")
+
+
+#Plot outputs
+x<-as.data.frame(summary(l))
+stan_plot(l,pars = c(zz) ,fill_color = "purple", )
+stan_plot(l2,pars = c(zz) ,fill_color = "purple", )
+########Compare outputs to overall smokiness
+datAll<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
[…]

### 2020-03-26 Rcode/RegressionFirstCut.R

@@ -1,0 +1,76 @@
+library(tidyverse)
+library(lme4)
+library(rstanarm)
+library(MASS)
+options(mc.cores = parallel::detectCores())
+dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+dat$Year<-as.factor(dat$Year)
+#models below increase in complexity..Currently, non of  them are converging using 
+#the lme4 package or with rstanarm
+#this may be a computational problem on my end though, rather than a model specification problem
+
+#glm with independent intercept for each park for each season
+fit<-glmer.nb(RecreationVisits~stdsmoke+CatColS+(1|CatColS),data=dat)
+
+#glm with independent intercept for each park for each Month
+fit2<-glmer.nb(RecreationVisits~stdsmoke+CatColM+(1|CatColM),data=dat)
+
+
+#glm with independent intercept for each park for each Season and
+# random slope for smoke for each park
+fit3<-glmer.nb(RecreationVisits~stdsmoke+CatColS+(stdsmoke|CatColS),data=dat)
+
+#glm with independent intercept for each park for each Month and
+# random slope for smoke for each park
+#this is the preferred model
+fit4<-glmer.nb(RecreationVisits~stdsmoke+CatColM+(stdsmoke|CatColM),data=dat)
+
+
+#any of these models can be run using stan by specifying 
+#stan_glmer.nb instead of the glmer.nb function from lme4
+
+
+fit5<-glmer.nb(RecreationVisits~(1|CatColS)+(stdsmoke|UnitCode), data=dat)
+save(fit5,file="fit5.rda")
+summary(fit5)
+
+#best
+fit6<-glmer.nb(RecreationVisits~(1|CatColM)+(stdsmoke|UnitCode), data=dat)
+save(fit6,file="fit6.rda")
+
+fit7<-glmer.nb(RecreationVisits~(stdsmoke|CatColM), data=dat)
+save(fit7,file="fit7.rda")
+
+fit8<-glm.nb(RecreationVisits~stdsmoke+Month+UnitCode,data=dat)
+save(fit8,file="fit8.rda")
+
+fit9<-glm.nb(RecreationVisits~stdsmoke+Season+UnitCode,data=dat)
+save(fit9,file="fit9.rda")
+
+
+coef(fit6)
+
+fit<-glmer.nb(RecreationVisits~
+                     (1|CatColM)+
+                     (stdsmoke|UnitCode)+
+                     Year,
+                   data=dat)
+
+
+
+
[…]

### 2020-03-26 Rcode/Sloppy plots and models.R

@@ -1,0 +1,63 @@
+library(tidyverse)
+library(lme4)
+library(rstanarm)
+library(MASS)
+dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+
+fit2<-glmer.nb(RecreationVisits~stdsmoke+(Season|UnitCode),data=dat)
+
+fit3<-glmer.nb(RecreationVisits~ScaledSmoke+(1|CatColM),data=dat)
+fit4<-glm(RecreationVisits~Smoke,family="poisson",data=dat)
+
+newdata <- data.frame(
+  ScaledSmoke = seq(from = min(dat[dat$CatColM=="YOSE07",]$Smoke), to = max(dat[dat$CatColM=="YOSE07",]$Smoke), length.out = 100),
+  CatColM = rep("YOSE07",100))
+  
+newdata2 <- cbind(newdata, predict(fit3, newdata))
+
+nd<-data.frame(Smoke=0.00316, CatColM="ARCH06")
+predict(fit2, nd)
+
+
+
+ggplot(data=dat[1:6959,],aes(x=Smoke,y=RecreationVisits))+
+  geom_smooth(aes(color=Month),method="lm", se=F)+
+facet_wrap(~ParkName, scales="free")
+  
+ggplot(data=dat[6960:13955,],aes(x=Smoke,y=RecreationVisits))+
+  geom_smooth(aes(color=Month),method="lm", se=F)+
+  facet_wrap(~ParkName,, scales="free")
+
+ggplot(data=dat[13956:20891,],aes(x=Smoke,y=RecreationVisits))+
+  geom_smooth(aes(color=Month),method="lm", se=F)+
+  facet_wrap(~ParkName,, scales="free")
+
+ggplot(data=dat[20892:nrow(dat),],aes(x=Smoke,y=RecreationVisits))+
+  geom_smooth(aes(color=Month),method="lm", se=F)+
+  facet_wrap(~ParkName, scales="free")
+  
+ggplot(data=dat[dat$UnitCode %in% c("YOSE", "YELL","ACAD","GRSM","JOTR","SEQU"),],aes(x=Smoke,y=RecreationVisits))+
+  geom_smooth(aes(color=Month),method="lm", se=F)+
+  geom_point(aes(color=Month))+facet_wrap(~ParkName, scales="free")
+  
[…]

### 2020-03-26 Rcode/SmokePointBreak.R

@@ -1,0 +1,104 @@
+library(tidyverse)
+library(rstan)
+dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+dat<-dat%>% filter(UnitCode !=c("GRSM"))
+x<-dat%>%group_by(UnitCode)%>%summarize(max=max(RecreationVisits))%>%arrange(desc(max))
+x<-x[1:6,]$UnitCode
+dat<-dat%>%filter(UnitCode %in% x)
+
+dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+x<-dat%>%group_by(UnitCode,Season)%>%
+  summarise(mv=max(RecreationVisits))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(mv==max(mv))%>%mutate(cc=paste0(UnitCode,Season))
+
+x2<-dat%>%group_by(UnitCode,Season)%>%
+  summarise(mv=min(RecreationVisits))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(mv==min(mv))%>%mutate(cc=paste0(UnitCode,Season))
+
+dat$SeasType<-ifelse(dat$CatColS %in% x$cc,
+                     "High",ifelse(dat$CatColS %in% x2$cc,"Low","Shoulder"))
+
+dat<-dat%>%filter(SeasType =="High")
+
+data_list <- list(
+  N = nrow(dat),
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$RecreationVisits,
+  smoke = dat$stdsmokepark,
+  pcode = as.numeric(as.factor(dat$UnitCode )))
+
+options(mc.cores=3)
+
+l<- stan( file="SmokePointBreak.stan" , data=data_list,chains=3 )
+
+print( l , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
+
+
+
+data_list <- list(
+  N = nrow(dat),
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$RecreationVisits,
+  smoke = dat$stdsmokepark,
+  pcode = as.numeric(as.factor(dat$UnitCode )),
+  NSPint = length(unique(dat$CatColS)),
+  SPint = as.numeric(as.factor(dat$CatColS ))
+  )
+
+options(mc.cores=3)
[…]

### 2020-03-26 Rcode/TempStdMod.R

@@ -1,0 +1,39 @@
+dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+x<-dat%>%group_by(UnitCode,Season)%>%
+  summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(mv==max(mv))%>%mutate(cc=paste0(UnitCode,Season))
+
+x2<-dat%>%group_by(UnitCode,Season)%>%
+  summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(mv==min(mv))%>%mutate(cc=paste0(UnitCode,Season))
+
+dat$SeasType<-ifelse(dat$CatColS %in% x$cc,
+                     "High",ifelse(dat$CatColS %in% x2$cc,"Low","Shoulder"))
+
+stdize<-function(x) {return((x-mean(x)/(2*sd(x))))}
+
+dat<-dat%>%group_by(as.character(Year))%>%mutate(stdsmoketime=stdize(Smoke*10e8))
+
+
+dat<-dat%>%filter(SeasType =="High")
+
+data_list <- list(
+  N = nrow(dat),
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$RecreationVisits,
+  smoke = dat$stdsmoketime,
+  pcode = as.numeric(as.factor(dat$UnitCode )))
+
+options(mc.cores=3)
+
+l3<- stan( file="SmokeNegBinom.stan" , data=data_list,chains=3 ,warmup = 1000 ,iter = 10000)
+
+print( l3 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
+
+save(l3, file="stdsmokeTimeNbinom.rda")
+stan_plot(l3,pars = c(zz) ,fill_color = "purple", )
+
+mean(as.data.frame(summary(l3))[2:61,1])
+
+
+

### 2020-03-26 Rcode/VidDataCleaning.R

@@ -1,0 +1,12 @@
+library(tidyverse)
+VisDat<-read_csv("~/Smoke_Proj/Data/VisDat_NoHeader.csv")
+View(VisDat)
+
+#this data is massive.
+#Let's chop it down a bit
+names(VisDat)
+VisDat<-VisDat %>% 
+  select(ParkName,UnitCode,Year,Month,RecreationVisits)
+
+write.csv(VisDat, file="~/Smoke_Proj/Data/VisitationDataClean.csv")
+nrow(VisDat)

### 2020-03-26 Rcode/scrappyMods.R

@@ -1,0 +1,33 @@
+library(tidyverse)
+library(lme4)
+library(rstanarm)
+library(MASS)
+
+dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+dat$Year<-as.factor(dat$Year)
+dat$catY<-paste0(dat$UnitCode,dat$Year)
+MattDat<-dat[seq(1,nrow(dat),10),]
+
+
+fitM<-stan_glmer.nb(RecreationVisits~(stdsmoke|UnitCode)+(Season|catY), chains=1,data=MattDat)
+coef(fitM)
+
+fitM<-stan_glmer(RecreationVisits~(stdsmoke|UnitCode)+(Season|catY),
+                 family="poisson",chains=1,data=MattDat)
+coef(fitM)
+fitz<-stan_glmer(RecreationVisits~(stdsmoke|UnitCode)+
+                      (UnitCode:Season|Year), chains=1,family="poisson",data=MattDat)
+
+MattDat<-dat[seq(1,nrow(dat),10),]
+
+
+
+
+
+
+
+
+
+
+
+

### 2020-03-28 Rcode/NegBinomRanInterceptRanSlopeModelsDone.R
x

@@ -1,4 +1,4 @@
-dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+dat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")
 x<-dat%>%group_by(UnitCode,Season)%>%
   summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
   group_by(UnitCode)%>%filter(mv==max(mv))%>%mutate(cc=paste0(UnitCode,Season))

### 2020-03-28 Rcode/NegBinomRanInterceptRanSlopeModelsDone.R
model runs now

@@ -1,3 +1,6 @@
+library(tidyverse)
+library(rstan)
+library(shinystan)
 dat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")
 x<-dat%>%group_by(UnitCode,Season)%>%
   summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
@@ -21,7 +24,7 @@ data_list <- list(
 
 options(mc.cores=3)
 
-l<- stan( file="SmokeNegBinom.stan" , data=data_list,chains=3 ,warmup = 1000 ,iter = 10000)
+l<- stan( file="~/SmokeProject/StanCode/SmokeNegBinom.stan" , data=data_list,chains=3 ,warmup = 1000 ,iter = 10000)
 
 print( l , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
 
@@ -34,7 +37,7 @@ data_list2 <- list(
   smoke = dat$stdsmoke,
   pcode = as.numeric(as.factor(dat$UnitCode )))
 
-l2<- stan( file="SmokeNegBinom.stan" , data=data_list2,chains=3 ,warmup=1000 ,iter = 6000)
+l2<- stan( file="~/SmokeProject/StanCode/SmokeNegBinom.stan" , data=data_list2,chains=3 ,warmup=1000 ,iter = 6000)
 print( l2 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
 
 save(l2, file="stdsmokeNbinom.rda")
@@ -42,10 +45,11 @@ save(l2, file="stdsmokeNbinom.rda")
 
 #Plot outputs
 x<-as.data.frame(summary(l))
+zz<-row.names(x[2:61,])
 stan_plot(l,pars = c(zz) ,fill_color = "purple", )
 stan_plot(l2,pars = c(zz) ,fill_color = "purple", )
 ########Compare outputs to overall smokiness
-datAll<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
+datAll<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")
 MeanSlopes<-data.frame(Park=unique(dat$UnitCode),slopeSmokePark=(as.data.frame(summary(l))[2:61,1]),
                        slopeSmokeOverall=(as.data.frame(summary(l2))[2:61,1]),
                        MedSmoke= (summarise(group_by(dat,UnitCode),MedSmoke=median(stdsmoke)))$MedSmoke,
[…]

### 2020-03-30 Laura_Stoddard_VIP_Timeline.Rmd
markdown document with instructions for laura

---
title: "Laura Stoddard VIP Timeline"
author: "Matt Clark"
date: "3/30/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rstanarm)

load("feefit.rda")
SMdat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")
FEdat<-read_csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
```

Schedule
======

| Week| Focus |Deliverable
:---:|:---:|:---
3/30/2020 | Data Cleaning | Cleaned up dataset with projected fees for relevant parks
4/6/2020 | Plotting 1| Previous smoke effects plot
4/13/2020 | Plotting 2| Projected smoke effects plot
4/20/2020 | Write-up / Catch-up| Brief write-up of results
4/27/2020 | Lightening talk | Lightening talk

Weekly Specifics
======

3/30/2020 - Data Cleaning
------

This week, it would be great if you could clean up the dataset that you collected and fill in what we think the entry fees may have been for each park for the years we don't have data on.

To do this, you will need to do 4 main steps.

1. Upload the data
2. Remove rows that we don't know
3. Make a very simple model
4. Use that model to predict what the unknown fees likely would have been

I'll give you some guidance on how to do all 4 of these steps, but I do kind of want you to struggle through some of it on your own. Here's an example of how you would do this for just 2 parks for 2 years. I want you to take these steps and apply them to all parks for all the years we don't know

```{r}
library(tidyverse)
FEdat<-read_csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv") #you can call this whatever you want. Keep in mind that your file path will be different than mine!
head(FEdat) #make sure the data came through

```

Now that we've loaded the data, we want to remove the single vehicle column. We aren't using it yet and it has rows with NAs that the single visitor column doesn't. We will filter by NAs later so we want this column gone.

```{r}
d<-FEdat %>% dplyr::select(Park,Year,SingleVisitorEntryFee) #keep only these columns
d<-na.omit(d) #get rid of rows with any NA
head(d) #check it out
```

Now we're going to make a model based on these data. Think of this as y=mx+b that you learned in algebra. Y is the predicted fee, x is the year, m is the slope, and b is an intercept. In this case we also have a x2 and m2 which are a 1 or 0 for each park and a slope (m2) for each park.

This is what's called an generalized linear model. The "poisson" part just tells the computer that we are modeling values that are discrete counts (1,4,7,etc.)

```{r, eval=FALSE}
feefit<-stan_glm(SingleVisitorEntryFee~Year+Park,family="poisson",data=d,iter=6000)
```

Now we can use this model to predict the fees for the parks we don't know.

Here's an example of that

```{r}
preddata<-data.frame(Year = c(2012,2016), Park = c("Arches NP", "Acadia NP")) #make a small, 2 row data frame

apply(posterior_predict(feefit,preddata),2,mean)     #predict our fees for thse parks for those years based on our model.

```
Notice that this gives predictions for those two parks for those years.

We want predictions for every row that we are missing. We definitely don't want to type this in by hand.

On your own, subset the original datframe by just the rows which have NA in the fee column. Use this new dataframe in the place of preddat in the example.

#### Hint: 
```{r eval=FALSE}
NAdata<-data%>%dplyr::filter(fee == NA)

apply(posterior_predict(feefit,NAdata),2,mean)  
```
You should then be able to add this output as a column to the dataframe with all nas that you made.

#### Hint: 
```{r eval=FALSE}
NAdata$PredictedVis<-apply(posterior_predict(feefit,NAdata),2,mean) 
```

Ideally, you will then use R to fill in all of the nas in the original data frame with our predicted values. You should also make a new column indicating if the value is real or a projection.

#### Hint: 
```{r eval=FALSE}
merge(FEdata,NAdata, by=1:2)

```

### 2020-03-30 Rcode/NegBinomRanInterceptRanSlopeModelsDone.R
param outputs as csv

@@ -44,6 +44,8 @@ save(l2, file="stdsmokeNbinom.rda")
 
 
 #Plot outputs
+load("stdsmokeparkNbinom.rda")
+load("stdsmokeNbinom.rda")
 x<-as.data.frame(summary(l))
 zz<-row.names(x[2:61,])
 stan_plot(l,pars = c(zz) ,fill_color = "purple", )
@@ -54,6 +56,8 @@ MeanSlopes<-data.frame(Park=unique(dat$UnitCode),slopeSmokePark=(as.data.frame(s
                        slopeSmokeOverall=(as.data.frame(summary(l2))[2:61,1]),
                        MedSmoke= (summarise(group_by(dat,UnitCode),MedSmoke=median(stdsmoke)))$MedSmoke,
                        MedSmokeAll= (summarise(group_by(datAll,UnitCode),MedSmoke=median(stdsmoke)))$MedSmoke) 
+
+write.csv(MeanSlopes,file="ParameterOutputs")
 
 cor(MeanSlopes$MedSmokeAll,MeanSlopes$slopeSmokeOverall)
 ggplot(MeanSlopes,aes(x=MedSmokeAll,y=slopeSmokeOverall))+

### 2020-03-30 Rcode/MergeSmokeVisData.R
code for making main df

@@ -60,5 +60,17 @@ mdat<-mdat%>%group_by(UnitCode)%>%
 
 #delete int vector
 mdat<-mdat[,-11]
+
+#add vis season
+x<-mdat%>%group_by(UnitCode,Season)%>%
+  summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(mv==max(mv))%>%mutate(cc=paste0(UnitCode,Season))
+
+x2<-mdat%>%group_by(UnitCode,Season)%>%
+  summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(mv==min(mv))%>%mutate(cc=paste0(UnitCode,Season))
+
+mdat$SeasType<-ifelse(mdat$CatColS %in% x$cc,
+                     "High",ifelse(mdat$CatColS %in% x2$cc,"Low","Shoulder"))
 #create the csv
 write.csv(mdat, file="MergedDataComplete.csv")

### 2020-03-30 Rcode/NegBinomRanInterceptRanSlopeModelsDone.R
x

@@ -57,7 +57,7 @@ MeanSlopes<-data.frame(Park=unique(dat$UnitCode),slopeSmokePark=(as.data.frame(s
                        MedSmoke= (summarise(group_by(dat,UnitCode),MedSmoke=median(stdsmoke)))$MedSmoke,
                        MedSmokeAll= (summarise(group_by(datAll,UnitCode),MedSmoke=median(stdsmoke)))$MedSmoke) 
 
-write.csv(MeanSlopes,file="ParameterOutputs")
+write.csv(MeanSlopes,file="ParameterOutputs.csv")
 
 cor(MeanSlopes$MedSmokeAll,MeanSlopes$slopeSmokeOverall)
 ggplot(MeanSlopes,aes(x=MedSmokeAll,y=slopeSmokeOverall))+

### 2020-04-27 Intro_Background.Rmd
moved a bunch of markdown documents

---
title: "Introduction/Background"
author: "Matt Clark & Alex Killion"
date: "4/27/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Intro/ Background
Wildfires have been increasing in frequency and severity across the U.S. primarily in the West. While the direct impacts of worsening wildfires have been extensively measured and reported both in the academic literature and the popular media, the indirect, or secondary effects are not as well understood. One of these secondary effects is the production of widespread wildfire smoke. Smoke, which is a mixture of gases and particulate matter, poses a considerable risk to human health. Because of these health risks, people who live in areas affected by wildfire smoke are frequently cautioned against recreating outdoors following fire events. It is anecdotally understood that individuals abide by these warnings and limit time spent outdoors when large amounts of ambient smoke are present. We do not however, currently understand the exact relationship between wildfire smoke and population level recreation behavior. For example, we do not know if the effects are linear or if there is a threshold after which the majority of individuals will decide to stay indoors, but not before. Furthermore, the theory of 'shifting baselines' suggests that individuals may tolerate normally intolerable degrees of environmental disturbance when the disturbance is framed against an already highly disturbed location or point in time. In other words, human response to envionmetal disturbance is relative.

To get an estimate of how populations behave as a result of wildfire smoke, we will use the visitation records collected by the National Park Service. We will couple these data with spatial data on the Dust Surface Mass Concentration at or below 2.5, the primary health concern generated by wildfire smoke. These spatial data were obtained from NASA's Modern-Era Retrospective analysis for Research and Applications version 2 (MERRA-2) satellite.

Increasing wildfire severity has further garnered attention for the financial burden it poses to the federal government. The United States spends approximately $1 billion every year to fight wildfires. This however, only approximates the direct costs of wildfire relief. Given that outdoor recreation is significantly affected by wildfire smoke, a secondary goal of this project will be to generate an order of magnitude estimate on National Park Service revenue lost due to increased wildfire smoke.

## 4/27 Meeting Notes

* Top 10 most visited parks?
* Remove relativelysmokeless parks? cutoff?
* Keep only parks where smoke season overlaps with high visitation season
* Selection bias? 
  + Keep parks in areas where wildfires are expected to increase?
  + Stratify on the independent variable, not the depemdent
  + Design analysis to ensure that the experimental treatment behaves consistently

* How is increasing smoke affecting parks in smoke vulnerable places? How is this expected to continue into the future?

Parks are the crown jewl of the america public lands system. They are already financially burdened. Think of pictures of overwhelmed trashcans, bears, condors eating trash, etc. One of the biggest things that affect them financially negatively and they have no control over is smoke. They don't control surrounding forest service land etc. Climate change might affect the financial strain on the park service, here's how?

* Path Forward
1. subset highsmoke/high season
2. varying effects of smoke and distance from airport, varying intercept for each park, varying slope on AR term
3. Alex will look for future fire vulnerability 
4. Alex will look into distance from airport

### 2020-04-27 Markdown Documents/Hypotheses_Rationale.Rmd
moved a bunch of markdown documents

@@
 ```
-## Testing - What is the best way to replicate a Word comment bubble when editing specific text in document?
 
@@
 
+Increasing wildfire severity has further garnered attention for the financial burden it poses to the federal government.
+The United States spends approximately $1 billion every year to fight wildfires.
+This however, only approximates the direct costs of wildfire relief.
+Given that outdoor recreation is significantly affected by wildfire smoke, a secondary goal of this project will be to generate an order of magnitude estimate on National Park Service revenue lost due to increased wildfire smoke.
+
 ## Hypotheses

### 2020-04-27 Markdown Documents/Laura_Stoddard_VIP_Timeline.Rmd
moved a bunch of markdown documents

--- 2020-03-30 Laura_Stoddard_VIP_Timeline.Rmd
@@
 
-load("feefit.rda")
+load("~/SmokeProject/ModelObjects/feefit.rda")
 SMdat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")
@@
 
-This is what's called an generalized linear model. The "poisson" part just tells the computer that we are modeling values that are discrete counts (1,4,7,etc.)
+This is what's called an generalized linear model. The "Poisson" part just tells the computer that we are modeling values that are discrete counts (1,4,7,etc.)
 
 ```{r, eval=FALSE}
-feefit<-stan_glm(SingleVisitorEntryFee~Year+Park,family="poisson",data=d,iter=6000)
+feefit<-stan_glm(SingleVisitorEntryFee~Year+Park,family="Poisson",data=d,iter=6000)
 ```
@@
 ```
+
+4/6/2020 Plotting 1
+------
+
+We will use the package ggplot2 to plot the data that you have collected/extrapolated.
+
+If you remember from before, the ggplot() function follows the [grammar of graphics](https://towardsdatascience.com/a-comprehensive-guide-to-the-grammar-of-graphics-for-effective-visualization-of-multi-dimensional-1f92b4ed4149), a theory for how to visualize data that relies on stacking layers on top of each other.
+
+When you create ggplot() code, the top most code will be the first layer created.
+For our purposed, we will always want this to be the axes of the figure.
+We will do this by specifying the data and the x and y axes.
+Remember that the x and y axes need to be wrapped in an aestetic function aes().
+This just tells ggplot to look for these things in the dataset that we specified.
+
+```{r}
+library(ggplot2) #load the package
+
+ggplot(data= d, aes(x=Year,y=SingleVisitorEntryFee))
+
+```
+
+You'll see that this just "sets the stage" for the rest of the plot. We haven't told it what kind of plot we want yet.
+
+Next, we will layer the data on top of this background.
+
+```{r}
+ggplot(data= d, aes(x=as.character(Year),y=SingleVisitorEntryFee))+
+  geom_boxplot()
+```
+
+Is this the best way to visualize this simple dataset? probably not. Can you think of  better way?
+
+There is a ton of flexibility in ggplot2. See if you can make this plot better.
+
+Once you do that, see if you can plot your data that you created by extrapolating from the fee data we had.
+
+I like to hand draw my plots first so I know what I'm aiming for.
+
+If you want help and inspiration check out the [Cookbook for ggplot](http://www.cookbook-r.com/Graphs/) and the [ggplot gallery](https://www.r-graph-gallery.com/).
+
+4/13/2020 Plotting 2
+------

### 2020-04-27 Markdown Documents/progress4_7.Rmd
moved a bunch of markdown documents

---
title: "4/7 progress update"
author: "Matt Clark"
date: "4/7/2020"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(rstan)
load("~/SmokeProject/ModelObjects/stdsmokeNbinom.rda")
load("~/SmokeProject/ModelObjects/stdsmokeparkNbinom.rda")

```

## Current progress

So to get rid of the noise that was introduced by the low visitation and low smoke months, we subset by each park's high season.

```{r}
dat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")

dat<-dat%>%filter(SeasType =="High")
```

Then I ran the following stan model on both the smoke values standardized overall and the smoke values standardized by park.

```{stan output.var='mod', eval = FALSE, tidy = FALSE}
data {
  int N;
  real smoke[N];
  int count[N];
  int Nprk;
  int pcode[N];
}
parameters {
  real intercept;
  real slope1[Nprk]; //the regression parameters

  real<lower=0> phi; //the overdispersion parameters
  real<lower=0> sigma_pr;
  real ran_intercept[Nprk];

}
model {  
  intercept ~ normal(0,1); //prior for the intercept following Gelman 2008
  slope1 ~ cauchy(0,2.5); //prior for the slopes following Gelman 2008

  phi ~ cauchy(0, 3);
  sigma_pr ~normal(0,1);

for(i in 1:N){

  count[i] ~ neg_binomial_2(exp(intercept+slope1[pcode[i]]*smoke[i]+ran_intercept[pcode[i]]) ,phi);

}
for(j in 1:Nprk){
  ran_intercept[j]~normal(0,sigma_pr);
}
}

```

This model converged for both the smoke variable standardized overall and standardized by park. 
Here are the parameter estimates standardized overall

```{r include=FALSE}
x<-as.data.frame(summary(l))
zz<-row.names(x[2:61,])
```

Standardized overall (mean = -0.55)
```{r echo=FALSE}
stan_plot(l2,pars = c(zz) ,fill_color = "purple", )
```

Standardized by park (mean = -0.107)
```{r echo=FALSE}
stan_plot(l,pars = c(zz) ,fill_color = "purple", )
```

These seem to indicate that we have a stronger effect of smoke overall, rather than a relative effect by park. To double check this, we can look to see if the parameter estimates by park vary in a systematic way with the amount of smoke a park gets.

```{r include=FALSE}
datAll<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")
MeanSlopes<-data.frame(Park=unique(dat$UnitCode),slopeSmokePark=(as.data.frame(summary(l))[2:61,1]),
                       slopeSmokeOverall=(as.data.frame(summary(l2))[2:61,1]),
                       MedSmoke= (summarise(group_by(dat,UnitCode),MedSmoke=median(stdsmoke)))$MedSmoke,
                       MedSmokeAll= (summarise(group_by(datAll,UnitCode),MedSmoke=median(stdsmoke)))$MedSmoke) 
```

```{r echo = FALSE}
ggplot(MeanSlopes,aes(x=MedSmoke,y=slopeSmokeOverall))+
  geom_point()+ggtitle("Median high season smokiness by slope of smoke in high season\nCor = 0.37")
```

This indicates to me that we do not see any effect of shifting baselines. Instead, we see that overall smokiness is an important predictor of visitation, regardless of the average smokiness of the place.

## Next Steps

Now that we have this functioning model and a preliminary result (No evidence of spatial shifting baselines for smoke tolerance), I want to 'backcast' previous visitation to each park given a minimum level of smoke

I decided to do this in R because stan was giving me a ton of trouble. I think this is correct, but it could definitely use a double check.

First I extracted the posteriors from the model:
```{r}
ext_fit<-rstan::extract(l2)

intercept_post<-ext_fit$intercept
slope1_post<-ext_fit$slope1
phi_post<-ext_fit$phi
sigma_pr_post<-ext_fit$sigma_pr
ran_intercept_post<-ext_fit$ran_intercept
lp__post<-ext_fit$lp__

sp<-as.data.frame(slope1_post)
ri<-as.data.frame(ran_intercept_post)

```

Then I wrote a sampeling statement

```{r}
gen_quant_r <- function(x,y) { #x here is the input smoke values #y is park
  lin_comb <- mean(intercept_post) + #mean intercept posterior
    x*mean(sp[,paste0("V",y)])+ #smoke value times the posterior slope for smoke for that park
    mean(ri[,paste0("V",y)]) #random intercept for the given park
  mu <- exp(lin_comb)  
  phi<-mean(phi_post) #mean value of the overdispersion parameter
  dist <- rnbinom(n=length(intercept_post),size= phi, mu=mu) #sample from a negative binomial
  out<-mean(dist) #mean of sample
  return(out)
}
```

```{r include = FALSE}
params<-read_csv("~/SmokeProject/Data/ParameterOutputs.csv")[,-c(1,3,5,6)]
fulldat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")[,-1]
fulldat<-fulldat%>%select(UnitCode,Year,Month,ParkName,Smoke,stdsmoke,RecreationVisits,SeasType)%>%
  filter(SeasType =="High")%>%select(UnitCode,Year,Month,ParkName,Smoke,RecreationVisits,stdsmoke)
names(params)[1]<-"UnitCode"

fulldat<-left_join(fulldat,params,by="UnitCode")
x<-as.data.frame(summary(l2))
x$parameter<-row.names(x)
#get just means for parks who's par estimates don't overlap zero
x<-x[c(2:61),c(1,4,8,32)]%>%filter((summary.2.5. >0 & summary.97.5.>0) |
                    (summary.2.5. <0 & summary.97.5.<0) )

Parks <- as.numeric(str_remove(substr(x$parameter,8,9),"]"))

fulldat$PF<-as.numeric(as.factor(fulldat$UnitCode))
sigdat<-fulldat%>% filter(PF %in% Parks)
```

Now we can test this on the actual data and see if we're close at all.

Note here that the 'sigdat' dataset that we're using is only the parks that had 95% CIs that did not overlap zero 
```{r}
empty_vec <- c()
for(i in 1:nrow(sigdat)){
sigdat2<-sigdat[i,]%>%dplyr::mutate(x=gen_quant_r(stdsmoke,PF))
empty_vec <- c(empty_vec,sigdat2$x)
  }

plot(empty_vec,sigdat$RecreationVisits)
cor(empty_vec,sigdat$RecreationVisits)
```

So that looks ..okay. What if we test it on the minimimum smoke value for each park?

```{r}
sigdat<-sigdat%>%group_by(UnitCode)%>%
  mutate(minsmoke=min(stdsmoke))

predNoSmoke <- c()
for(i in 1:nrow(sigdat)){
  sigdat2<-sigdat[i,]%>%dplyr::mutate(x=gen_quant_r(minsmoke,PF))
  predNoSmoke <- c(predNoSmoke,sigdat2$x)
}
```

```{r include = F}
sigdat$predNoSmoke<-predNoSmoke
sigdat$date<-zoo::as.yearmon(paste0(sigdat$Month,sigdat$Year),"%m%Y")
```

Let's plot the minimimum smoke visitation estimate for each park (green) by the actual visitation (purple) and plot the smoke value on there too (black).

These look really sloppy but I think they get the job done for now.

```{r echo = FALSE}
sigdat%>%filter(UnitCode %in% c("ACAD","BADL","CRLA","CUVA", "GLAC","GLBA","HAVO","KEFJ","MACA","OLYM","ROMO","SHEN","THRO","MORA","KICA"))%>%
ggplot(.,aes(x=date))+
  geom_line(aes(y=RecreationVisits),color="purple",size=1,alpha=0.99)+
  geom_line(aes(y=predNoSmoke),color="green",size=2,alpha=0.5)+
  geom_line(aes(y=Smoke*1e14))+
  facet_wrap(~UnitCode,scales = "free")+
  theme_classic()

sigdat%>%filter(!UnitCode %in% c("ACAD","BADL","CRLA","CUVA", "GLAC","GLBA","HAVO","KEFJ","MACA","OLYM","ROMO","SHEN","THRO","MORA","KICA"))%>%
ggplot(.,aes(x=date))+
  geom_line(aes(y=RecreationVisits),color="purple",size=1,alpha=0.99)+
  geom_line(aes(y=predNoSmoke),color="green",size=2,alpha=0.5)+
  geom_line(aes(y=Smoke*1e13))+
  facet_wrap(~UnitCode,scales = "free")+
  theme_classic()

```

I'm not totally sure where to go from here. These estimates don't look great, probably because visitation is obviously much more complicated than just how smokey it is. Maybe we should add an autoregressive term for each park?

Or maybe we're going too far and the spatial shifting baselines stuff is enough for a paper?

Anyway, I'm curious to hear your thoughts overall and especially on the posterior predictive stuff.

Thanks!!

### 2020-04-27 PlotsForPaper.R

@@ -1,197 +1,0 @@
-library(tidyverse)
-allData<-read_csv("Data/MergedDataComplete.csv")[,-1]
-predDat<-read_csv("~/SmokeProject/Data/VisitationEstimatesMinSmokeSignificantParks.csv")[,-1]
-predDat$date<-zoo::as.yearmon(predDat$date)
-allData$date<-zoo::as.yearmon(paste0(allData$Month,allData$Year),"%m%Y")
-allData<-allData%>% filter(SeasType =="High")
- #all sig parks
-predDat.interp <- predDat %>% 
-   split(.$UnitCode) %>% 
-   map_df(~data.frame(UnitCode = approx(.x$date, .x$No_Smoke_Mean_est, n = 80), 
-                      nat = approx(.x$date, .x$RecreationVisits, n = 80), 
-                      UnitCode = .x$UnitCode[1]))
-
-
-ggplot(predDat.interp, aes(nat.x, nat.y)) +  facet_wrap(~UnitCode,scales="free") +
-  geom_ribbon(aes(ymin = nat.y, ymax = pmin(UnitCode.y, nat.y), fill = "Observed greater"),alpha=0.9) +
-  geom_ribbon(aes(ymin = UnitCode.y, ymax = pmin(UnitCode.y, nat.y), fill = "Predicted greater"),alpha=0.9) +
-  geom_line(aes(linetype = "Observed Visitation"),size=1.15) +
-  geom_line(aes(nat.x, UnitCode.y, linetype = "Predicted Visitation No Smoke"),size=1.15) +
-  scale_fill_manual(values=cols)+theme_classic()+xlab("Time")+ylab("Recreational Visits")
-
-ggplot(predDat.interp, aes(nat.x, nat.y)) +  facet_wrap(~UnitCode,scales="free") +
-  geom_ribbon(aes(ymin = nat.y, ymax = pmin(UnitCode.y, nat.y), fill = "Observed greater"),alpha=0.5) +
-  geom_ribbon(aes(ymin = UnitCode.y, ymax = pmin(UnitCode.y, nat.y), fill = "Predicted greater"),alpha=0.5) +
-  geom_line(color="#b2df8a",size=1.15) +
-  geom_line(aes(nat.x, UnitCode.y), color="#1f78b4",size=1.15) +
-  scale_fill_manual(values=cols)+theme_classic()+xlab("Time")+ylab("Recreational Visits")
-
-
-#individual
[…]

### 2020-04-27 SmokeModelAutoregressive.R

@@ -1,141 +1,0 @@
-library(tidyverse)
-library(rstan)
-library(shinystan)
-dat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")
-#stdize<-function(x) {return((x-mean(x)/(2*sd(x))))}
-
-dat<-dat%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
-dat<-dat%>%filter(SeasType =="High")
-
-dat<-na.omit(dat)
-
-dat$stdARV<-as.vector(scale(dat$ARVis))
-  
-
-data_list2 <- list(
-  N = nrow(dat),
-  previs = dat$stdARV,
-  Nprk = length(unique(dat$UnitCode)),
-  count = dat$RecreationVisits,
-  smoke = dat$stdsmoke,
-  pcode = as.numeric(as.factor(dat$UnitCode )))
-
-A2<- stan( file="~/SmokeProject/StanCode/SmokeNegBinomAutoregressive.stan" , 
-           data=data_list2,chains=3 ,warmup=1000,iter = 10000, control = list(max_treedepth = 25))
-
-print( A2 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
-
-save(A2, file="stdsmokeNbinomAutoregressive.rda")
-
-
-#Plot outputs
-load("~/SmokeProject/ModelObjects/stdsmokeNbinomAutoregressive.rda")
-x<-as.data.frame(summary(A2))
-zz<-row.names(x[2:61,])
-stan_plot(A2,pars = c(zz) ,fill_color = "purple", )
-
-
-########Compare predictions to overall smokiness##
-x$parameter<-row.names(x)
-#get just means for parks who's par estimates don't overlap zero
-x<-x[c(2:61),c(1,4,8,32)]%>%filter((summary.2.5. >0 & summary.97.5.>0) |
-                                     (summary.2.5. <0 & summary.97.5.<0) )
-
-
-Parks <- as.numeric(str_remove(substr(x$parameter,8,9),"]"))
-
-
-dat$PF<-as.numeric(as.factor(dat$UnitCode))
-
-
-
-ext_fit<-rstan::extract(A2)
-
-intercept_post<-ext_fit$intercept
-slope1_post<-ext_fit$slope1
[…]

### 2020-04-27 apply parameter estimates to vis data.R

@@ -1,122 +1,0 @@
-library(tidyverse)
-library(rstan)
-params<-read_csv("~/SmokeProject/Data/ParameterOutputs.csv")[,-c(1,3,5,6)]
-fulldat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")[,-1]
-fulldat<-fulldat%>%select(UnitCode,Year,Month,ParkName,Smoke,stdsmoke,RecreationVisits,SeasType)%>%
-  filter(SeasType =="High")%>%select(UnitCode,Year,Month,ParkName,Smoke,RecreationVisits,stdsmoke)
-names(params)[1]<-"UnitCode"
-
-fulldat<-left_join(fulldat,params,by="UnitCode")
-
-load("~/SmokeProject/ModelObjects/stdsmokeNbinom.rda")
-print( l2 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
-stan_plot(l2,pars = c(zz) ,fill_color = "purple", )
-
-x<-as.data.frame(summary(l2))
-x$parameter<-row.names(x)
-#get just means for parks who's par estimates don't overlap zero
-x<-x[c(2:61),c(1,4,8,32)]%>%filter((summary.2.5. >0 & summary.97.5.>0) |
-                    (summary.2.5. <0 & summary.97.5.<0) )
-
-
-Parks <- as.numeric(str_remove(substr(x$parameter,8,9),"]"))
-
-fulldat$PF<-as.numeric(as.factor(fulldat$UnitCode))
-sigdat<-fulldat%>% filter(PF %in% Parks)
-
-
-ext_fit<-rstan::extract(l2)
-
-intercept_post<-ext_fit$intercept
-slope1_post<-ext_fit$slope1
-phi_post<-ext_fit$phi
-sigma_pr_post<-ext_fit$sigma_pr
-ran_intercept_post<-ext_fit$ran_intercept
-lp__post<-ext_fit$lp__
-
-sp<-as.data.frame(slope1_post)
-ri<-as.data.frame(ran_intercept_post)
-
-
-
-##just one estimate
-gen_quant_r <- function(x,y) { #x here is the input smoke values #y is park
-  lin_comb <- mean(intercept_post) + 
-    x*mean(sp[,paste0("V",y)])+
-    mean(ri[,paste0("V",y)])
-  mu <- exp(lin_comb)
-  #phi<-sample(phi_post,size=length(x))
-  phi<-mean(phi_post)
[…]

### 2020-04-27 Rcode/PlotsForPaper.R
moved/organized files

@@ -1,0 +1,202 @@
+library(tidyverse)
+allData<-read_csv("Data/MergedDataComplete.csv")[,-1]
+predDat<-read_csv("~/SmokeProject/Data/VisitationEstimatesMinSmokeSignificantParks.csv")[,-1]
+predDat$date<-zoo::as.yearmon(predDat$date)
+allData$date<-zoo::as.yearmon(paste0(allData$Month,allData$Year),"%m%Y")
+allData<-allData%>% filter(SeasType =="High")
+ #all sig parks
+predDat.interp <- predDat %>% 
+   split(.$UnitCode) %>% 
+   map_df(~data.frame(UnitCode = approx(.x$date, .x$No_Smoke_Mean_est, n = 80), 
+                      nat = approx(.x$date, .x$RecreationVisits, n = 80), 
+                      UnitCode = .x$UnitCode[1]))
+
+
+ggplot(predDat.interp, aes(nat.x, nat.y)) +  facet_wrap(~UnitCode,scales="free") +
+  geom_ribbon(aes(ymin = nat.y, ymax = pmin(UnitCode.y, nat.y), fill = "Observed greater"),alpha=0.9) +
+  geom_ribbon(aes(ymin = UnitCode.y, ymax = pmin(UnitCode.y, nat.y), fill = "Predicted greater"),alpha=0.9) +
+  geom_line(aes(linetype = "Observed Visitation"),size=1.15) +
+  geom_line(aes(nat.x, UnitCode.y, linetype = "Predicted Visitation No Smoke"),size=1.15) +
+  scale_fill_manual(values=cols)+theme_classic()+xlab("Time")+ylab("Recreational Visits")
+
+ggplot(predDat.interp, aes(nat.x, nat.y)) +  facet_wrap(~UnitCode,scales="free") +
+  geom_ribbon(aes(ymin = nat.y, ymax = pmin(UnitCode.y, nat.y), fill = "Observed greater"),alpha=0.5) +
+  geom_ribbon(aes(ymin = UnitCode.y, ymax = pmin(UnitCode.y, nat.y), fill = "Predicted greater"),alpha=0.5) +
+  geom_line(color="#b2df8a",size=1.15) +
+  geom_line(aes(nat.x, UnitCode.y), color="#1f78b4",size=1.15) +
+  scale_fill_manual(values=cols)+theme_classic()+xlab("Time")+ylab("Recreational Visits")
+
+
+#individual
[…]

### 2020-04-27 Rcode/SmokeModelAutoregressive.R

@@ -1,0 +1,141 @@
+library(tidyverse)
+library(rstan)
+library(shinystan)
+dat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")
+#stdize<-function(x) {return((x-mean(x)/(2*sd(x))))}
+
+dat<-dat%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
+dat<-dat%>%filter(SeasType =="High")
+
+dat<-na.omit(dat)
+
+dat$stdARV<-as.vector(scale(dat$ARVis))
+  
+
+data_list2 <- list(
+  N = nrow(dat),
+  previs = dat$stdARV,
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$RecreationVisits,
+  smoke = dat$stdsmoke,
+  pcode = as.numeric(as.factor(dat$UnitCode )))
+
+A2<- stan( file="~/SmokeProject/StanCode/SmokeNegBinomAutoregressive.stan" , 
+           data=data_list2,chains=3 ,warmup=1000,iter = 10000, control = list(max_treedepth = 25))
+
+print( A2 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
+
+save(A2, file="stdsmokeNbinomAutoregressive.rda")
+
+
+#Plot outputs
+load("~/SmokeProject/ModelObjects/stdsmokeNbinomAutoregressive.rda")
+x<-as.data.frame(summary(A2))
+zz<-row.names(x[2:61,])
+stan_plot(A2,pars = c(zz) ,fill_color = "purple", )
+
+
+########Compare predictions to overall smokiness##
+x$parameter<-row.names(x)
+#get just means for parks who's par estimates don't overlap zero
+x<-x[c(2:61),c(1,4,8,32)]%>%filter((summary.2.5. >0 & summary.97.5.>0) |
+                                     (summary.2.5. <0 & summary.97.5.<0) )
+
+
+Parks <- as.numeric(str_remove(substr(x$parameter,8,9),"]"))
+
+
+dat$PF<-as.numeric(as.factor(dat$UnitCode))
+
+
+
+ext_fit<-rstan::extract(A2)
+
+intercept_post<-ext_fit$intercept
+slope1_post<-ext_fit$slope1
[…]

### 2020-04-27 Rcode/apply parameter estimates to vis data.R

@@ -1,0 +1,122 @@
+library(tidyverse)
+library(rstan)
+params<-read_csv("~/SmokeProject/Data/ParameterOutputs.csv")[,-c(1,3,5,6)]
+fulldat<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")[,-1]
+fulldat<-fulldat%>%select(UnitCode,Year,Month,ParkName,Smoke,stdsmoke,RecreationVisits,SeasType)%>%
+  filter(SeasType =="High")%>%select(UnitCode,Year,Month,ParkName,Smoke,RecreationVisits,stdsmoke)
+names(params)[1]<-"UnitCode"
+
+fulldat<-left_join(fulldat,params,by="UnitCode")
+
+load("~/SmokeProject/ModelObjects/stdsmokeNbinom.rda")
+print( l2 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
+stan_plot(l2,pars = c(zz) ,fill_color = "purple", )
+
+x<-as.data.frame(summary(l2))
+x$parameter<-row.names(x)
+#get just means for parks who's par estimates don't overlap zero
+x<-x[c(2:61),c(1,4,8,32)]%>%filter((summary.2.5. >0 & summary.97.5.>0) |
+                    (summary.2.5. <0 & summary.97.5.<0) )
+
+
+Parks <- as.numeric(str_remove(substr(x$parameter,8,9),"]"))
+
+fulldat$PF<-as.numeric(as.factor(fulldat$UnitCode))
+sigdat<-fulldat%>% filter(PF %in% Parks)
+
+
+ext_fit<-rstan::extract(l2)
+
+intercept_post<-ext_fit$intercept
+slope1_post<-ext_fit$slope1
+phi_post<-ext_fit$phi
+sigma_pr_post<-ext_fit$sigma_pr
+ran_intercept_post<-ext_fit$ran_intercept
+lp__post<-ext_fit$lp__
+
+sp<-as.data.frame(slope1_post)
+ri<-as.data.frame(ran_intercept_post)
+
+
+
+##just one estimate
+gen_quant_r <- function(x,y) { #x here is the input smoke values #y is park
+  lin_comb <- mean(intercept_post) + 
+    x*mean(sp[,paste0("V",y)])+
+    mean(ri[,paste0("V",y)])
+  mu <- exp(lin_comb)
+  #phi<-sample(phi_post,size=length(x))
+  phi<-mean(phi_post)
[…]

### 2020-04-27 Rcode/EntryFeeBackPredicting.R
small changes

@@ -40,6 +40,16 @@ names(NamesData)[2]<-"Park" #match the column names so we can merge
 NamesData<-data.frame(UnitCode=unique(NamesData$UnitCode),Park=unique(NamesData$Park))#get just each park once
 FEdat<-merge(FEdat, NamesData, by = "Park") #merge
 
+
+#plotting examples
+FEdat%>%
+  ggplot(., aes(x=Year,y=SingleVisitorFee,by=UnitCode))+  #note that when importing data from a pipe %>% into ggplot
+  geom_smooth(aes(color=Source),method="lm",se=F)         # you need a . as a data place holder
+
+FEdat%>%filter(Park=="Zion NP")%>%
+  ggplot(., aes(x=Year,y=SingleVisitorFee))+
+  geom_line(aes(color=Source))
+
 #now add the predicted visitation data
 VEstDat<-read.csv("~/SmokeProject/Data/VisitationEstimatesMinSmokeSignificantParks.csv")[,-1]
 
@@ -62,12 +72,7 @@ summaryFinancialLoss<-dat %>% group_by(UnitCode)%>%summarise(totrec=sum(incomeOb
   mutate(IncomeDif=totrec-totpred,IncomePrecDiff=round((totrec/totpred)*100,digits=2))
 
 
-#some example plots
-library(ggplot2) 
-FEdat%>%filter(Park=="Zion NP")%>%
-  ggplot(., aes(x=Year,y=SingleVisitorFee))+
-  geom_line(aes(color=Source))
  
-FEdat%>%filter(Source=="Observed")%>%
-  ggplot(., aes(x=Year,y=SingleVisitorFee,by=Park))+
-  geom_smooth(method="lm",se=F)
+
+ 
+

### 2020-05-08 Markdown Documents/Laura_Stoddard_VIP_Timeline.Rmd
cleaning up

@@
 date: "3/30/2020"
-output: html_document
+output:
+  word_document: default
+  html_document: default
 ---
@@
 We will do this by specifying the data and the x and y axes.
-Remember that the x and y axes need to be wrapped in an aestetic function aes().
+Remember that the x and y axes need to be wrapped in an aesthetic function aes().
 This just tells ggplot to look for these things in the dataset that we specified.
@@
 ------
+
+Now that you have some experience with the basics of plotting, let's work on cleaning up our figures so they look good and could go into a publication.
+
+When we're dealing with grouped data like this park data, we often want to facet the data so it's in small multiples, or repeated figures showing the same data for different groups.
+
+Let's first subset the parks data so that there's only 6 parks.
+
+```{r}
+d<-d%>%filter(Park %in% c("Acadia NP", "Yosemite NP", "Glacier NP","Zion NP", "Arches NP","Joshua Tree NP"))
+```
+
+Now we might want to look at the entry fees for each of these parks in their own figure. It would be a pain to code those figures for each park. Luckily we can use faceting.
+
+```{r}
+ggplot(d,aes(x=as.character(Year),y=SingleVisitorEntryFee))+
+  geom_point()+
+  facet_wrap(~Park) #this is the part that does the faceting
+```
+
+An easy thing we can do to make our plots look better is just to apply the build in ggplot themes to them. Here's some examples
+
+```{r}
+ggplot(d,aes(x=as.character(Year),y=SingleVisitorEntryFee))+
+  geom_point()+
+  facet_wrap(~Park)+theme_classic()
+```
+
+```{r}
+ggplot(d,aes(x=as.character(Year),y=SingleVisitorEntryFee))+
+  geom_point()+
+  facet_wrap(~Park)+theme_dark()
+```
+
+```{r}
+ggplot(d,aes(x=as.character(Year),y=SingleVisitorEntryFee))+
+  geom_point()+
+  facet_wrap(~Park)+theme_bw()
+```
+
+These themes are great, but usually, we want to customize them a bit more.
+
+We can do that by creating our own themes and stacking them on top.
+
+There's a TON of flexibility here.
+No one has all these options memorized.
+A good rule is that if you have an idea of what you want to do in ggplot it's almost definitely possible and someone has probably thought of it before you.
+Googling things is usually a good place to start.
+
+```{r}
+mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+                  plot.title = element_text( size=18, color="black",face="bold"),
+                  axis.title.x = element_text( size=18),
+                  axis.title.y = element_text( size=18),
+                  axis.text=(element_text(color="black", size=8)),
+                axis.text.x = element_text(angle = 90, hjust = 1),
+                  legend.title = element_text(colour="black", size=18),
+                  legend.text = element_text( size = 14))
+
+```
+
+Now that I have that saved, I can just add it to the end of my plot to make the changes I want
+
+```{r}
+ggplot(d,aes(x=as.character(Year),y=SingleVisitorEntryFee))+
+  geom_point()+
+  facet_wrap(~Park)+theme_classic()+
+  mytheme
+```
+
+We've really just scratched the surface here, but hopefully this is enough to get you started. Don't hesitate to reach out when you have questions!

### 2020-05-08 DiferencefromTrend.R

@@ -1,0 +1,76 @@
+library(tidyverse)
+library(lme4)
+library(rstan)
+Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
+Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
+#add smoke season
+#x<-Data%>%group_by(UnitCode,Season)%>%
+ # summarise(ms=mean(Smoke))%>%ungroup()%>%
+  #group_by(UnitCode)%>%filter(ms==max(ms))%>%mutate(cc=paste0(UnitCode,Season))
+#Data<-Data%>%filter( SmokeSeas =="High")
+
+
+#Data$SmokeSeas<-ifelse(Data$CatColS %in% x$cc,
+ #                      "High","NotHigh")
+
+#filter by high season
+Data<-Data%>%filter( SeasType =="High")
+
+#filter by smoke affected regions
+Data<-Data%>%
+  filter(Region %in% c("Intermountain","Pacific West"))
+
+
+#get vis and smoke trends
+fits <- lmList(stdsmoke ~ date | UnitCode, data=Data) 
+Data$trendsmoke<-predict(fits,date=allData$date)
+
+fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
+Data$trendvis<-predict(fits,date=Data$date)
+
+
+#plot it to make sure it worked
+Data%>%filter(UnitCode=="YELL")%>%
+ggplot(.,aes(x=date,y=stdsmoke))+geom_point()+geom_smooth(se=F,method = "lm")+
+  geom_point(aes(x=date,y=trendsmoke),color="green")
+
+
+Data%>%filter(UnitCode=="YOSE")%>%
+  ggplot(.,aes(x=date,y=RecreationVisits))+geom_point()+geom_smooth(se=F,method = "lm")+
+  geom_point(aes(x=date,y=trendvis),color="green")+ggtitle("Yosemite NP")
+
+
+Data$SmokeDiff<-Data$stdsmoke-Data$trendsmoke
+Data$VisDiff<-as.integer(Data$RecreationVisits-round(Data$trendvis,digits=0))
+
+
+#Model
+dat<-Data
+
+data_list3 <- list(
+  N = nrow(dat),
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$VisDiff,
+  smoke = dat$SmokeDiff,
[…]

### 2020-05-08 Rcode/EntryFeeBackPredicting.R

@@ -18,11 +18,23 @@ head(d) #check it out
 fillData<-FEdat[,1:2]
 
 
-FEdat$PredictedFee<-apply(posterior_predict(feefit,fillData),2,mean) #assuming this is for a new column
+FEdat$PredictedFee<-round(apply(posterior_predict(feefit,fillData),2,mean),digits = 0) #assuming this is for a new column
 
 xx<-na.omit(FEdat) #remove na's to compare accuracy of predictions and observed data
 cor(xx$PredictedFee,xx$SingleVisitorEntryFee) #correlation of predicted to observed
 plot(xx$PredictedFee,xx$SingleVisitorEntryFee) #plot it
+
+xx%>%
+  ggplot(.,aes(x=SingleVisitorEntryFee,y=round(PredictedFee),digits=0))+
+  geom_point()+
+  geom_smooth(method="lm", se=FALSE, colour="red")+
+  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(round(xx$PredictedFee,digits=0),xx$SingleVisitorEntryFee),digits=2)))+
+  ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+
+  theme(plot.title = element_text(hjust = 0.5))+
+  scale_y_continuous(name="Predicted Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))+
+  scale_x_continuous(name="Observed Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))
+
+
 
 #fill the na's leave it if it's not an na
 FEdat$SingleVisitorFee<-ifelse(is.na(FEdat$SingleVisitorEntryFee),FEdat$PredictedFee,FEdat$SingleVisitorEntryFee)
@@ -35,7 +47,7 @@ FEdat<-FEdat[,-c(3,4)]
 
 
 #the FEdat data doesn't have a unit code column. Let's add that
-NamesData<-read_csv("Data/MergedDataComplete.csv")[,-c(1,3,4,6:13)]
+NamesData<-read_csv("Data/MergedDataComplete.csv")[,-c(1,3,4,6:15)]
 names(NamesData)[2]<-"Park" #match the column names so we can merge
 NamesData<-data.frame(UnitCode=unique(NamesData$UnitCode),Park=unique(NamesData$Park))#get just each park once
 FEdat<-merge(FEdat, NamesData, by = "Park") #merge

### 2020-05-08 Rcode/MergeSmokeVisData.R

@@ -35,8 +35,10 @@ mdat<-mdat[,-c(4,7)]
 
 ##standardization function doing some weird rounding thing
 ## Doing it in 2 steps doesn't yeild the same problem
-zz<-mdat$Smoke-mean(mdat$Smoke)
-mdat$stdsmoke<-zz/(2*sd(mdat$Smoke))
+#zz<-mdat$Smoke-mean(mdat$Smoke)
+#mdat$stdsmoke<-zz/(2*sd(mdat$Smoke))
+mdat$stdsmoke<-as.vector(scale(mdat$Smoke))
+
 
 #make month categorical
 mdat$Month<-as.character(mdat$Month)
@@ -51,15 +53,15 @@ mdat$CatColS<-paste0(mdat$UnitCode,mdat$Season)
 mdat$CatColM<-paste0(mdat$UnitCode,mdat$Month)
 
 #create a variable for stdized smoke by park
-sdd1<-function(x) {return(x-mean(x))}
-sdd2<-function(x) {return(x/(2*sd(x)))}
+#sdd1<-function(x) {return(x-mean(x))}
+#sdd2<-function(x) {return(x/(2*sd(x)))}
 
-mdat<-mdat%>%group_by(UnitCode)%>%
-  mutate(deletelater=sdd1(Smoke))%>%
-  mutate(stdsmokepark=sdd2(deletelater))
+#mdat<-mdat%>%group_by(UnitCode)%>%
+  #mutate(deletelater=sdd1(Smoke))%>%
+  #mutate(stdsmokepark=sdd2(deletelater))
 
 #delete int vector
-mdat<-mdat[,-11]
+#mdat<-mdat[,-11]
 
 #add vis season
 x<-mdat%>%group_by(UnitCode,Season)%>%
@@ -72,5 +74,11 @@ x2<-mdat%>%group_by(UnitCode,Season)%>%
 
 mdat$SeasType<-ifelse(mdat$CatColS %in% x$cc,
                      "High",ifelse(mdat$CatColS %in% x2$cc,"Low","Shoulder"))
+
+#regions_states
+rd<-read.csv("~/SmokeProject/Data/States_Regions.csv")[-c(44,57),-1]
+
+mdat<-merge(mdat,rd,by="UnitCode")
+
 #create the csv
 write.csv(mdat, file="MergedDataComplete.csv")

### 2020-05-08 VariousARModels.R

@@ -1,0 +1,131 @@
+library(tidyverse)
+library(rstan)
+Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
+
+
+
+#add smoke season
+x<-Data%>%group_by(UnitCode,Season)%>%
+  summarise(ms=mean(Smoke))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(ms==max(ms))%>%mutate(cc=paste0(UnitCode,Season))
+
+
+
+Data$SmokeSeas<-ifelse(Data$CatColS %in% x$cc,
+                      "High","NotHigh")
+
+
+
+#Data<-Data%>%filter( SmokeSeas =="High")
+Data<-Data%>%filter( SeasType =="High")
+
+Data<-Data%>%
+  filter(Region %in% c("Intermountain","Pacific West"))
+##
+
+dat<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
+
+dat<-na.omit(dat)
+
+dat$stdARV<-as.vector(scale(dat$ARVis))
+
+
+data_list2 <- list(
+  N = nrow(dat),
+  previs = dat$stdARV,
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$RecreationVisits,
+  smoke = dat$stdsmoke,
+  pcode = as.numeric(as.factor(dat$UnitCode )))
+
+options(mc.cores=parallel::detectCores()-1)
+
+A3<- stan( file="~/SmokeProject/StanCode/SmokeNegBinomAutoregressive.stan" , 
+           data=data_list2,chains=3 ,warmup=1000,iter = 8000, control = list(max_treedepth = 25))
+
+
+
+print( A3 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
+
+save(A3, file="stdsmokeNbinomAutoregressiveSubset.rda")
+
+x<-as.data.frame(summary(A3))
+zz<-row.names(x[2:34,])
+stan_plot(A3,pars = c(zz) ,fill_color = "purple", )
+
+## no AR term
+dat<-Data
+
+data_list3 <- list(
+  N = nrow(dat),
+  Nprk = length(unique(dat$UnitCode)),
+  count = dat$RecreationVisits,
+  smoke = dat$stdsmoke,
+  pcode = as.numeric(as.factor(dat$UnitCode )))
+
[…]

### 2020-05-12 DiferencefromTrend.R
Residuals...taking a long time to run

@@ -3,6 +3,8 @@ library(lme4)
 library(rstan)
 Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
 Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
+dat<-dat%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
+dat<-na.omit(dat)
 #add smoke season
 #x<-Data%>%group_by(UnitCode,Season)%>%
  # summarise(ms=mean(Smoke))%>%ungroup()%>%
@@ -22,7 +24,7 @@ Data<-Data%>%
 
 
 #get vis and smoke trends
-fits <- lmList(stdsmoke ~ date | UnitCode, data=Data) 
+fits <- lmList(Smoke ~ date | UnitCode, data=Data) 
 Data$trendsmoke<-predict(fits,date=allData$date)
 
 fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
@@ -31,7 +33,7 @@ Data$trendvis<-predict(fits,date=Data$date)
 
 #plot it to make sure it worked
 Data%>%filter(UnitCode=="YELL")%>%
-ggplot(.,aes(x=date,y=stdsmoke))+geom_point()+geom_smooth(se=F,method = "lm")+
+ggplot(.,aes(x=date,y=Smoke))+geom_point()+geom_smooth(se=F,method = "lm")+
   geom_point(aes(x=date,y=trendsmoke),color="green")
 
 
@@ -40,9 +42,18 @@ Data%>%filter(UnitCode=="YOSE")%>%
   geom_point(aes(x=date,y=trendvis),color="green")+ggtitle("Yosemite NP")
 
 
-Data$SmokeDiff<-Data$stdsmoke-Data$trendsmoke
-Data$VisDiff<-as.integer(Data$RecreationVisits-round(Data$trendvis,digits=0))
+Data$SmokeDiff<-scale(Data$Smoke-Data$trendsmoke)
+Data$VisDiff<-Data$RecreationVisits-Data$trendvis
 
+Data%>%
+  ggplot(.,aes(x=SmokeDiff,y=VisDiff))+
+  geom_point(aes(color=UnitCode))+
+  geom_smooth(aes(color=UnitCode),se=F,method = "lm")+theme_classic()
+
+Data%>%
+  ggplot(.,aes(x=SmokeDiff,y=VisDiff))+
+  geom_point()+
[…]

### 2020-05-12 DiferencefromTrend.R
I think it's working now. Vis residuals predicted by smoke.

@@ -30,7 +30,9 @@ Data$trendsmoke<-predict(fits,date=allData$date)
 fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
 Data$trendvis<-predict(fits,date=Data$date)
 
-
+Data$SmokeDiff<-scale(Data$Smoke-Data$trendsmoke)
+Data$VisDiff<-Data$RecreationVisits-Data$trendvis
+dat<-Data
 #plot it to make sure it worked
 Data%>%filter(UnitCode=="YELL")%>%
 ggplot(.,aes(x=date,y=Smoke))+geom_point()+geom_smooth(se=F,method = "lm")+
@@ -55,64 +57,40 @@ Data%>%
   geom_point()+
   geom_smooth(se=F,method = "lm")+theme_classic()+facet_wrap(~Region)
 
-#Model
-dat<-Data
 
-data_list3 <- list(
+
+############################
+
+
+
+
+
+
+data_list4 <- list(
   N = nrow(dat),
   Nprk = length(unique(dat$UnitCode)),
-  vis = dat$VisDiff,
-  smoke = dat$SmokeDiff,
+  vis = as.vector(dat$VisDiff),
+  smoke = dat$stdsmoke,
   pcode = as.numeric(as.factor(dat$UnitCode )))
 
 options(mc.cores=parallel::detectCores()-1)
-rstan_options(auto_write=T)
-
-M1<- stan( file="~/SmokeProject/StanCode/Residuals.stan" , 
-           data=data_list3,chains=3 ,warmup=50,iter = 200, control = list(max_treedepth = 25))
-
-############################
-Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
-Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
-Data<-Data%>%filter( SeasType =="High")
-fits <- lmList(Smoke ~ date | UnitCode, data=Data) 
-Data$trendsmoke<-predict(fits,date=allData$date)
-
-fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
-Data$trendvis<-predict(fits,date=Data$date)
-Data$SmokeDiff<-scale(Data$Smoke-Data$trendsmoke)
-Data$VisDiff<-Data$RecreationVisits-Data$trendvis
-dat<-Data
-
-data_list3 <- list(
-  N = nrow(dat),
[…]

### 2020-05-13 DiferencefromTrend.R
random changes

@@ -3,18 +3,19 @@ library(lme4)
 library(rstan)
 Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
 Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
-dat<-dat%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
-dat<-na.omit(dat)
+Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
+Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
+Data<-na.omit(Data)
 #add smoke season
 #x<-Data%>%group_by(UnitCode,Season)%>%
  # summarise(ms=mean(Smoke))%>%ungroup()%>%
   #group_by(UnitCode)%>%filter(ms==max(ms))%>%mutate(cc=paste0(UnitCode,Season))
-#Data<-Data%>%filter( SmokeSeas =="High")
 
 
 #Data$SmokeSeas<-ifelse(Data$CatColS %in% x$cc,
  #                      "High","NotHigh")
 
+#Data<-Data%>%filter( SmokeSeas =="High")
 #filter by high season
 Data<-Data%>%filter( SeasType =="High")
 
@@ -34,18 +35,17 @@ Data$SmokeDiff<-scale(Data$Smoke-Data$trendsmoke)
 Data$VisDiff<-Data$RecreationVisits-Data$trendvis
 dat<-Data
 #plot it to make sure it worked
-Data%>%filter(UnitCode=="YELL")%>%
+Data%>%filter(UnitCode=="BIBE")%>%
 ggplot(.,aes(x=date,y=Smoke))+geom_point()+geom_smooth(se=F,method = "lm")+
   geom_point(aes(x=date,y=trendsmoke),color="green")
 
 
-Data%>%filter(UnitCode=="YOSE")%>%
+Data%>%filter(UnitCode=="BIBE")%>%
   ggplot(.,aes(x=date,y=RecreationVisits))+geom_point()+geom_smooth(se=F,method = "lm")+
   geom_point(aes(x=date,y=trendvis),color="green")+ggtitle("Yosemite NP")
 
 
-Data$SmokeDiff<-scale(Data$Smoke-Data$trendsmoke)
-Data$VisDiff<-Data$RecreationVisits-Data$trendvis
+
 
 Data%>%
   ggplot(.,aes(x=SmokeDiff,y=VisDiff))+
[…]

### 2020-05-14 HESBoelWorkingScript.R
rstanarm script & model of residuals

@@ -1,0 +1,45 @@
+library(tidyverse)
+library(lme4)
+library(rstan)
+Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
+Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
+Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
+Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
+Data<-na.omit(Data)
+#add smoke season
+#x<-Data%>%group_by(UnitCode,Season)%>%
+# summarise(ms=mean(Smoke))%>%ungroup()%>%
+#group_by(UnitCode)%>%filter(ms==max(ms))%>%mutate(cc=paste0(UnitCode,Season))
+
+
+#Data$SmokeSeas<-ifelse(Data$CatColS %in% x$cc,
+#                      "High","NotHigh")
+
+#Data<-Data%>%filter( SmokeSeas =="High")
+#filter by high season
+Data<-Data%>%filter( SeasType =="High")
+
+#filter by smoke affected regions
+Data<-Data%>%
+  filter(Region %in% c("Intermountain","Pacific West"))
+
+
+#get vis and smoke trends
+fits <- lmList(Smoke ~ date | UnitCode, data=Data) 
+Data$trendsmoke<-predict(fits,date=allData$date)
+
+fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
+Data$trendvis<-predict(fits,date=Data$date)
+
+Data$SmokeDiff<-scale(Data$Smoke-Data$trendsmoke)
+Data$VisDiff<-Data$RecreationVisits-Data$trendvis
+dat<-Data
+
+library(rstanarm)
+options(mc.cores=parallel::detectCores())
+fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+save(fit,file="residuals.rda")
+
+
+bayesplot::color_scheme_set("viridisC")
+(trace <- plot(fit, "trace", pars = c("b[stdsmoke UnitCode:ZION]","b[stdsmoke UnitCode:YELL]","b[stdsmoke UnitCode:SEQU]","(Intercept)")))

### 2020-05-14 HESBoelWorkingScript.R
added vis predictions and plotz

@@ -43,3 +43,31 @@ save(fit,file="residuals.rda")
 
 bayesplot::color_scheme_set("viridisC")
 (trace <- plot(fit, "trace", pars = c("b[stdsmoke UnitCode:ZION]","b[stdsmoke UnitCode:YELL]","b[stdsmoke UnitCode:SEQU]","(Intercept)")))
+
+
+#look at the predictions
+
+dat<-dat%>%group_by(UnitCode)%>%mutate(MinSmoke=min(stdsmoke))
+
+minsmokedata<-data.frame(UnitCode=dat$UnitCode,stdsmoke=dat$MinSmoke)
+
+
+dat$PredNoSmoke<-apply(posterior_predict(fit,minsmokedata),2,mean)
+
+dat$PredNoSmoke<-dat$trendvis+dat$PredNoSmoke
+
+dat%>%filter(UnitCode =="YOSE")%>%
+ggplot(.,aes(x=date))+
+  geom_line(aes(y=RecreationVisits),color="green")+
+  geom_line(aes(y=PredNoSmoke),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")
+
+
+
+  ggplot(dat,aes(x=date))+
+  geom_line(aes(y=RecreationVisits),color="green")+
+  geom_line(aes(y=PredNoSmoke),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")+facet_wrap(~UnitCode,scales="free")
+
+  p <- plot(fit, regex_pars = "stdsmoke UnitCode",
+            prob = 0.5, prob_outer = 0.9)
+
+

### 2020-05-14 HESBoelWorkingScript.R
hes bowl script

@@ -24,22 +24,18 @@ Data<-Data%>%
   filter(Region %in% c("Intermountain","Pacific West"))
 
 
-#get vis and smoke trends
-fits <- lmList(Smoke ~ date | UnitCode, data=Data) 
-Data$trendsmoke<-predict(fits,date=allData$date)
+#get vis trends
 
 fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
 Data$trendvis<-predict(fits,date=Data$date)
-
-Data$SmokeDiff<-scale(Data$Smoke-Data$trendsmoke)
 Data$VisDiff<-Data$RecreationVisits-Data$trendvis
 dat<-Data
 
 library(rstanarm)
 options(mc.cores=parallel::detectCores())
-fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
-save(fit,file="residuals.rda")
-
+#fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fit,file="residuals.rda")
+load("~/SmokeProject/ModelObjects/residuals.rda")
 
 bayesplot::color_scheme_set("viridisC")
 (trace <- plot(fit, "trace", pars = c("b[stdsmoke UnitCode:ZION]","b[stdsmoke UnitCode:YELL]","b[stdsmoke UnitCode:SEQU]","(Intercept)")))

### 2020-05-14 Manuscript.Rmd
mnuscript. rmd

---
title: Estimating the Financial Impact of Wildfire Smoke on US National parks
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: University of Michigan
  name: Alexander Killion
- affiliation: Boise State University
  name: Matthew A. Williamson
- affiliation: Boise State University
  name: Vicken?
- affiliation: Boise State University
  name: Laura? (undergrad who will continue working over summer)
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: /Users/Matthewclark989/Documents/LatexPractice/templatedocs/svm_ms_template.tex
  html_document:
    df_print: paged
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
keywords: pandoc, r markdown, knitr
bibliography: /Users/Matthewclark989/Documents/SmokeProject/References.bib
thanks: If we want to thank anyone? Maybe Moji?.
abstract: National Parks are important. National Parks are low on funding which affects their ability to do their job. A key way they get this funding is through visitation. Wildfire smoke affects visitation greatly and is out of ther control. Wildfires are expected to increase in the Pacific West and Intermountain regions. Over the last 38 years, we estimate that wildfire smoke has led to a visitation decrease of XX visitors in these regions. This visitation dampening has cost the National Park Service $XX based on our estimates. Wildfires and the smoke it produces is supposed to increase over the next 50 years. Assuming a linear trend in smoke, visitation overall, and park fees, we might see a financial loss to the park service of $XX. Our recomendation is that....NPS work with the forest service to prevent fires in areas close to parks?? Do they already do that??    
---
# Introduction 
Public lands are a cornerstone of American culture and identity. The crown jewls of the American public lands system are the National Parks. National Parks have recently drawn public attention for having a lack of funding which negatively affects their ability to protect America's most cherished natural resources and provide visitors with desired recreational experiences. While the National Park Service makes countless decisions regarding where to allocate resources and how to best invest its limited funding, other monetary sources and sinks are out of the control of the National Park Service.
This is how you add a reference (e.g., [@fisichelli_protected_2015; @fann_ch_2016])

# Methods
```{r include=FALSE}
library(tidyverse)
library(lme4)
library(rstanarm)
library(rstan)
load("~/SmokeProject/ModelObjects/feefit.rda")
load("~/SmokeProject/ModelObjects/residuals.rda")
```
## Data Collection

### Visitor Data

### Visitation Fees

We started with an incomplete dataset. Laura will write this part with information on collecting this data.
```{r include=FALSE}
FEdat<-read.csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
FEdat<-FEdat %>% dplyr::select(Park,Year,SingleVisitorEntryFee) #keep only these columns
```

```{r echo=FALSE}
head(FEdat[c(1009,256,10,800,655),])
```

We removed the NA's from this dataset built a linear model which estimates the fees based on the date. 
```{r, echo=F}
d<-na.omit(FEdat) #get rid of rows with any NA
```

$$
\begin{aligned}
y_{tp} &\sim \begin{cases}\operatorname{Pois}(\mu,\lambda)\\ \mu = B_{0}+ B_{1}*X_{1} +B_{2}*X_{2}
\end{cases}
\end{aligned}
$$

We operationalized this model in rstanarm
```{r eval=FALSE}
feefit<-stan_glm(SingleVisitorEntryFee~Year+Park,family="poisson",data=d,chains=3,iter=6000)
```

We then used this to predict what the fees ould have been for every year we didn't have
```{r}
fillData<-FEdat[,1:2]

FEdat$PredictedFee<-round(apply(posterior_predict(feefit,fillData),2,mean),digits = 0) #assuming this is for a new column

```

Our predicted fees correlate well with our observed fees that we have
```{r echo = FALSE}
xx<-na.omit(FEdat)
 mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                  plot.title = element_text( size=18, color="black",face="bold"),
                  axis.title.x = element_text( size=18),
                  axis.title.y = element_text( size=18),
                  axis.text=(element_text(color="black", size=14)),
                  legend.title = element_text(colour="black", size=18),
                  legend.text = element_text( size = 14))
xx%>%
  ggplot(.,aes(x=SingleVisitorEntryFee,y=round(PredictedFee),digits=0))+
  geom_jitter(size=2, alpha=0.3)+
  geom_smooth(method="lm", se=FALSE, colour="red")+
  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(round(xx$PredictedFee,digits=0),xx$SingleVisitorEntryFee),digits=2)))+
  ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(name="Predicted Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))+
  scale_x_continuous(name="Observed Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))
```
### Smoke Data

Alex to complete this section

## Analysis

### Smoke Effects Past
We subset the data to only include high season visitation for parks in the Intermountain and Pacific West regions. The regions projected to be most havily affected by climate change.

```{r include=FALSE}
library(tidyverse)
library(lme4)
library(rstan)
Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
Data<-na.omit(Data)

#filter by high season
Data<-Data%>%filter( SeasType =="High")

#filter by smoke affected regions
Data<-Data%>%
  filter(Region %in% c("Intermountain","Pacific West"))

```

To account for the linear trends of visitation, we first calculated the trends, then found the residuals of each datapoint from the trendline.
$$
\begin{aligned}
Resid_{y} &= \begin{cases}y_{obs}-y_{trend}\\ y_{trend} \sim \operatorname{Norm}(\mu_{park},\sigma)\\ \mu_{park} = B_{0}+ B_{1}*X_{1}
\end{cases}
\end{aligned}
$$

```{r warning=FALSE}
fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
Data$trendvis<-predict(fits,date=Data$date)
Data$VisDiff<-Data$RecreationVisits-Data$trendvis
dat<-Data
```

We then used a hierarchical modeling structure to understand these residuals as a result of the max observed smoke values in each park in each month.
$$
\begin{aligned}
Resid_{y} &\sim \begin{cases}\operatorname{Norm}(\mu,\sigma)\\ \mu = B_{0Park}+ B_{1Park}*X_{1Park} \\ \sigma\sim Norm(0,1) 
\end{cases}
\end{aligned}
$$

This was also operationalized in `rstanarm`
```{r,eval=FALSE}
fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)

```

### Smoke, Fee, & Visitation Projections

# Results

For all significant effects (90% C.I. doesn't overlap 0), we see a negative effect of smoke on the residual visitation to National Parks.

```{r,echo=FALSE}
 plot(fit, regex_pars = "stdsmoke UnitCode",
            prob = 0.5, prob_outer = 0.9)
```

We can then go back and apply these estimates effects to the observed visitation given a minimum level of smoke observed for each park to estimate what the vititation may have been given optimal air quality.

Here's an example from a few parks

```{r include=FALSE}

dat<-dat%>%group_by(UnitCode)%>%mutate(MinSmoke=min(stdsmoke))
```

```{r include=FALSE}
minsmokedata<-data.frame(UnitCode=dat$UnitCode,stdsmoke=dat$MinSmoke)
```

```{r include=FALSE}
dat$PredNoSmoke<-apply(posterior_predict(fit,minsmokedata),2,mean)
```

```{r include=FALSE}
dat$PredNoSmoke<-dat$trendvis+dat$PredNoSmoke
```

```{r echo=FALSE, warning=FALSE}
dat%>%filter(UnitCode == "ROMO")%>%
ggplot(.,aes(x=date))+
  geom_line(aes(y=RecreationVisits),color="green")+
  geom_line(aes(y=PredNoSmoke),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")

dat%>%filter(UnitCode == "YOSE")%>%
ggplot(.,aes(x=date))+
  geom_line(aes(y=RecreationVisits),color="green")+
  geom_line(aes(y=PredNoSmoke),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")

dat%>%filter(UnitCode == "ZION")%>%
ggplot(.,aes(x=date))+
  geom_line(aes(y=RecreationVisits),color="green")+
  geom_line(aes(y=PredNoSmoke),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")

```

# Discussion

Discussion here

### 2020-05-14 Rcode/EntryFeeBackPredicting.R
point to jitter

@@ -7,6 +7,7 @@ FEdat<-read.csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
 head(FEdat)
 
 FEdat<-FEdat %>% dplyr::select(Park,Year,SingleVisitorEntryFee) #keep only these columns
+head(FEdat[c(1009,256,10,800,655),])
 d<-na.omit(FEdat) #get rid of rows with any NA
 head(d) #check it out
 
@@ -26,7 +27,7 @@ plot(xx$PredictedFee,xx$SingleVisitorEntryFee) #plot it
 
 xx%>%
   ggplot(.,aes(x=SingleVisitorEntryFee,y=round(PredictedFee),digits=0))+
-  geom_point()+
+  geom_jitter(size=2, alpha=0.3)+
   geom_smooth(method="lm", se=FALSE, colour="red")+
   geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(round(xx$PredictedFee,digits=0),xx$SingleVisitorEntryFee),digits=2)))+
   ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+

### 2020-05-15 Manuscript.Rmd
manuscript and R script

@@
 output:
+  html_document:
+    df_print: paged
   word_document: default
@@
     template: /Users/Matthewclark989/Documents/LatexPractice/templatedocs/svm_ms_template.tex
-  html_document:
-    df_print: paged
 fontfamily: mathpazo
@@
 Wildfires are expected to increase in the Pacific West and Intermountain regions.
-Over the last 38 years, we estimate that wildfire smoke has led to a visitation decrease of XX visitors in these regions.
-This visitation dampening has cost the National Park Service $XX based on our estimates.
+Over the last 38 years, we estimate that wildfire smoke has led to a 12% overall visitation decrease to National Parks in these regions.
+When accounting for visitation fees, this loss may have cost the National Park Service as much as $580 million in total.
 Wildfires and the smoke it produces is supposed to increase over the next 50 years.
@@
 ---
-# Introduction 
+# 1. Introduction 
 Public lands are a cornerstone of American culture and identity.
-The crown jewls of the American public lands system are the National Parks.
+The crown jewels of the American public lands system are the National Parks.
 National Parks have recently drawn public attention for having a lack of funding which negatively affects their ability to protect America's most cherished natural resources and provide visitors with desired recreational experiences.
@@
 
-# Methods
-```{r include=FALSE}
+# 2. Methods
+```{r include=FALSE, cache=TRUE}
 library(tidyverse)
@@
 library(rstan)
+library(zoo)
+library(formattable)
 load("~/SmokeProject/ModelObjects/feefit.rda")
@@
 ```
-## Data Collection
+## 2.1. Data Collection
 
-### Visitor Data
+### 2.1.1. Visitor Data
 
-### Visitation Fees
+### 2.1.2. Visitation Fees
 
 We started with an incomplete dataset. Laura will write this part with information on collecting this data.
-```{r include=FALSE}
+```{r include=FALSE, cache=TRUE}
 FEdat<-read.csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, cache=TRUE}
 head(FEdat[c(1009,256,10,800,655),])
@@
 
-We removed the NA's from this dataset built a linear model which estimates the fees based on the date. 
-```{r, echo=F}
+We removed the NA's from this dataset built a linear model which estimates the fees based on the date and the park. 
+```{r, echo=FALSE, cache=TRUE}
 d<-na.omit(FEdat) #get rid of rows with any NA
@@
 ```{r eval=FALSE}
-feefit<-stan_glm(SingleVisitorEntryFee~Year+Park,family="poisson",data=d,chains=3,iter=6000)
+feefit<-stan_glm(SingleVisitorEntryFee~Year+Park,family="Poisson",data=d,chains=3,iter=6000)
 ```
 
-We then used this to predict what the fees ould have been for every year we didn't have
-```{r}
+We then used this to predict what the fees would have been for every year we didn't have
+```{r cache=TRUE}
 fillData<-FEdat[,1:2]
@@
 Our predicted fees correlate well with our observed fees that we have
-```{r echo = FALSE}
+```{r echo = FALSE, cache=TRUE}
 xx<-na.omit(FEdat)
@@
 ```
-### Smoke Data
+
+### 2.1.3. Smoke Data
 
@@
 
-## Analysis
+## 2.2. Analysis
 
-### Smoke Effects Past
-We subset the data to only include high season visitation for parks in the Intermountain and Pacific West regions. The regions projected to be most havily affected by climate change.
+### 2.2.1. Smoke Effects Past
+We subset the data to only include high season visitation for parks in the Inter-mountain and Pacific West regions. The regions projected to be most heavily affected by climate change.
 
-```{r include=FALSE}
+```{r include=FALSE, cache=TRUE}
 library(tidyverse)
@@
 
-```{r warning=FALSE}
+```{r warning=FALSE, cache=TRUE}
 fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
@@
 This was also operationalized in `rstanarm`
-```{r,eval=FALSE}
+```{r eval=FALSE, cache=TRUE}
 fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
@@
 
-### Smoke, Fee, & Visitation Projections
+### 2.2.2. Smoke, Fee, & Visitation Projections
 
-# Results
+# 3. Results
 
@@
 
-```{r,echo=FALSE}
+```{r echo=FALSE, cache=TRUE}
  plot(fit, regex_pars = "stdsmoke UnitCode",
@@
 
-We can then go back and apply these estimates effects to the observed visitation given a minimum level of smoke observed for each park to estimate what the vititation may have been given optimal air quality.
+We can then go back and apply these estimates effects to the observed visitation given a minimum level of smoke observed for each park to estimate what the visitation may have been given optimal air quality.
 
@@
 
-```{r include=FALSE}
+```{r include=FALSE, cache=TRUE}
 
-dat<-dat%>%group_by(UnitCode)%>%mutate(MinSmoke=min(stdsmoke))
+dat<-read.csv("~/SmokeProject/Data/FullDataWithMinSmokePredictions.csv")
+dat$date<-zoo::as.yearmon(dat$date)
 ```
 
-```{r include=FALSE}
-minsmokedata<-data.frame(UnitCode=dat$UnitCode,stdsmoke=dat$MinSmoke)
-```
-
-```{r include=FALSE}
-dat$PredNoSmoke<-apply(posterior_predict(fit,minsmokedata),2,mean)
-```
-
-```{r include=FALSE}
-dat$PredNoSmoke<-dat$trendvis+dat$PredNoSmoke
-```
-
-```{r echo=FALSE, warning=FALSE}
-dat%>%filter(UnitCode == "ROMO")%>%
+```{r echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE}
+library(scales)
+dat%>%filter(UnitCode =="ROMO")%>%
 ggplot(.,aes(x=date))+
-  geom_line(aes(y=RecreationVisits),color="green")+
-  geom_line(aes(y=PredNoSmoke),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")
+  geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
+  geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
+  geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
+  geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
+  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
+  scale_color_identity(name = "",
+                       breaks = c("black", "#1f78b4"),
+                       labels = c("Observed", "Median Estimate"),
+                       guide = "legend")+ggtitle("Rocky Mountain National Park")
 
@@
 ggplot(.,aes(x=date))+
-  geom_line(aes(y=RecreationVisits),color="green")+
-  geom_line(aes(y=PredNoSmoke),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")
+  geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
+  geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
+  geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
+  geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
+  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
+  scale_color_identity(name = "",
+                       breaks = c("black", "#1f78b4"),
+                       labels = c("Observed", "Median Estimate"),
+                       guide = "legend")+ggtitle("Yosemite National Park")
 
@@
 ggplot(.,aes(x=date))+
-  geom_line(aes(y=RecreationVisits),color="green")+
-  geom_line(aes(y=PredNoSmoke),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")
+  geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
+  geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
+  geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
+  geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
+  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
+  scale_color_identity(name = "",
+                       breaks = c("black", "#1f78b4"),
+                       labels = c("Observed", "Median Estimate"),
+                       guide = "legend")+ggtitle("Zion National Park *not significant")
 
@@
 
-# Discussion
+We can apply these visitation estimates to the fee data/predictions also. This gives us the following table.
+```{r echo=FALSE, cache=TRUE}
+fitint<-as.data.frame(posterior_interval(fit))
+names(fitint)<-c("CI5","CI95")
+fitint$Par<-row.names(fitint)
+fitsig<-fitint[grep("stdsmoke UnitCode",row.names(fitint)),]%>% 
+  filter(CI5<0 & CI95<0)
+
+fitsig$Par<-substr(fitsig$Par,21,24)
+
+#predicted vis fees 
+FEdat<-read.csv("~/SmokeProject/Data/ParkFees.csv")
+
+#get the visitation observed and predicted yearly totals (same scale as fee data)
+VEstDat<-dat%>%group_by(UnitCode,Year)%>%
+  summarise(RecVisits=sum(RecreationVisits),PredVisits = sum(PredNoSmoke))%>%
+  filter(UnitCode %in% fitsig$Par)
+
+#only include significant parks and years we predicted for
+FEdat<-FEdat%>%filter(UnitCode%in%fitsig$Par)
+
+#merge the fee and visitation data
+financialData<-merge(FEdat, VEstDat, by = c("UnitCode","Year"))
+
+#multiply the fees and visitation (predicted and observed)
+financialData<-financialData%>%mutate(incomeObserved= round(SingleVisitorFee*RecVisits,digits = 2),incomePred= round(SingleVisitorFee*PredVisits,digits=2))
+
+#get the dollar estimates for smoke loss for each park)
+summaryFinancialLoss<-financialData %>% group_by(UnitCode)%>%summarise(totrec=sum(incomeObserved),totpred=sum(incomePred))%>%
+  mutate(IncomeDif=totrec-totpred,IncomePrecDiff=round((1-(totrec/totpred))*100,digits=2))
+
+summaryFinancialLoss$ParkName<-unique(dat[dat$UnitCode %in% summaryFinancialLoss$UnitCode,]$ParkName)
+
+summaryFinancialLoss<-summaryFinancialLoss[,c(6,4,5)]
+summaryFinancialLoss$IncomeDif<-currency(summaryFinancialLoss$IncomeDif)
+names(summaryFinancialLoss)<-(c("National Park","Income Loss Estimate (1980-2018)", "Percent Income Loss (1980-2018)"))
+summaryFinancialLoss$`Percent Income Loss (1980-2018)`<-paste0(summaryFinancialLoss$`Percent Income Loss (1980-2018)`,'%')
+formattable::formattable(summaryFinancialLoss)
+```
+
+# 4. Discussion
 

### 2020-05-15 HESBoelWorkingScript.R

@@ -6,16 +6,7 @@ Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
 Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
 Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
 Data<-na.omit(Data)
-#add smoke season
-#x<-Data%>%group_by(UnitCode,Season)%>%
-# summarise(ms=mean(Smoke))%>%ungroup()%>%
-#group_by(UnitCode)%>%filter(ms==max(ms))%>%mutate(cc=paste0(UnitCode,Season))
 
-
-#Data$SmokeSeas<-ifelse(Data$CatColS %in% x$cc,
-#                      "High","NotHigh")
-
-#Data<-Data%>%filter( SmokeSeas =="High")
 #filter by high season
 Data<-Data%>%filter( SeasType =="High")
 
@@ -48,14 +39,41 @@ dat<-dat%>%group_by(UnitCode)%>%mutate(MinSmoke=min(stdsmoke))
 minsmokedata<-data.frame(UnitCode=dat$UnitCode,stdsmoke=dat$MinSmoke)
 
 
-dat$PredNoSmoke<-apply(posterior_predict(fit,minsmokedata),2,mean)
+dat$PredNoSmoke50CI<-apply(posterior_predict(fit,minsmokedata),2,median)
+dat$PredNoSmoke95CI<-apply(posterior_predict(fit,minsmokedata),2,function(x){quantile(x,0.95)})
+dat$PredNoSmoke5CI<-apply(posterior_predict(fit,minsmokedata),2,function(x){quantile(x,0.05)})
+dat$PredNoSmoke75CI<-apply(posterior_predict(fit,minsmokedata),2,function(x){quantile(x,0.75)})
+dat$PredNoSmoke25CI<-apply(posterior_predict(fit,minsmokedata),2,function(x){quantile(x,0.25)})
 
-dat$PredNoSmoke<-dat$trendvis+dat$PredNoSmoke
 
-dat%>%filter(UnitCode =="YOSE")%>%
+dat$PredNoSmoke50CI<-dat$trendvis+dat$PredNoSmoke50CI
+dat$PredNoSmoke95CI<-dat$trendvis+dat$PredNoSmoke95CI
[…]

### 2020-05-15 Rcode/EntryFeeBackPredicting.R
small changes

@@ -7,7 +7,7 @@ FEdat<-read.csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
 head(FEdat)
 
 FEdat<-FEdat %>% dplyr::select(Park,Year,SingleVisitorEntryFee) #keep only these columns
-head(FEdat[c(1009,256,10,800,655),])
+#head(FEdat[c(1009,256,10,800,655),])
 d<-na.omit(FEdat) #get rid of rows with any NA
 head(d) #check it out
 
@@ -18,8 +18,8 @@ head(d) #check it out
 
 fillData<-FEdat[,1:2]
 
-
-FEdat$PredictedFee<-round(apply(posterior_predict(feefit,fillData),2,mean),digits = 0) #assuming this is for a new column
+load("~/SmokeProject/ModelObjects/feefit.rda")
+FEdat$PredictedFee<-apply(posterior_predict(feefit,fillData),2,median) #assuming this is for a new column
 
 xx<-na.omit(FEdat) #remove na's to compare accuracy of predictions and observed data
 cor(xx$PredictedFee,xx$SingleVisitorEntryFee) #correlation of predicted to observed
@@ -48,12 +48,12 @@ FEdat<-FEdat[,-c(3,4)]
 
 
 #the FEdat data doesn't have a unit code column. Let's add that
-NamesData<-read_csv("Data/MergedDataComplete.csv")[,-c(1,3,4,6:15)]
-names(NamesData)[2]<-"Park" #match the column names so we can merge
+NamesData<-read_csv("Data/MergedDataComplete.csv")
+names(NamesData)[5]<-"Park" #match the column names so we can merge
 NamesData<-data.frame(UnitCode=unique(NamesData$UnitCode),Park=unique(NamesData$Park))#get just each park once
 FEdat<-merge(FEdat, NamesData, by = "Park") #merge
 
-
+write.csv(FEdat,file = "ParkFees.csv")
 #plotting examples
 FEdat%>%
   ggplot(., aes(x=Year,y=SingleVisitorFee,by=UnitCode))+  #note that when importing data from a pipe %>% into ggplot

### 2020-05-15 CrossValidation.R
added cross validation. cor=0.36..pretty good! for just smoke

@@ -1,0 +1,58 @@
+library(tidyverse)
+library(lme4)
+library(rstan)
+Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
+Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
+
+#filter by high season
+Data<-Data%>%filter( SeasType =="High")
+
+#filter by smoke affected regions
+Data<-Data%>%
+  filter(Region %in% c("Intermountain","Pacific West"))
+
+
+#get vis trends
+
+fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
+Data$trendvis<-predict(fits,date=Data$date)
+Data$VisDiff<-Data$RecreationVisits-Data$trendvis
+dat<-Data
+
+#pull out last 5 years
+train<-dat%>%filter(!Year %in% c("2018","2017","2016","2015","2014"))
+test<-dat%>%filter(Year %in% c("2018","2017","2016","2015","2014"))
+
+library(rstanarm)
+options(mc.cores=parallel::detectCores())
+
+#fitTrain<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+save(fitTrain,file="CVResidualMod.rds")
+preds<-apply(posterior_predict(fitTrain,test),2,median)
+reals<-test$VisDiff
+
+d<-as.data.frame(cbind(as.vector(preds),reals))
+names(d)<-c("preds","reals")
+mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+                plot.title = element_text( size=18, color="black",face="bold"),
+                axis.title.x = element_text( size=18),
+                axis.title.y = element_text( size=18),
+                axis.text=(element_text(color="black", size=14)),
+                legend.title = element_text(colour="black", size=18),
+                legend.text = element_text( size = 14))
+d%>%
+  ggplot(.,aes(x=reals,y=preds))+
+  geom_jitter(size=2, alpha=0.3)+
+  geom_smooth(method="lm", se=FALSE, colour="red")+
+  geom_text(aes(x=-1e+05,y=1e+05),size=10,label=paste("cor =",round(cor(d$preds,d$reals),digits=2)))+
+  ggtitle("CV Results")+theme_classic()+mytheme+
+  theme(plot.title = element_text(hjust = 0.5))
+
+
+
+
+
+
+
+
+

### 2020-05-15 HESBoelWorkingScript.R

@@ -3,9 +3,9 @@ library(lme4)
 library(rstan)
 Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
 Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
-Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
-Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
-Data<-na.omit(Data)
+#Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
+#Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
+#Data<-na.omit(Data)
 
 #filter by high season
 Data<-Data%>%filter( SeasType =="High")

### 2020-05-18 Markdown Documents/Manuscript.Rmd
moving things around

--- 2020-05-15 Manuscript.Rmd
@@
 
-```{r echo=FALSE, cache=TRUE}
-head(FEdat[c(1009,256,10,800,655),])
-```
-
 We removed the NA's from this dataset built a linear model which estimates the fees based on the date and the park. 
@@
 We operationalized this model in rstanarm
-```{r eval=FALSE}
-feefit<-stan_glm(SingleVisitorEntryFee~Year+Park,family="Poisson",data=d,chains=3,iter=6000)
-```
 
 We then used this to predict what the fees would have been for every year we didn't have
-```{r cache=TRUE}
+```{r cache=TRUE, include=FALSE}
 fillData<-FEdat[,1:2]
 
-FEdat$PredictedFee<-round(apply(posterior_predict(feefit,fillData),2,mean),digits = 0) #assuming this is for a new column
+FEdat$PredictedFee<-apply(posterior_predict(feefit,fillData),2,mean) #assuming this is for a new column
 
@@
 xx%>%
-  ggplot(.,aes(x=SingleVisitorEntryFee,y=round(PredictedFee),digits=0))+
+  ggplot(.,aes(x=SingleVisitorEntryFee,y=PredictedFee))+
   geom_jitter(size=2, alpha=0.3)+
   geom_smooth(method="lm", se=FALSE, colour="red")+
-  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(round(xx$PredictedFee,digits=0),xx$SingleVisitorEntryFee),digits=2)))+
+  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(xx$PredictedFee,xx$SingleVisitorEntryFee),digits=2)))+
   ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+
@@
 
-```{r warning=FALSE, cache=TRUE}
+```{r warning=FALSE, include=FALSE, cache=TRUE}
 fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
@@
 This was also operationalized in `rstanarm`
-```{r eval=FALSE, cache=TRUE}
+```{r eval=FALSE, cache=TRUE,include=FALSE}
 fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=dat,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
@@
 
-### 2.2.2. Smoke, Fee, & Visitation Projections
+### 2.2.2. Model Validation
+
+### 2.2.3. Smoke, Fee, & Visitation Projections
 
@@
 
+## 3.1. Parameter Estimates
 For all significant effects (90% C.I. doesn't overlap 0), we see a negative effect of smoke on the residual visitation to National Parks.
@@
 
+## 3.2. Backcasting
 We can then go back and apply these estimates effects to the observed visitation given a minimum level of smoke observed for each park to estimate what the visitation may have been given optimal air quality.
@@
 VEstDat<-dat%>%group_by(UnitCode,Year)%>%
-  summarise(RecVisits=sum(RecreationVisits),PredVisits = sum(PredNoSmoke))%>%
+  summarise(RecVisits=sum(RecreationVisits),PredVisits = sum(PredNoSmoke50CI))%>%
   filter(UnitCode %in% fitsig$Par)
@@
 ```
+## 3.3. Model Validation
+
+## 3.4. Future Projections
 

### 2020-05-18 Markdown Documents/Manuscript.Rmd
Started Cross validation section

@@
 load("~/SmokeProject/ModelObjects/residuals.rda")
+load("~/SmokeProject/ModelObjects/CVResidualMod.rds")
 ```
@@
 ## 3.3. Model Validation
+To assess the validity of our model, we used a k-fold cross validation technique.
 

### 2020-05-18 Rcode/CrossValidation.R

@@ -27,7 +27,9 @@ library(rstanarm)
 options(mc.cores=parallel::detectCores())
 
 #fitTrain<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
-save(fitTrain,file="CVResidualMod.rds")
+#save(fitTrain,file="CVResidualMod.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod.rds")
+
 preds<-apply(posterior_predict(fitTrain,test),2,median)
 reals<-test$VisDiff
 

### 2020-05-18 Rcode/CrossValidation.R
added k fold validation. k=7

@@ -20,18 +20,86 @@ Data$VisDiff<-Data$RecreationVisits-Data$trendvis
 dat<-Data
 
 #pull out last 5 years
-train<-dat%>%filter(!Year %in% c("2018","2017","2016","2015","2014"))
-test<-dat%>%filter(Year %in% c("2018","2017","2016","2015","2014"))
+train1<-dat%>%filter(!Year %in% c("2018","2017","2016","2015","2014"))
+test1<-dat%>%filter(Year %in% c("2018","2017","2016","2015","2014"))
 
 library(rstanarm)
 options(mc.cores=parallel::detectCores())
 
-#fitTrain<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#fitTrain<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train1,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
 #save(fitTrain,file="CVResidualMod.rds")
 load("~/SmokeProject/ModelObjects/CVResidualMod.rds")
 
-preds<-apply(posterior_predict(fitTrain,test),2,median)
-reals<-test$VisDiff
+preds1<-apply(posterior_predict(fitTrain,test1),2,median)
+reals1<-test1$VisDiff
+
+
+## 2013:2009
+
+train2<-dat%>%filter(!Year %in% c("2013","2012","2011","2010","2009"))
+test2<-dat%>%filter(Year %in% c("2013","2012","2011","2010","2009"))
+fitTrain2<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train2,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+save(fitTrain2,file="CVResidualMod2.rds")
+#load("~/SmokeProject/ModelObjects/CVResidualMod2.rds")
+
+preds2<-apply(posterior_predict(fitTrain2,test2),2,median)
+reals2<-test2$VisDiff
+
+# 2008:2004
+train3<-dat%>%filter(!Year %in% c("2008","2007","2006","2005","2004"))
+test3<-dat%>%filter(Year %in% c("2008","2007","2006","2005","2004"))
+fitTrain3<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train3,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
[…]

### 2020-05-18 Markdown Documents/Manuscript.Rmd

@@
 library(rstan)
-Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
+Data<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")[,-1]
 Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")

### 2020-05-18 Markdown Documents/Manuscript.Rmd
added outline to the intro

@@
 # 1. Introduction
-Public lands are a cornerstone of American culture and identity.
-The crown jewels of the American public lands system are the National Parks.
-National Parks have recently drawn public attention for having a lack of funding which negatively affects their ability to protect America's most cherished natural resources and provide visitors with desired recreational experiences.
-While the National Park Service makes countless decisions regarding where to allocate resources and how to best invest its limited funding, other monetary sources and sinks are out of the control of the National Park Service.
+
 This is how you add a reference (e.g., [@fisichelli_protected_2015; @fann_ch_2016])
+
+Paragraphs:
+1. Public lands are important
+2. NPs are specifically high value
+3. NPs are under funded. This underfunding prevents them from doing those important jobs
+4. NPs get a lot of their funding from visitation. Vis is influenced by a lot of things. One of these things that's particurlarly out of their control is smoke from adjacent forested areas.
+5. Smoke is bad for recreation and should decrease vis.
+6. Smoke events have been increasing over the last 50 years. Particularly, these increases have been in the west.
+7. These smoke events are expected to worsen via climate change over the next 50 years.
+8. Logically, smoke should influence visitation to parks. This decrease should affect income. We seek to assess this effect in a systematic way and backcast what the monetary effects might have been.
+9. Given the trends we observe, we project out into the future what the effects might be over the next 50 years.
 

### 2020-05-18 Rcode/CrossValidation.R
minor changes

@@ -38,9 +38,9 @@ reals1<-test1$VisDiff
 
 train2<-dat%>%filter(!Year %in% c("2013","2012","2011","2010","2009"))
 test2<-dat%>%filter(Year %in% c("2013","2012","2011","2010","2009"))
-fitTrain2<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train2,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
-save(fitTrain2,file="CVResidualMod2.rds")
-#load("~/SmokeProject/ModelObjects/CVResidualMod2.rds")
+#fitTrain2<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train2,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain2,file="CVResidualMod2.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod2.rds")
 
 preds2<-apply(posterior_predict(fitTrain2,test2),2,median)
 reals2<-test2$VisDiff
@@ -48,9 +48,9 @@ reals2<-test2$VisDiff
 # 2008:2004
 train3<-dat%>%filter(!Year %in% c("2008","2007","2006","2005","2004"))
 test3<-dat%>%filter(Year %in% c("2008","2007","2006","2005","2004"))
-fitTrain3<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train3,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
-save(fitTrain3,file="CVResidualMod3.rds")
-#load("~/SmokeProject/ModelObjects/CVResidualMod3.rds")
+#fitTrain3<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train3,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain3,file="CVResidualMod3.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod3.rds")
 
 preds3<-apply(posterior_predict(fitTrain3,test3),2,median)
 reals3<-test3$VisDiff
@@ -96,8 +96,8 @@ preds7<-apply(posterior_predict(fitTrain7,test7),2,median)
 reals7<-test7$VisDiff
 
 #combine them all
-preds<-c(preds1,preds2,preds3,preds4,preds5)
-reals<-c(reals1,reals2,reals3,reals4,reals5)
[…]

### 2020-05-18 Rcode/MeanSmokePlots.R
Deleting old stuff

@@ -1,68 +1,0 @@
-library(tidyverse)
-dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
-
-
-ggplot(data=dat,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
-  geom_point(alpha=0.2)+
-  geom_smooth(method="lm",color="red",se=FALSE)+
-  theme_classic()+ggtitle("max")
-
-ggplot(data=dat,aes(x=stdsmoke,y=RecreationVisits,by=UnitCode))+
-  geom_point(alpha=0.2)+
-  geom_smooth(color="red",se=FALSE,alpha=0.1)+
-  theme_classic()+ggtitle("max")
-
-
-
-
-ggplot(data=dat,aes(x=Year,y=stdsmoke,by=UnitCode))+
-  geom_jitter(alpha=0.2)+
-  geom_smooth(method="lm",color="red",se=FALSE,alpha=0.2)+
-  theme_classic()
-
-ggplot(data=dat,aes(x=Year,y=RecreationVisits,by=UnitCode))+
-  geom_jitter(alpha=0.2)+
-  geom_smooth(method="lm",color="red",se=FALSE,alpha=0.2)+
-  theme_classic()
-
-
-
-dat$LE<- ifelse(datmax$stdsmoke >=2,"True","False")
-
-ggplot(data=dat,aes(x=LE,y=RecreationVisits))+
-  geom_boxplot()+
-  theme_classic()
-
-
-stdize<-function(x) {return((x-mean(x)/(2*sd(x))))}
-
-dat<-dat%>%group_by(UnitCode)%>%
-  mutate(stdsmokepark=stdize(stdsmoke))
-
-ggplot(data=datmax,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
-  geom_point(alpha=0.2)+
-  geom_smooth(method="lm",color="red",se=FALSE,alpha=0.2)+
-  theme_classic()+ggtitle("max")
-
-ggplot(data=datmax,aes(x=stdsmokepark,y=RecreationVisits,by=UnitCode))+
-  geom_point(alpha=0.2)+
-  geom_smooth(color="red",se=FALSE,alpha=0.2)+
-  theme_classic()+ggtitle("max")
-
-
-#stop here
-
-library(zoo)
-dat$Date<-zoo::as.yearmon(paste(as.numeric(dat$Year),
[…]

### 2020-05-18 Rcode/RegressionFirstCut.R

@@ -1,76 +1,0 @@
-library(tidyverse)
-library(lme4)
-library(rstanarm)
-library(MASS)
-options(mc.cores = parallel::detectCores())
-dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
-dat$Year<-as.factor(dat$Year)
-#models below increase in complexity..Currently, non of  them are converging using 
-#the lme4 package or with rstanarm
-#this may be a computational problem on my end though, rather than a model specification problem
-
-#glm with independent intercept for each park for each season
-fit<-glmer.nb(RecreationVisits~stdsmoke+CatColS+(1|CatColS),data=dat)
-
-#glm with independent intercept for each park for each Month
-fit2<-glmer.nb(RecreationVisits~stdsmoke+CatColM+(1|CatColM),data=dat)
-
-
-#glm with independent intercept for each park for each Season and
-# random slope for smoke for each park
-fit3<-glmer.nb(RecreationVisits~stdsmoke+CatColS+(stdsmoke|CatColS),data=dat)
-
-#glm with independent intercept for each park for each Month and
-# random slope for smoke for each park
-#this is the preferred model
-fit4<-glmer.nb(RecreationVisits~stdsmoke+CatColM+(stdsmoke|CatColM),data=dat)
-
-
-#any of these models can be run using stan by specifying 
-#stan_glmer.nb instead of the glmer.nb function from lme4
-
-
-fit5<-glmer.nb(RecreationVisits~(1|CatColS)+(stdsmoke|UnitCode), data=dat)
-save(fit5,file="fit5.rda")
-summary(fit5)
-
-#best
-fit6<-glmer.nb(RecreationVisits~(1|CatColM)+(stdsmoke|UnitCode), data=dat)
-save(fit6,file="fit6.rda")
-
-fit7<-glmer.nb(RecreationVisits~(stdsmoke|CatColM), data=dat)
-save(fit7,file="fit7.rda")
-
-fit8<-glm.nb(RecreationVisits~stdsmoke+Month+UnitCode,data=dat)
-save(fit8,file="fit8.rda")
-
-fit9<-glm.nb(RecreationVisits~stdsmoke+Season+UnitCode,data=dat)
-save(fit9,file="fit9.rda")
-
-
-coef(fit6)
-
-fit<-glmer.nb(RecreationVisits~
-                     (1|CatColM)+
-                     (stdsmoke|UnitCode)+
-                     Year,
-                   data=dat)
-
-
-
-
[…]

### 2020-05-18 Rcode/Sloppy plots and models.R

@@ -1,63 +1,0 @@
-library(tidyverse)
-library(lme4)
-library(rstanarm)
-library(MASS)
-dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
-
-fit2<-glmer.nb(RecreationVisits~stdsmoke+(Season|UnitCode),data=dat)
-
-fit3<-glmer.nb(RecreationVisits~ScaledSmoke+(1|CatColM),data=dat)
-fit4<-glm(RecreationVisits~Smoke,family="poisson",data=dat)
-
-newdata <- data.frame(
-  ScaledSmoke = seq(from = min(dat[dat$CatColM=="YOSE07",]$Smoke), to = max(dat[dat$CatColM=="YOSE07",]$Smoke), length.out = 100),
-  CatColM = rep("YOSE07",100))
-  
-newdata2 <- cbind(newdata, predict(fit3, newdata))
-
-nd<-data.frame(Smoke=0.00316, CatColM="ARCH06")
-predict(fit2, nd)
-
-
-
-ggplot(data=dat[1:6959,],aes(x=Smoke,y=RecreationVisits))+
-  geom_smooth(aes(color=Month),method="lm", se=F)+
-facet_wrap(~ParkName, scales="free")
-  
-ggplot(data=dat[6960:13955,],aes(x=Smoke,y=RecreationVisits))+
-  geom_smooth(aes(color=Month),method="lm", se=F)+
-  facet_wrap(~ParkName,, scales="free")
-
-ggplot(data=dat[13956:20891,],aes(x=Smoke,y=RecreationVisits))+
-  geom_smooth(aes(color=Month),method="lm", se=F)+
-  facet_wrap(~ParkName,, scales="free")
-
-ggplot(data=dat[20892:nrow(dat),],aes(x=Smoke,y=RecreationVisits))+
-  geom_smooth(aes(color=Month),method="lm", se=F)+
-  facet_wrap(~ParkName, scales="free")
-  
-ggplot(data=dat[dat$UnitCode %in% c("YOSE", "YELL","ACAD","GRSM","JOTR","SEQU"),],aes(x=Smoke,y=RecreationVisits))+
-  geom_smooth(aes(color=Month),method="lm", se=F)+
-  geom_point(aes(color=Month))+facet_wrap(~ParkName, scales="free")
-  
[…]

### 2020-05-18 Rcode/TempStdMod.R

@@ -1,39 +1,0 @@
-dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
-x<-dat%>%group_by(UnitCode,Season)%>%
-  summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
-  group_by(UnitCode)%>%filter(mv==max(mv))%>%mutate(cc=paste0(UnitCode,Season))
-
-x2<-dat%>%group_by(UnitCode,Season)%>%
-  summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
-  group_by(UnitCode)%>%filter(mv==min(mv))%>%mutate(cc=paste0(UnitCode,Season))
-
-dat$SeasType<-ifelse(dat$CatColS %in% x$cc,
-                     "High",ifelse(dat$CatColS %in% x2$cc,"Low","Shoulder"))
-
-stdize<-function(x) {return((x-mean(x)/(2*sd(x))))}
-
-dat<-dat%>%group_by(as.character(Year))%>%mutate(stdsmoketime=stdize(Smoke*10e8))
-
-
-dat<-dat%>%filter(SeasType =="High")
-
-data_list <- list(
-  N = nrow(dat),
-  Nprk = length(unique(dat$UnitCode)),
-  count = dat$RecreationVisits,
-  smoke = dat$stdsmoketime,
-  pcode = as.numeric(as.factor(dat$UnitCode )))
-
-options(mc.cores=3)
-
-l3<- stan( file="SmokeNegBinom.stan" , data=data_list,chains=3 ,warmup = 1000 ,iter = 10000)
-
-print( l3 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
-
-save(l3, file="stdsmokeTimeNbinom.rda")
-stan_plot(l3,pars = c(zz) ,fill_color = "purple", )
-
-mean(as.data.frame(summary(l3))[2:61,1])
-
-
-

### 2020-05-18 Rcode/VariousARModels.R

@@ -1,131 +1,0 @@
-library(tidyverse)
-library(rstan)
-Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
-
-
-
-#add smoke season
-x<-Data%>%group_by(UnitCode,Season)%>%
-  summarise(ms=mean(Smoke))%>%ungroup()%>%
-  group_by(UnitCode)%>%filter(ms==max(ms))%>%mutate(cc=paste0(UnitCode,Season))
-
-
-
-Data$SmokeSeas<-ifelse(Data$CatColS %in% x$cc,
-                      "High","NotHigh")
-
-
-
-#Data<-Data%>%filter( SmokeSeas =="High")
-Data<-Data%>%filter( SeasType =="High")
-
-Data<-Data%>%
-  filter(Region %in% c("Intermountain","Pacific West"))
-##
-
-dat<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
-
-dat<-na.omit(dat)
-
-dat$stdARV<-as.vector(scale(dat$ARVis))
-
-
-data_list2 <- list(
-  N = nrow(dat),
-  previs = dat$stdARV,
-  Nprk = length(unique(dat$UnitCode)),
-  count = dat$RecreationVisits,
-  smoke = dat$stdsmoke,
-  pcode = as.numeric(as.factor(dat$UnitCode )))
-
-options(mc.cores=parallel::detectCores()-1)
-
-A3<- stan( file="~/SmokeProject/StanCode/SmokeNegBinomAutoregressive.stan" , 
-           data=data_list2,chains=3 ,warmup=1000,iter = 8000, control = list(max_treedepth = 25))
-
-
-
-print( A3 , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
-
-save(A3, file="stdsmokeNbinomAutoregressiveSubset.rda")
-
-x<-as.data.frame(summary(A3))
-zz<-row.names(x[2:34,])
-stan_plot(A3,pars = c(zz) ,fill_color = "purple", )
-
-## no AR term
-dat<-Data
-
-data_list3 <- list(
-  N = nrow(dat),
-  Nprk = length(unique(dat$UnitCode)),
-  count = dat$RecreationVisits,
-  smoke = dat$stdsmoke,
-  pcode = as.numeric(as.factor(dat$UnitCode )))
-
[…]

### 2020-05-18 Rcode/scrappyMods.R

@@ -1,33 +1,0 @@
-library(tidyverse)
-library(lme4)
-library(rstanarm)
-library(MASS)
-
-dat<-read_csv("~/Smoke_Proj/Data/MergedDataComplete.csv")
-dat$Year<-as.factor(dat$Year)
-dat$catY<-paste0(dat$UnitCode,dat$Year)
-MattDat<-dat[seq(1,nrow(dat),10),]
-
-
-fitM<-stan_glmer.nb(RecreationVisits~(stdsmoke|UnitCode)+(Season|catY), chains=1,data=MattDat)
-coef(fitM)
-
-fitM<-stan_glmer(RecreationVisits~(stdsmoke|UnitCode)+(Season|catY),
-                 family="poisson",chains=1,data=MattDat)
-coef(fitM)
-fitz<-stan_glmer(RecreationVisits~(stdsmoke|UnitCode)+
-                      (UnitCode:Season|Year), chains=1,family="poisson",data=MattDat)
-
-MattDat<-dat[seq(1,nrow(dat),10),]
-
-
-
-
-
-
-
-
-
-
-
-

### 2020-05-19 Rcode/CrossValidation.R
Done with CV

@@ -58,9 +58,9 @@ reals3<-test3$VisDiff
 # 2003:1999
 train4<-dat%>%filter(!Year %in% c("2003","2002","2001","2000","1999"))
 test4<-dat%>%filter(Year %in% c("2003","2002","2001","2000","1999"))
-fitTrain4<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train4,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
-save(fitTrain4,file="CVResidualMod4.rds")
-#load("~/SmokeProject/ModelObjects/CVResidualMod4.rds")
+#fitTrain4<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train4,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain4,file="CVResidualMod4.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod4.rds")
 
 preds4<-apply(posterior_predict(fitTrain4,test4),2,median)
 reals4<-test4$VisDiff
@@ -68,9 +68,9 @@ reals4<-test4$VisDiff
 # 1998:1994
 train5<-dat%>%filter(!Year %in% c("1998","1997","1996","1995","1994"))
 test5<-dat%>%filter(Year %in% c("1998","1997","1996","1995","1994"))
-fitTrain5<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train5,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
-save(fitTrain5,file="CVResidualMod5.rds")
-#load("~/SmokeProject/ModelObjects/CVResidualMod5.rds")
+#fitTrain5<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train5,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain5,file="CVResidualMod5.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod5.rds")
 
 preds5<-apply(posterior_predict(fitTrain5,test5),2,median)
 reals5<-test5$VisDiff
@@ -78,9 +78,9 @@ reals5<-test5$VisDiff
 # 1993:1989
 train6<-dat%>%filter(!Year %in% c("1993","1992","1991","1990","1989"))
 test6<-dat%>%filter(Year %in% c("1993","1992","1991","1990","1989"))
[…]

### 2020-05-19 Markdown Documents/Manuscript.Rmd
added train/test figure
added CV table

@@
 2. NPs are specifically high value
-3. NPs are under funded. This underfunding prevents them from doing those important jobs
-4. NPs get a lot of their funding from visitation. Vis is influenced by a lot of things. One of these things that's particurlarly out of their control is smoke from adjacent forested areas.
+3. NPs are under funded. This under funding prevents them from doing those important jobs
+4. NPs get a lot of their funding from visitation. Vis is influenced by a lot of things. One of these things that's particularly out of their control is smoke from adjacent forested areas.
 5. Smoke is bad for recreation and should decrease vis.
@@
   geom_jitter(size=2, alpha=0.3)+
-  geom_smooth(method="lm", se=FALSE, colour="red")+
+  geom_smooth(method="lm", se=FALSE, colour="black",size=2,linetype="longdash")+
   geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(xx$PredictedFee,xx$SingleVisitorEntryFee),digits=2)))+
@@
 
-To account for the linear trends of visitation, we first calculated the trends, then found the residuals of each datapoint from the trendline.
+To account for the linear trends of visitation, we first calculated the trends, then found the residuals of each datapoint from the trend line.
 $$
@@
 
+```{r echo=FALSE, message=FALSE,cache=TRUE}
+knitr::include_graphics(path.expand("~/SmokeProject/Train_Test.png"))
+```
+
 ### 2.2.3. Smoke, Fee, & Visitation Projections
@@
 
-```{r echo=FALSE, cache=TRUE}
- plot(fit, regex_pars = "stdsmoke UnitCode",
+```{r echo=FALSE, cache=TRUE, message=FALSE,warning=FALSE}
+PNames<-rev(unique(dat$ParkName))
+
+pars<-plot(fit, regex_pars = "stdsmoke UnitCode",
      prob = 0.5, prob_outer = 0.9)
+
+mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+                plot.title = element_text( size=18, color="black",face="bold"),
+                axis.title.x = element_text( size=18),
+                axis.title.y = element_text( size=18),
+                axis.text=(element_text(color="black", size=14)),
+                legend.title = element_text(colour="black", size=18),
+                legend.text = element_text( size = 14))
+```
+
+```{r echo=FALSE, cache=TRUE, message=FALSE,warning=FALSE, fig.height = 9, fig.width = 12}
+pars+ggplot2::ggtitle("Posterior medians with 50% and 90% intervals")+
+  ggplot2::scale_y_discrete(name ="Park",labels=PNames)+
+  mytheme+scale_x_continuous(name="Parameter Value", labels = scales::comma)
 ```
@@
 
-```{r echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE}
+```{r echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE, fig.height = 6, fig.width = 12}
 library(scales)
@@
 
+```{r echo=FALSE, message=FALSE,cache=TRUE}
+library(tidyverse)
+library(rstanarm)
+#pull out last 5 years
+train1<-dat%>%filter(!Year %in% c("2018","2017","2016","2015","2014"))
+test1<-dat%>%filter(Year %in% c("2018","2017","2016","2015","2014"))
+
+#fitTrain<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train1,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain,file="CVResidualMod.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod.rds")
+
+preds1<-apply(posterior_predict(fitTrain,test1),2,median)
+reals1<-test1$VisDiff
+
+## 2013:2009
+
+train2<-dat%>%filter(!Year %in% c("2013","2012","2011","2010","2009"))
+test2<-dat%>%filter(Year %in% c("2013","2012","2011","2010","2009"))
+#fitTrain2<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train2,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain2,file="CVResidualMod2.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod2.rds")
+
+preds2<-apply(posterior_predict(fitTrain2,test2),2,median)
+reals2<-test2$VisDiff
+
+# 2008:2004
+train3<-dat%>%filter(!Year %in% c("2008","2007","2006","2005","2004"))
+test3<-dat%>%filter(Year %in% c("2008","2007","2006","2005","2004"))
+#fitTrain3<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train3,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain3,file="CVResidualMod3.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod3.rds")
+
+preds3<-apply(posterior_predict(fitTrain3,test3),2,median)
+reals3<-test3$VisDiff
+
+# 2003:1999
+train4<-dat%>%filter(!Year %in% c("2003","2002","2001","2000","1999"))
+test4<-dat%>%filter(Year %in% c("2003","2002","2001","2000","1999"))
+#fitTrain4<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train4,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain4,file="CVResidualMod4.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod4.rds")
+
+preds4<-apply(posterior_predict(fitTrain4,test4),2,median)
+reals4<-test4$VisDiff
+
+# 1998:1994
+train5<-dat%>%filter(!Year %in% c("1998","1997","1996","1995","1994"))
+test5<-dat%>%filter(Year %in% c("1998","1997","1996","1995","1994"))
+#fitTrain5<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train5,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain5,file="CVResidualMod5.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod5.rds")
+
+preds5<-apply(posterior_predict(fitTrain5,test5),2,median)
+reals5<-test5$VisDiff
+
+# 1993:1989
+train6<-dat%>%filter(!Year %in% c("1993","1992","1991","1990","1989"))
+test6<-dat%>%filter(Year %in% c("1993","1992","1991","1990","1989"))
+#fitTrain6<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train6,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain6,file="CVResidualMod6.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod6.rds")
+
+preds6<-apply(posterior_predict(fitTrain6,test6),2,median)
+reals6<-test6$VisDiff
+
+# 1988:1980
+train7<-dat%>%filter(!Year %in% c("1988","1987","1986","1985","1984","1983","1982","1981","1980"))
+test7<-dat%>%filter(Year %in% c("1988","1987","1986","1985","1984","1983","1982","1981","1980"))
+#fitTrain7<-stan_lmer(VisDiff~stdsmoke|UnitCode,data=train7,adapt_delta = 0.99,chains=8,warmup=2000,iter=8000)
+#save(fitTrain7,file="CVResidualMod7.rds")
+load("~/SmokeProject/ModelObjects/CVResidualMod7.rds")
+
+preds7<-apply(posterior_predict(fitTrain7,test7),2,median)
+reals7<-test7$VisDiff
+
+#combine them all
+preds<-c(preds1,preds2,preds3,preds4,preds5,preds6,preds7)
+reals<-c(reals1,reals2,reals3,reals4,reals5,reals6,reals7)
+
+d<-as.data.frame(cbind(as.vector(preds),reals))
+names(d)<-c("preds","reals")
+mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+                plot.title = element_text( size=18, color="black",face="bold"),
+                axis.title.x = element_text( size=18),
+                axis.title.y = element_text( size=18),
+                axis.text=(element_text(color="black", size=14)),
+                legend.title = element_text(colour="black", size=18),
+                legend.text = element_text( size = 14))
+d%>%
+  ggplot(.,aes(x=reals,y=preds))+
+  geom_point(size=2.5, alpha=0.2)+
+  geom_smooth(method="lm", se=FALSE, colour="black",size=2,linetype="longdash")+
+  geom_text(aes(x=3e+05,y=-2e+05),size=10,label=paste("cor =",round(cor(d$preds,d$reals),digits=2)))+
+  ggtitle("CV Results")+theme_classic()+mytheme+scale_x_continuous("Observed Residual Visitation", labels = scales::comma)+
+  scale_y_continuous(name="Predicted Residual Visitation", labels = scales::comma)+
+  theme(plot.title = element_text(hjust = 0.5))
+
+```
+
+```{r echo=FALSE, message=FALSE,cache=TRUE}
+x<-data.frame(Model=c(1:7), 
+              "TimePeriod"=c("1980 - 1988",
+                                         "1989 - 1993",
+                                         "1994 - 1998",
+                                         "1999 - 2003",
+                           "2004 - 2008",
+                           "2009 - 2013",
+                           "2014 - 2018"), 
+              Correlation=c(cor(as.vector(preds1),reals1),
+                                                   cor(as.vector(preds2),reals2),
+                                                   cor(as.vector(preds3),reals3),
+                                                   cor(as.vector(preds4),reals4),
+                                                   cor(as.vector(preds5),reals5),
+                                                   cor(as.vector(preds6),reals6),
+                                                   cor(as.vector(preds7),reals7)))
+
+x$r2<-round(x$Correlation^2,digits=2)
+x$Correlation<-round(x$Correlation,digits=2)
+names(x)<-c("Model","Time Period", "Correlation", "R^2")
+formattable::formattable(x)
+```
 ## 3.4. Future Projections

### 2020-05-19 Rcode/CrossValidation.R
added CV table

@@ -108,6 +108,28 @@ cor(as.vector(preds5),reals5)
 cor(as.vector(preds6),reals6)
 cor(as.vector(preds7),reals7)
 
+x<-data.frame(Model=c(1:7), 
+              "TimePeriod"=c("1980 - 1988",
+                                         "1989 - 1993",
+                                         "1994 - 1998",
+                                         "1999 - 2003",
+                           "2004 - 2008",
+                           "2009 - 2013",
+                           "2014 - 2018"), 
+              Correlation=c(cor(as.vector(preds1),reals1),
+                                                   cor(as.vector(preds2),reals2),
+                                                   cor(as.vector(preds3),reals3),
+                                                   cor(as.vector(preds4),reals4),
+                                                   cor(as.vector(preds5),reals5),
+                                                   cor(as.vector(preds6),reals6),
+                                                   cor(as.vector(preds7),reals7)))
+
+x$r2<-round(x$Correlation^2,digits=2)
+x$Correlation<-round(x$Correlation,digits=2)
+names(x)<-c("Model","Time Period", "Correlation", "R^2")
+formattable::formattable(x)
+
+
 
 d<-as.data.frame(cbind(as.vector(preds),reals))
 names(d)<-c("preds","reals")

### 2020-05-19 Rcode/apply parameter estimates to vis data.R
nothing

@@ -39,6 +39,7 @@ ri<-as.data.frame(ran_intercept_post)
 
 
 
+
 ##just one estimate
 gen_quant_r <- function(x,y) { #x here is the input smoke values #y is park
   lin_comb <- mean(intercept_post) + 

### 2020-05-19 Markdown Documents/Manuscript.Rmd
Plots

@@
 Wildfires and the smoke it produces is supposed to increase over the next 50 years.
-Assuming a linear trend in smoke, visitation overall, and park fees, we might see a financial loss to the park service of $XX.
+Assuming a linear trend in smoke, visitation overall, and park fees, we expect to see significant financial losses to the National Park Service.
 Our recomendation is that....NPS work with the forest service to prevent fires in areas close to parks??
@@
 ```
-
-### 2.2.3. Smoke, Fee, & Visitation Projections
 
@@
 ```
-## 3.4. Future Projections
 

### 2020-05-19 Rcode/PlotsForPaper.R

@@ -1,17 +1,17 @@
 library(tidyverse)
 allData<-read_csv("Data/MergedDataComplete.csv")[,-1]
-predDat<-read_csv("~/SmokeProject/Data/VisitationEstimatesMinSmokeSignificantParks.csv")[,-1]
+predDat<-read_csv("~/SmokeProject/Data/FullDataWithMinSmokePredictions.csv")[,-1]
 predDat$date<-zoo::as.yearmon(predDat$date)
 allData$date<-zoo::as.yearmon(paste0(allData$Month,allData$Year),"%m%Y")
 allData<-allData%>% filter(SeasType =="High")
  #all sig parks
 predDat.interp <- predDat %>% 
    split(.$UnitCode) %>% 
-   map_df(~data.frame(UnitCode = approx(.x$date, .x$No_Smoke_Mean_est, n = 80), 
+   map_df(~data.frame(UnitCode = approx(.x$date, .x$PredNoSmoke50CI, n = 80), 
                       nat = approx(.x$date, .x$RecreationVisits, n = 80), 
                       UnitCode = .x$UnitCode[1]))
 
-
+cols <- c("Observed greater" = "#b2df8a", "Predicted greater" = "#1f78b4")
 ggplot(predDat.interp, aes(nat.x, nat.y)) +  facet_wrap(~UnitCode,scales="free") +
   geom_ribbon(aes(ymin = nat.y, ymax = pmin(UnitCode.y, nat.y), fill = "Observed greater"),alpha=0.9) +
   geom_ribbon(aes(ymin = UnitCode.y, ymax = pmin(UnitCode.y, nat.y), fill = "Predicted greater"),alpha=0.9) +
@@ -199,4 +199,42 @@ allData<-allData%>%filter(UnitCode %in% unique(predDat$UnitCode))
     xlab("") +
     ylab("Value of Y")
   
+  
+  
+  
+  
+  
+  ## 
+  
+  p1<-predDat%>%filter(UnitCode == "ROMO")%>%
+    ggplot(.,aes(x=date))+
+    geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
+    geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
+    geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
[…]

### 2020-05-20 Rcode/PlotsForPaper.R
formatting/tweaking plots

@@ -1,6 +1,6 @@
 library(tidyverse)
 allData<-read_csv("Data/MergedDataComplete.csv")[,-1]
-predDat<-read_csv("~/SmokeProject/Data/FullDataWithMinSmokePredictions.csv")[,-1]
+predDat<-read_csv("Data/FullDataWithMinSmokePredictions.csv")[,-1]
 predDat$date<-zoo::as.yearmon(predDat$date)
 allData$date<-zoo::as.yearmon(paste0(allData$Month,allData$Year),"%m%Y")
 allData<-allData%>% filter(SeasType =="High")
@@ -206,25 +206,30 @@ allData<-allData%>%filter(UnitCode %in% unique(predDat$UnitCode))
   
   ## 
   
-  p1<-predDat%>%filter(UnitCode == "ROMO")%>%
+  p1<-predDat%>%filter(UnitCode == "OLYM")%>%
     ggplot(.,aes(x=date))+
     geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
     geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
-    geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
+    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
+    geom_point(aes(y=RecreationVisits,color="black"),size=2)+
     geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
-    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
+    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
+    scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
     scale_color_identity(name = "",
                          breaks = c("black", "#1f78b4"),
-                         labels = c("Observed", "Median Estimate"),
-                         guide = "legend")+ggtitle("Rocky Mountain National Park")+theme(legend.position="top")
+                         labels = c("Observed", "Median Estimate Minimum Smoke"),
[…]

### 2020-05-20 Markdown Documents/Manuscript.Rmd
Updated Plots

@@
 library(scales)
-dat%>%filter(UnitCode =="ROMO")%>%
+p1<-predDat%>%filter(UnitCode == "GLAC")%>%
     ggplot(.,aes(x=date))+
@@
     geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
-  geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
+    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
+    geom_point(aes(y=RecreationVisits,color="black"),size=2)+
     geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
-  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
+    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
+    scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
     scale_color_identity(name = "",
                          breaks = c("black", "#1f78b4"),
-                       labels = c("Observed", "Median Estimate"),
-                       guide = "legend")+ggtitle("Rocky Mountain National Park")
+                         labels = c("Observed", "Median Estimate Minimum Smoke"),
+                         guide = "legend")+ggtitle("Glacier National Park")+theme(legend.position="top")+
+    theme(axis.title.x=element_blank(),
+          axis.text.x=element_blank(),
+          axis.ticks.x=element_blank(),
+          axis.line.x = element_blank())
 
-dat%>%filter(UnitCode == "YOSE")%>%
+p2<-predDat%>%filter(UnitCode == "GLAC")%>%group_by(Year)%>%
+    summarise(meansmoke=mean(Smoke))%>%
+    ggplot(.,aes(x=Year,y=meansmoke))+
+    geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
+    geom_point(aes(color =(meansmoke)),size=5)+
+    scale_color_gradient2(low ="#2b83ba" , mid = "#ffffbf", high = "#d7191c",name= "Difference from trend",breaks=c(0.050,-0.035),
+                          labels=c("Above","Below")
+    ) + 
+    #guides(color = FALSE) +
+    guides(alpha = FALSE) + 
+    theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
+   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(0, 9e-10))+mytheme
+
+gA <- ggplotGrob(p1)
+gB <- ggplotGrob(p2)
+grid::grid.newpage()
+grid::grid.draw(rbind(gA, gB))
+```
+
+```{r echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE, fig.height = 6, fig.width = 12}
+library(scales)
+p1<-predDat%>%filter(UnitCode == "YELL")%>%
     ggplot(.,aes(x=date))+
@@
     geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
-  geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
+    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
+    geom_point(aes(y=RecreationVisits,color="black"),size=2)+
     geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
-  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
+    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
+    scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
     scale_color_identity(name = "",
                          breaks = c("black", "#1f78b4"),
-                       labels = c("Observed", "Median Estimate"),
-                       guide = "legend")+ggtitle("Yosemite National Park")
+                         labels = c("Observed", "Median Estimate Minimum Smoke"),
+                         guide = "legend")+ggtitle("Yellowstone National Park")+theme(legend.position="top")+
+    theme(axis.title.x=element_blank(),
+          axis.text.x=element_blank(),
+          axis.ticks.x=element_blank(),
+          axis.line.x = element_blank())
 
-dat%>%filter(UnitCode == "ZION")%>%
-ggplot(.,aes(x=date))+
-  geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
-  geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
-  geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
-  geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
-  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
-  scale_color_identity(name = "",
-                       breaks = c("black", "#1f78b4"),
-                       labels = c("Observed", "Median Estimate"),
-                       guide = "legend")+ggtitle("Zion National Park *not significant")
+p2<-predDat%>%filter(UnitCode == "YELL")%>%group_by(Year)%>%
+    summarise(meansmoke=mean(Smoke))%>%
+    ggplot(.,aes(x=Year,y=meansmoke))+
+    geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
+    geom_point(aes(color =(meansmoke)),size=5)+
+    scale_color_gradient2(low ="#2b83ba" , mid = "#ffffbf", high = "#d7191c",name= "Difference from trend",breaks=c(0.050,-0.035),
+                          labels=c("Above","Below")
+    ) + 
+    #guides(color = FALSE) +
+    guides(alpha = FALSE) + 
+    theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
+   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(0, 9e-10))+mytheme
 
+gA <- ggplotGrob(p1)
+gB <- ggplotGrob(p2)
+grid::grid.newpage()
+grid::grid.draw(rbind(gA, gB))
 ```

### 2020-05-20 Markdown Documents/Manuscript.Rmd
new model and plots

--- 2020-05-19 Markdown Documents/Manuscript.Rmd
@@
 ```{r echo=FALSE, cache=TRUE, message=FALSE,warning=FALSE}
-PNames<-rev(unique(dat$ParkName))
+PNames<-unique(dat$ParkName)
+
+pp<-as.data.frame(posterior_interval(fit))
+limits<-row.names(pp[grep("stdsmoke UnitCode",row.names(pp)),])
 
@@
                 legend.text = element_text( size = 14))
+
 ```
@@
 pars+ggplot2::ggtitle("Posterior medians with 50% and 90% intervals")+
-  ggplot2::scale_y_discrete(name ="Park",labels=PNames)+
+  ggplot2::scale_y_discrete(name ="Park",labels=rev(PNames),limits=rev(limits))+
   mytheme+scale_x_continuous(name="Parameter Value", labels = scales::comma)
@@
 library(scales)
-dat%>%filter(UnitCode =="ROMO")%>%
+p1<-dat%>%filter(UnitCode == "GLAC")%>%
     ggplot(.,aes(x=date))+
@@
     geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
-  geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
+    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
+    geom_point(aes(y=RecreationVisits,color="black"),size=2)+
     geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
-  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
+    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
+    scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
     scale_color_identity(name = "",
                          breaks = c("black", "#1f78b4"),
-                       labels = c("Observed", "Median Estimate"),
-                       guide = "legend")+ggtitle("Rocky Mountain National Park")
+                         labels = c("Observed", "Median Estimate Minimum Smoke"),
+                         guide = "legend")+ggtitle("Glacier National Park")+theme(legend.position="top")+
+    theme(axis.title.x=element_blank(),
+          axis.text.x=element_blank(),
+          axis.ticks.x=element_blank(),
+          axis.line.x = element_blank())
 
-dat%>%filter(UnitCode == "YOSE")%>%
+p2<-dat%>%filter(UnitCode == "GLAC")%>%group_by(Year)%>%
+    summarise(meansmoke=mean(Smoke))%>%
+    ggplot(.,aes(x=Year,y=meansmoke))+
+    geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
+    geom_point(aes(color =(meansmoke)),size=5)+
+    scale_color_gradient2(low ="#2b83ba" , mid = "#ffffbf", high = "#d7191c",name= "Difference from trend",breaks=c(0.050,-0.035),
+                          labels=c("Above","Below")
+    ) + 
+    #guides(color = FALSE) +
+    guides(alpha = FALSE) + 
+    theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
+   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(0, 1.25e-9))+mytheme
+
+gA <- ggplotGrob(p1)
+gB <- ggplotGrob(p2)
+grid::grid.newpage()
+grid::grid.draw(rbind(gA, gB))
+```
+
+```{r echo=FALSE, warning=FALSE, cache=TRUE,message=FALSE, fig.height = 6, fig.width = 12}
+library(scales)
+p1<-dat%>%filter(UnitCode == "OLYM")%>%
     ggplot(.,aes(x=date))+
@@
     geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
-  geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
+    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
+    geom_point(aes(y=RecreationVisits,color="black"),size=2)+
     geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
-  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
+    mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
+    scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
     scale_color_identity(name = "",
                          breaks = c("black", "#1f78b4"),
-                       labels = c("Observed", "Median Estimate"),
-                       guide = "legend")+ggtitle("Yosemite National Park")
+                         labels = c("Observed", "Median Estimate Minimum Smoke"),
+                         guide = "legend")+ggtitle("Olympic National Park")+theme(legend.position="top")+
+    theme(axis.title.x=element_blank(),
+          axis.text.x=element_blank(),
+          axis.ticks.x=element_blank(),
+          axis.line.x = element_blank())
 
-dat%>%filter(UnitCode == "ZION")%>%
-ggplot(.,aes(x=date))+
-  geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
-  geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
-  geom_line(aes(y=RecreationVisits,color="black"),size=1.25)+
-  geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
-  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+xlab("Date")+
-  scale_color_identity(name = "",
-                       breaks = c("black", "#1f78b4"),
-                       labels = c("Observed", "Median Estimate"),
-                       guide = "legend")+ggtitle("Zion National Park *not significant")
+p2<-dat%>%filter(UnitCode == "OLYM")%>%group_by(Year)%>%
+    summarise(meansmoke=mean(Smoke))%>%
+    ggplot(.,aes(x=Year,y=meansmoke))+
+    geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
+    geom_point(aes(color =(meansmoke)),size=5)+
+    scale_color_gradient2(low ="#2b83ba" , mid = "#ffffbf", high = "#d7191c",name= "Difference from trend",breaks=c(0.050,-0.035),
+                          labels=c("Above","Below")
+    ) + 
+    #guides(color = FALSE) +
+    guides(alpha = FALSE) + 
+    theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
+   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(0, 1.0e-9))+mytheme
 
+gA <- ggplotGrob(p1)
+gB <- ggplotGrob(p2)
+grid::grid.newpage()
+grid::grid.draw(rbind(gA, gB))
 ```

### 2020-05-20 Rcode/PlotsForPaper.R

@@ -206,7 +206,7 @@ allData<-allData%>%filter(UnitCode %in% unique(predDat$UnitCode))
   
   ## 
   
-  p1<-predDat%>%filter(UnitCode == "OLYM")%>%
+  p1<-predDat%>%filter(UnitCode == "GLAC")%>%
     ggplot(.,aes(x=date))+
     geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
     geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
@@ -218,7 +218,7 @@ allData<-allData%>%filter(UnitCode %in% unique(predDat$UnitCode))
     scale_color_identity(name = "",
                          breaks = c("black", "#1f78b4"),
                          labels = c("Observed", "Median Estimate Minimum Smoke"),
-                         guide = "legend")+ggtitle("Rocky Mountain National Park")+theme(legend.position="top")+
+                         guide = "legend")+ggtitle("Yellowstone National Park")+theme(legend.position="top")+
     theme(axis.title.x=element_blank(),
           axis.text.x=element_blank(),
           axis.ticks.x=element_blank(),
@@ -227,7 +227,7 @@ allData<-allData%>%filter(UnitCode %in% unique(predDat$UnitCode))
 
   
   
-p2<-predDat%>%filter(UnitCode == "OLYM")%>%group_by(Year)%>%
+p2<-predDat%>%filter(UnitCode == "GLAC")%>%group_by(Year)%>%
     summarise(meansmoke=mean(Smoke))%>%
     ggplot(.,aes(x=Year,y=meansmoke))+
     geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
@@ -238,7 +238,7 @@ p2<-predDat%>%filter(UnitCode == "OLYM")%>%group_by(Year)%>%
     #guides(color = FALSE) +
     guides(alpha = FALSE) + 
     theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
-   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(0, 9e-10))+mytheme
+   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0))+mytheme
   
 
 gA <- ggplotGrob(p1)

### 2020-05-20 Markdown Documents/Manuscript.Rmd
Working on intro

@@
 
-Paragraphs:
+Points:
 1. Public lands are important
@@
 9. Given the trends we observe, we project out into the future what the effects might be over the next 50 years.
+
+Twenty eight percent of the land area of the United States is designated as public land [@vincent_federal_2020].
+These public lands are the critical backdrop for the majority wildlife conservation and outdoor recreation in the country [@romagosa_inside_2015; @leemans_millennium_2003; @dudley_national_2011].
+In a 2016 study, 93% of respondents indicated that they believe public open spaces and recreational areas should be protected, regardless of whether or not they visited them personally [@haefele_total_2016].
+The crown jewel of the United States Public lands system is the National Park Service.
+The national park system alone received approximately 318 million recreational visitors and generated over $40 billion in economic output to local economies in 2018 [@thomas_2018_2018].
+
+National parks undeniably represent a critical infrastructure for supporting ecological integrity and outdoor recreation in the United States, but have recently drawn public attention for considerable underfunding [@simmonds_crisis_2018;@duncan_opinion_2016].
+Most noteably for the $12 billion repair backlog for basic maintence and resource protection [@nps_nps_2018].
+In additiion to necessary repairs, the National Park Service reports a 14% reduction in staff funding over the last 5 years [@doi_fiscal_2019].
+While the majority of National Park Service funding comes from congressional appropriations, recent bugetary restrictions have put pressure on parks to gather an increasing amount of their budget from recreational visitor fees such as admissions and camping.
+The income from these fees has been steadily increasing over the last 10 years as parks have seen visitation increases while simultanously increasing visitor fees [@gao_national_2016].
 
@@
     ggplot(.,aes(x=Year,y=meansmoke))+
-    geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
+    geom_segment(aes(xend = Year, yend = 2.5e-10,color =(meansmoke)),size=1.5) +
     geom_point(aes(color =(meansmoke)),size=5)+
@@
     theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
-   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(0, 1.25e-9))+mytheme
+   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(2.5e-10, 1.25e-9))+mytheme
 
@@
     ggplot(.,aes(x=Year,y=meansmoke))+
-    geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
+    geom_segment(aes(xend = Year, yend = 1e-10,color =(meansmoke)),size=1.5) +
     geom_point(aes(color =(meansmoke)),size=5)+
@@
     theme_classic()+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
-   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(0, 1.0e-9))+mytheme
+   scale_y_continuous(name="Smoke (PM 2.5)",expand = c(0, 0), limits = c(1e-10, 9e-10))+mytheme
 

### 2020-05-20 Rcode/PlotsForPaper.R

@@ -206,7 +206,7 @@ allData<-allData%>%filter(UnitCode %in% unique(predDat$UnitCode))
   
   ## 
   
-  p1<-predDat%>%filter(UnitCode == "GLAC")%>%
+  p1<-predDat%>%filter(UnitCode == "OLYM")%>%
     ggplot(.,aes(x=date))+
     geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
     geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
@@ -218,7 +218,7 @@ allData<-allData%>%filter(UnitCode %in% unique(predDat$UnitCode))
     scale_color_identity(name = "",
                          breaks = c("black", "#1f78b4"),
                          labels = c("Observed", "Median Estimate Minimum Smoke"),
-                         guide = "legend")+ggtitle("Yellowstone National Park")+theme(legend.position="top")+
+                         guide = "legend")+ggtitle("Glacier National Park")+theme(legend.position="top")+
     theme(axis.title.x=element_blank(),
           axis.text.x=element_blank(),
           axis.ticks.x=element_blank(),
@@ -227,10 +227,10 @@ allData<-allData%>%filter(UnitCode %in% unique(predDat$UnitCode))
 
   
   
-p2<-predDat%>%filter(UnitCode == "GLAC")%>%group_by(Year)%>%
+p2<-predDat%>%filter(UnitCode == "OLYM")%>%group_by(Year)%>%
     summarise(meansmoke=mean(Smoke))%>%
     ggplot(.,aes(x=Year,y=meansmoke))+
-    geom_segment(aes(xend = Year, yend = 0,color =(meansmoke)),size=1.5) +
+    geom_segment(aes(xend = Year, yend = 1e-10,color =(meansmoke)),size=1.5) +
     geom_point(aes(color =(meansmoke)),size=5)+
     scale_color_gradient2(low ="#2b83ba" , mid = "#ffffbf", high = "#d7191c",name= "Difference from trend",breaks=c(0.050,-0.035),
                           labels=c("Above","Below")
@@ -238,7 +238,7 @@ p2<-predDat%>%filter(UnitCode == "GLAC")%>%group_by(Year)%>%
     #guides(color = FALSE) +
[…]

### 2020-05-20 Markdown Documents/Manuscript.Rmd
Another paragraph down

@@
 
-National parks undeniably represent a critical infrastructure for supporting ecological integrity and outdoor recreation in the United States, but have recently drawn public attention for considerable underfunding [@simmonds_crisis_2018;@duncan_opinion_2016].
-Most noteably for the $12 billion repair backlog for basic maintence and resource protection [@nps_nps_2018].
-In additiion to necessary repairs, the National Park Service reports a 14% reduction in staff funding over the last 5 years [@doi_fiscal_2019].
-While the majority of National Park Service funding comes from congressional appropriations, recent bugetary restrictions have put pressure on parks to gather an increasing amount of their budget from recreational visitor fees such as admissions and camping.
-The income from these fees has been steadily increasing over the last 10 years as parks have seen visitation increases while simultanously increasing visitor fees [@gao_national_2016].
+National parks undeniably represent a critical infrastructure for supporting ecological integrity and outdoor recreation in the United States, but have recently drawn public attention for considerable under funding [@simmonds_crisis_2018;@duncan_opinion_2016].
+Most notably for the $12 billion repair backlog for basic maintenance and resource protection [@nps_nps_2018].
+In addition to necessary repairs, the National Park Service reports a 14% reduction in staff funding over the last 5 years [@doi_fiscal_2019].
+While the majority of National Park Service funding comes from congressional appropriations, recent budgetary restrictions have put pressure on parks to gather an increasing amount of their budget from recreational visitor fees such as admissions and camping.
+The income from these fees has been steadily increasing over the last 10 years as parks have seen visitation increases while simultaneously increasing visitor fees [@gao_national_2016].
+
+As national parks become increasingly dependent on visitor fees in order to provide the suite of social and ecological services communities depend on, it is of paralleled importance to understand the drivers of fluctuation in recreational visits.
+Research has shown that economic considerations (e.g.
+park fees) are considerably less influential on park visitation than are recreational opportunities [@neuvonen_visits_2010;@stevens_declining_2014].
+Most factors which influence the recreational accessibility of a given park are either fixed (natural resources within the park area) or within the control of park management (permitting, trail building, etc.).
+Wildfire smoke however, shows considerable yearly variation and is largely produced by adjacent public lands and therefore out of the jurisdiction of park management [@dombeck_wildfire_2004;@collins_regional_2006].
 

### 2020-05-20 Markdown Documents/Manuscript.Rmd
Rough draft of intro done

@@
 
-This is how you add a reference (e.g., [@fisichelli_protected_2015; @fann_ch_2016])
-
-Points:
-1. Public lands are important
-2. NPs are specifically high value
-3. NPs are under funded. This under funding prevents them from doing those important jobs
-4. NPs get a lot of their funding from visitation. Vis is influenced by a lot of things. One of these things that's particularly out of their control is smoke from adjacent forested areas.
-5. Smoke is bad for recreation and should decrease vis.
-6. Smoke events have been increasing over the last 50 years. Particularly, these increases have been in the west.
-7. These smoke events are expected to worsen via climate change over the next 50 years.
-8. Logically, smoke should influence visitation to parks. This decrease should affect income. We seek to assess this effect in a systematic way and backcast what the monetary effects might have been.
-9. Given the trends we observe, we project out into the future what the effects might be over the next 50 years.
-
 Twenty eight percent of the land area of the United States is designated as public land [@vincent_federal_2020].
@@
 In a 2016 study, 93% of respondents indicated that they believe public open spaces and recreational areas should be protected, regardless of whether or not they visited them personally [@haefele_total_2016].
-The crown jewel of the United States Public lands system is the National Park Service.
+The crown jewel of the United States public lands system is the National Park Service.
 The national park system alone received approximately 318 million recreational visitors and generated over $40 billion in economic output to local economies in 2018 [@thomas_2018_2018].
@@
 While the majority of National Park Service funding comes from congressional appropriations, recent budgetary restrictions have put pressure on parks to gather an increasing amount of their budget from recreational visitor fees such as admissions and camping.
-The income from these fees has been steadily increasing over the last 10 years as parks have seen visitation increases while simultaneously increasing visitor fees [@gao_national_2016].
+The income from these fees has steadily increased over the last 10 years as park visitation and visitor fees have both surged [@gao_national_2016].
 
-As national parks become increasingly dependent on visitor fees in order to provide the suite of social and ecological services communities depend on, it is of paralleled importance to understand the drivers of fluctuation in recreational visits.
+As national parks become increasingly dependent on visitor fees in order to provide the suite of social and ecological services communities depend on, it is of paralleled importance to understand what drives fluctuation in recreational visits.
 Research has shown that economic considerations (e.g.
@@
 Wildfire smoke however, shows considerable yearly variation and is largely produced by adjacent public lands and therefore out of the jurisdiction of park management [@dombeck_wildfire_2004;@collins_regional_2006].
+
+Wildfire smoke is a mixture of many gasses aerated by biomass combustion across a landscape [@naeher_woodsmoke_2007].
+While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 μm in diameter (PM2.5) is particularly harmful to human health, causing between 260,000 - 600,000 global deaths annually [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007].
+Health concerns regarding wildfire smoke, specifically PM2.5, regurlarly drive calls for individuals to reduce outdoor recreation behavior [@oneill_summary_2013].
+
+Wildfires and the smoke they produce have been increasing in the United States over the last 50 years, specifically in the American West and Inter-Mountain regions [@kitzberger_contingent_2007;@westerling_increasing_2016].
+These regions are also projected to experience increased wildfire frequency and severity over the next 50 years as a result of climate change [@harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
+Anecdotally, it is known that wildfire smoke has impacted the recreation behavior of individuals in these areas and decreased visitation revenue to the national parks in both regions.
+This assumption however has not yet been demonstrated empirically.
+
+The primary goal of this project is to assess how visitation to national parks in the Inter-Mountain and Pacific regions has systematically varied with surface level concentrations of PM2.5 resulting from wildfire smoke from 1980 - 2018.
+Given the variability in recreational fees between parks and over time, the secondary goal of this project is to produce order of magnitude estimates of the financial losses that wildfire smoke has caused  park units during the same time period.
+The intention of this research is not to perform an economic analysis of smoke effects on each park specifically, but rather to provide an approximation of how wildfire smoke affects population level recreational decision making and what the downstream economic impacts have been historically and may be in the future.
 

### 2020-05-21 Markdown Documents/Manuscript.Rmd
small text change

@@
 The primary goal of this project is to assess how visitation to national parks in the Inter-Mountain and Pacific regions has systematically varied with surface level concentrations of PM2.5 resulting from wildfire smoke from 1980 - 2018.
-Given the variability in recreational fees between parks and over time, the secondary goal of this project is to produce order of magnitude estimates of the financial losses that wildfire smoke has caused  park units during the same time period.
+Given the variability in recreational fees between parks and over time, the secondary goal of this project is to produce order of magnitude estimates of the financial losses that wildfire smoke has caused  park units during the same period.
 The intention of this research is not to perform an economic analysis of smoke effects on each park specifically, but rather to provide an approximation of how wildfire smoke affects population level recreational decision making and what the downstream economic impacts have been historically and may be in the future.

### 2020-05-21 Markdown Documents/Manuscript.Rmd
added a sick dumbbell fig

@@
 
-We can apply these visitation estimates to the fee data/predictions also. This gives us the following table.
 ```{r echo=FALSE, cache=TRUE}
@@
 summaryFinancialLoss$`Percent Income Loss (1980-2018)`<-paste0(summaryFinancialLoss$`Percent Income Loss (1980-2018)`,'%')
+```
+
+```{r echo=FALSE, cache=TRUE, warning=FALSE, message=FALSE, fig.height = 6, fig.width = 12}
+dat%>%filter(UnitCode %in% fitsig$Par)%>%group_by(ParkName)%>%
+  summarise(totrec=sum(RecreationVisits),totpred=sum(PredNoSmoke50CI))%>%ungroup()%>%
+  mutate(diff=totpred-totrec)%>%
+ggplot(., 
+       aes(y = reorder(ParkName,diff),
+           x = totrec,
+           xend = totpred)) +  
+  ggalt::geom_dumbbell(size = 1.2,
+                       size_x = 5, 
+                       size_xend = 5,
+                       colour = "darkgrey", 
+                       colour_x = "black", 
+                       colour_xend = "#1f78b4",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey") +
+  scale_x_continuous( breaks=seq(10000000,90000000,20000000),labels = scales::comma) + 
+  labs(title = "Total Lost Visitation",
+       subtitle = "1980 to 2018",
+       x = "Total Visits",
+       y = "")+theme_classic()+mytheme
+
+```
+
+We can apply these visitation estimates to the fee data/predictions also. This gives us the following table.
+```{r echo=FALSE, cache=TRUE}
+
 formattable::formattable(summaryFinancialLoss)
 ```
+
 ## 3.3. Model Validation

### 2020-05-21 Markdown Documents/Manuscript.Rmd
cleaned up dumbbell plot

@@
                        colour_xend = "#1f78b4",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey") +
+   annotate("text", x = 60000000, y = "Glacier NP", label = "Observed Visitation",vjust=-2,hjust=1)+
+  annotate("text", x = 80000000, y = "Glacier NP", label = "Predicted Visitation No Smoke",vjust=-2)+
   scale_x_continuous( breaks=seq(10000000,90000000,20000000),labels = scales::comma) + 
@@
        x = "Total Visits",
-       y = "")+theme_classic()+mytheme
+       y = "")+theme_classic()+mytheme+theme(plot.subtitle = element_text(face = "bold"))
 

### 2020-05-21 Markdown Documents/Manuscript.Rmd
small plotting changes

--- 2020-05-21 Markdown Documents/Manuscript.Rmd
@@
                        colour_xend = "#1f78b4",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey") +
+   annotate("text", x = 60000000, y = "Glacier NP", label = "Observed Visitation",vjust=-1,hjust=1)+
+  annotate("text", x = 80000000, y = "Glacier NP", label = "Predicted Visitation No Smoke",vjust=-1)+
   scale_x_continuous( breaks=seq(10000000,90000000,20000000),labels = scales::comma) + 

### 2020-05-21 Rcode/HESBoelWorkingScript.R

@@ -130,3 +130,32 @@ formattable::formattable(summaryFinancialLoss)
 sum(dat$RecreationVisits)/sum(dat$PredNoSmoke50CI)
 sum(summaryFinancialLoss$IncomeDif)  
   
+
+
+# create dumbbell plot
+
+dat%>%filter(UnitCode %in% fitsig$Par)%>%group_by(ParkName)%>%
+  summarise(totrec=sum(RecreationVisits),totpred=sum(PredNoSmoke50CI))%>%ungroup()%>%
+  mutate(diff=totpred-totrec)%>%
+ggplot(., 
+       aes(y = reorder(ParkName,diff),
+           x = totrec,
+           xend = totpred)) +  
+  ggalt::geom_dumbbell(size = 1.2,
+                       size_x = 5, 
+                       size_xend = 5,
+                       colour = "darkgrey", 
+                       colour_x = "black", 
+                       colour_xend = "#1f78b4",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey") +
+   annotate("text", x = 60000000, y = "Glacier NP", label = "Observed Visitation",vjust=-2,hjust=1)+
+  annotate("text", x = 80000000, y = "Glacier NP", label = "Predicted Visitation No Smoke",vjust=-2)+
+  scale_x_continuous( breaks=seq(10000000,90000000,20000000),labels = scales::comma) + 
+  labs(title = "Total Lost Visitation",
+       subtitle = "1980 to 2018",
+       x = "Total Visits",
+       y = "")+theme_classic()+mytheme+theme(plot.subtitle = element_text(face = "bold"))
+
+
+
+
+

### 2020-05-21 Rcode/PlotsForPaper.R

@@ -186,7 +186,7 @@ allData<-allData%>%filter(UnitCode %in% unique(predDat$UnitCode))
   
   ##Lolipops  ...
   predDat%>%group_by(UnitCode)%>%
-    summarise(totrec=sum(RecreationVisits),totpred=sum(No_Smoke_Mean_est))%>%
+    summarise(totrec=sum(RecreationVisits),totpred=sum(PredNoSmoke50CI))%>%
   ggplot(.) +
     geom_segment( aes(x=UnitCode, xend=UnitCode, y=totrec, yend=totpred), color="grey") +
     geom_point( aes(x=UnitCode, y=totrec), color=rgb(0.2,0.7,0.1,0.5), size=3 ) +

### 2020-05-21 Markdown Documents/Manuscript.Rmd
Small plot anotation change

--- 2020-05-21 Markdown Documents/Manuscript.Rmd
@@
                        colour_xend = "#1f78b4",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey") +
+   annotate("text", x = 60000000, y = "Yosemite NP", label = "Observed visitation",vjust=-1,hjust=1)+
+  annotate("text", x = 73000000, y = "Yosemite NP", label = "Predicted visitation no smoke",vjust=-1)+
   scale_x_continuous( breaks=seq(10000000,90000000,20000000),labels = scales::comma) + 

### 2020-05-21 Markdown Documents/Manuscript.Rmd
small annotation changes

--- 2020-05-21 Markdown Documents/Manuscript.Rmd
@@
                        colour_xend = "#1f78b4",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey") +
-   annotate("text", x = 60000000, y = "Glacier NP", label = "Observed Visitation",vjust=-1,hjust=1)+
-  annotate("text", x = 80000000, y = "Glacier NP", label = "Predicted Visitation No Smoke",vjust=-1)+
+   annotate("text", x = 60000000, y = "Yosemite NP", label = "Observed visitation",vjust=-1,hjust=1,fontface=2)+
+  annotate("text", x = 75000000, y = "Yosemite NP", label = "Predicted visitation no smoke",vjust=-1,color="#1f78b4",fontface=2)+
   scale_x_continuous( breaks=seq(10000000,90000000,20000000),labels = scales::comma) + 

### 2020-05-21 Markdown Documents/Manuscript.Rmd
small plot chages

@@
   ggplot2::scale_y_discrete(name ="Park",labels=rev(PNames),limits=rev(limits))+
-  mytheme+scale_x_continuous(name="Parameter Value", labels = scales::comma)
+  mytheme+scale_x_continuous(name="Parameter Value", labels = scales::comma)+
+  theme(panel.grid.major.y = element_line(color = "lightgrey", linetype="dotted")) 
 ```
@@
     geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
-    geom_line(aes(y=RecreationVisits,color="grey"),size=1.25)+
+    geom_line(aes(y=RecreationVisits,color="darkgrey"),size=1.25)+
     geom_point(aes(y=RecreationVisits,color="black"),size=2)+
@@
     scale_color_identity(name = "",
-                         breaks = c("black", "#1f78b4"),
+                         breaks = c("darkgrey", "#1f78b4"),
                          labels = c("Observed", "Median Estimate Minimum Smoke"),

### 2020-05-21 Markdown Documents/Manuscript.Rmd
edited abstract

@@
 thanks: If we want to thank anyone? Maybe Moji?.
-abstract: National Parks are important.
-National Parks are low on funding which affects their ability to do their job.
-A key way they get this funding is through visitation.
-Wildfire smoke affects visitation greatly and is out of ther control.
-Wildfires are expected to increase in the Pacific West and Intermountain regions.
-Over the last 38 years, we estimate that wildfire smoke has led to a 12% overall visitation decrease to National Parks in these regions.
+abstract: The United States national park system serves as a global model for how natural resource protection can provide numerous social and ecological services while generating econimic returns on investment.
+In recent years parks have increasingly relied on visitation fees to fund consevation projects as well as basic infrastructure.
+In contrast with congressional funding, visitation is subject to fluctuation given climatic and environmental variability.
+While many environmental variables are relatively predictable and stable over time, wildfire smoke events have increased drmatically via climate change, particurlarly in the American West.
+In this project, we assess how recreational visitation to national parks has systematically varied with wildfire smoke.
+Over the last 38 years, we estimate that wildfire smoke has led to a 12% overall visitation decrease to National Parks in the Pacific West and Inter-Mountain Regions.
 When accounting for visitation fees, this loss may have cost the National Park Service as much as $580 million in total.
-Wildfires and the smoke it produces is supposed to increase over the next 50 years.
-Assuming a linear trend in smoke, visitation overall, and park fees, we expect to see significant financial losses to the National Park Service.
+Wildfire frequency and severity are expected to increase considerably in these regions over the next 50 years.
+Assuming current trends in smoke, park visitation, and park fees continue, we expect to see significant financial losses to the National Park Service via wildfire smoke in the coming years.
 Our recomendation is that....NPS work with the forest service to prevent fires in areas close to parks??

### 2020-05-26 Markdown Documents/Manuscript.Rmd
small changes

@@
 While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 μm in diameter (PM2.5) is particularly harmful to human health, causing between 260,000 - 600,000 global deaths annually [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007].
-Health concerns regarding wildfire smoke, specifically PM2.5, regurlarly drive calls for individuals to reduce outdoor recreation behavior [@oneill_summary_2013].
+Health concerns regarding wildfire smoke, specifically PM2.5, regularly drive calls for individuals to reduce outdoor recreation behavior [@oneill_summary_2013].
 

### 2020-05-26 Markdown Documents/Manuscript.Rmd
moved template

@@
     latex_engine: pdflatex
-    template: /Users/Matthewclark989/Documents/LatexPractice/templatedocs/svm_ms_template.tex
+    template: /Users/Matthewclark989/Documents/SmokeProject/svm_ms_template.tex
 fontfamily: mathpazo

### 2020-05-26 Markdown Documents/Manuscript.Rmd
Update Manuscript.Rmd

@@
 thanks: If we want to thank anyone? Maybe Moji?.
-abstract: The United States national park system serves as a global model for how natural resource protection can provide numerous social and ecological services while generating econimic returns on investment.
+abstract: The United States national park system serves as a global model for how natural resource protection can provide numerous social and ecological services while generating economic returns on investment.
 In recent years parks have increasingly relied on visitation fees to fund consevation projects as well as basic infrastructure.

### 2020-05-27 Markdown Documents/Manuscript.Rmd
small spacing change

@@
 ```
+
+```{r, include=FALSE}
+options(tinytex.verbose = TRUE)
+```
 ## 2.1. Data Collection
@@
 
+<!-- This is how you comment -->
+
 To account for the linear trends of visitation, we first calculated the trends, then found the residuals of each datapoint from the trend line.
@@
 pars+ggplot2::ggtitle("Posterior medians with 50% and 90% intervals")+
-  ggplot2::scale_y_discrete(name ="Park",labels=rev(PNames),limits=rev(limits))+
+  ggplot2::scale_y_discrete(name ="Park",labels=rev(PNames),limits=rev(limits))+theme_classic()+
   mytheme+scale_x_continuous(name="Parameter Value", labels = scales::comma)+

### 2020-05-27 Markdown Documents/Manuscript.Rmd
added 2 paragraphs to the intro

@@
 Assuming current trends in smoke, park visitation, and park fees continue, we expect to see significant financial losses to the National Park Service via wildfire smoke in the coming years.
-Our recomendation is that....NPS work with the forest service to prevent fires in areas close to parks??
-Do they already do that??
 ---
@@
 park fees) are considerably less influential on park visitation than are recreational opportunities [@neuvonen_visits_2010;@stevens_declining_2014].
-Most factors which influence the recreational accessibility of a given park are either fixed (natural resources within the park area) or within the control of park management (permitting, trail building, etc.).
-Wildfire smoke however, shows considerable yearly variation and is largely produced by adjacent public lands and therefore out of the jurisdiction of park management [@dombeck_wildfire_2004;@collins_regional_2006].
+Most factors which influence the recreational accessibility of a given park are either fixed (e.g.
+natural resources within the park area) or within the control of park management (permitting, trail building, etc.).
+Wildfires however, largely result from external forces and have been shown to negatively affect individual's willingness to engage in recreational travel [@thapa_wildfires_2013].
+Furthermore, a study of five national parks in Utah showed visitation losses from wildfire events even at a distance of up to 320 kilometers from the park boundary [@kim_wildfire_2019].
+Given the distance with which wildfires have been shown to impact park visitation, it can be assumed that individuals modify their recreational behavior not only as a result of wildfires themselves but also the smoke they produce.
 
@@
 While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 μm in diameter (PM2.5) is particularly harmful to human health, causing between 260,000 - 600,000 global deaths annually [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007].
-Health concerns regarding wildfire smoke, specifically PM2.5, regularly drive calls for individuals to reduce outdoor recreation behavior [@oneill_summary_2013].
+A 2012 study estimated that wildfire smoke in Los Angeles County resulted in public health costs of approximately $84 per exposed person each day [@richardson_hidden_2012].
+Health concerns regarding wildfire smoke, specifically PM2.5, regularly drive calls for individuals to reduce outdoor recreational behavior [@oneill_summary_2013].
+
+Recent literature reviews on recreational behavior have found a dearth of research both describing how individual's are affected by and respond to low air quality, as well as research which gives predictions for how climate change will affect the recreation landscape in specific areas [@otoole_climate_2019;@zajchowski_air_2019-1].
+Additionally, a survey of federal land managers indicated that decision makers in areas such as national parks feel that they have a shortage of information describing how air quality affects the recreational behavior of their visitors [@zajchowski_air_2019].
+It is of particular importance to understand how individuals respond to wildfire smoke because of the highly variable nature of wildfires, their relationship to climate change, and the great distances that smoke can travel during and after wildfire events [@dombeck_wildfire_2004; @collins_regional_2006].
 
@@
 These regions are also projected to experience increased wildfire frequency and severity over the next 50 years as a result of climate change [@harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
-Anecdotally, it is known that wildfire smoke has impacted the recreation behavior of individuals in these areas and decreased visitation revenue to the national parks in both regions.
-This assumption however has not yet been demonstrated empirically.
 
-The primary goal of this project is to assess how visitation to national parks in the Inter-Mountain and Pacific regions has systematically varied with surface level concentrations of PM2.5 resulting from wildfire smoke from 1980 - 2018.
+The primary goal of this project is to assess how visitation to national parks in the Inter-Mountain and Pacific West regions has systematically varied with surface level concentrations of PM2.5 resulting from wildfire smoke from 1980 - 2018.
 Given the variability in recreational fees between parks and over time, the secondary goal of this project is to produce order of magnitude estimates of the financial losses that wildfire smoke has caused park units during the same period.
-The intention of this research is not to perform an economic analysis of smoke effects on each park specifically, but rather to provide an approximation of how wildfire smoke affects population level recreational decision making and what the downstream economic impacts have been historically and may be in the future.
+The intention of this research is not to perform an in-depth economic analysis of smoke effects on each park specifically, but rather to provide an approximation of how wildfire smoke affects population level recreational decision making and what the downstream economic impacts have been historically and what the implications are for the future.
+Overall, this research aims to better equip protected area managers with the information necessary to best manage recreational resources in the face of increasing wildfire events and climatic variability generally.
 

### 2020-05-27 Markdown Documents/Manuscript.Rmd
changed template file paths.

@@
     latex_engine: pdflatex
-    template: /Users/Matthewclark989/Documents/SmokeProject/svm_ms_template.tex
+    template: ../svm_ms_template.tex
 fontfamily: mathpazo
@@
 keywords: pandoc, r markdown, knitr
-bibliography: /Users/Matthewclark989/Documents/SmokeProject/References.bib
+bibliography: ../References.bib
 thanks: If we want to thank anyone? Maybe Moji?.
@@
 The national park system alone received approximately 318 million recreational visitors and generated over $40 billion in economic output to local economies in 2018 [@thomas_2018_2018].
+
+<!-- regular html comment -->
 
@@
 ```
-
-<!-- This is how you comment -->
 

### 2020-05-28 Markdown Documents/Manuscript.Rmd
added difference columns and a barbell plot to lost revenue

@@
 
-```{r echo=FALSE, cache=TRUE, warning=FALSE, message=FALSE, fig.height = 6, fig.width = 12}
+```{r echo=FALSE, cache=TRUE, warning=FALSE, message=FALSE, fig.height = 8, fig.width = 13}
 dat%>%filter(UnitCode %in% fitsig$Par)%>%group_by(ParkName)%>%
@@
                        colour_xend = "#1f78b4",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey") +
+  geom_rect( aes(xmin=91000000, xmax=99000000, ymin=-Inf, ymax=Inf), fill="grey") +
+  geom_text( aes(label=paste0("-",round((diff/totrec)*100,digits=1), "%"), y=ParkName, x=95000000), fontface=2, size=4)+
+  annotate("text", x = 95000000, y = "Glacier NP", label = "Difference",vjust=-1.5,fontface=2)+
   annotate("text", x = 60000000, y = "Yosemite NP", label = "Observed visitation",vjust=-1,hjust=1,fontface=2)+
@@
 
-We can apply these visitation estimates to the fee data/predictions also. This gives us the following table.
-```{r echo=FALSE, cache=TRUE}
+We can apply these visitation estimates to the fee data/predictions also.
+```{r echo=FALSE, cache=TRUE, warning=FALSE, message=FALSE, fig.height = 8, fig.width = 13}
+financialData<-merge(dat[,c(2,5)],financialData,by="UnitCode")
+financialData%>%filter(UnitCode %in% fitsig$Par)%>%group_by(ParkName)%>%
+  summarise(sumincome=sum(incomeObserved),sumincomePred=sum(incomePred))%>%ungroup()%>%
+  mutate(diff=sumincomePred-sumincome)%>%
+  ggplot(., 
+         aes(y = reorder(ParkName,diff),
+             x = sumincome,
+             xend = sumincomePred)) +  
+  ggalt::geom_dumbbell(size = 1.2,
+                       size_x = 5, 
+                       size_xend = 5,
+                       colour = "darkgrey", 
+                       colour_x = "black", 
+                       colour_xend = "#b2df8a",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey")+
+  geom_rect( aes(xmin=100500000000, xmax=109500000000, ymin=-Inf, ymax=Inf), fill="grey") +
+  geom_text( aes(label=paste0("-",round((diff/sumincome)*100,digits=1), "%"), y=ParkName, x=105000000000), fontface=2, size=4)+
+  annotate("text", x = 105000000000, y = "Glacier NP", label = "Difference",vjust=-1.5,fontface=2)+
+  scale_x_continuous( labels = scales::dollar_format(prefix = "$")) +
+  annotate("text", x = 62000000000, y = "Yosemite NP", label = "Observed fee revenue",vjust=-1,hjust=1,fontface=2)+
+  annotate("text", x = 80000000000, y = "Yosemite NP", label = "Predicted fee revenue no smoke",vjust=-1,color="#b2df8a",fontface=2)+
+  labs(title = "Total Lost Revenue",
+       subtitle = "1980 to 2018",
+       x = "Total Visitor Revenue 1980 to 2018",
+       y = "")+theme_classic()+mytheme+theme(plot.subtitle = element_text(face = "bold"))
 
-formattable::formattable(summaryFinancialLoss)
 ```

### 2020-05-28 Rcode/HESBoelWorkingScript.R
dumbell plotz

@@ -53,8 +53,8 @@ dat$PredNoSmoke75CI<-dat$trendvis+dat$PredNoSmoke75CI
 dat$PredNoSmoke25CI<-dat$trendvis+dat$PredNoSmoke25CI
 
 
-write.csv(dat,file="Data/FullDataWithMinSmokePredictions.csv")
-
+#write.csv(dat,file="Data/FullDataWithMinSmokePredictions.csv")
+dat<-read.csv("Data/FullDataWithMinSmokePredictions.csv")
 mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                 plot.title = element_text( size=18, color="black",face="bold"),
                 axis.title.x = element_text( size=18),
@@ -100,7 +100,7 @@ FEdat<-read.csv("~/SmokeProject/Data/ParkFees.csv")
 
 #get the visitation observed and predicted yearly totals (same scale as fee data)
 VEstDat<-dat%>%group_by(UnitCode,Year)%>%
-  summarise(RecVisits=sum(RecreationVisits),PredVisits = sum(PredNoSmoke))%>%
+  summarise(RecVisits=sum(RecreationVisits),PredVisits = sum(PredNoSmoke50CI))%>%
   filter(UnitCode %in% fitsig$Par)
 
 #only include significant parks and years we predicted for
@@ -120,7 +120,7 @@ summaryFinancialLoss<-financialData %>% group_by(UnitCode)%>%summarise(totrec=su
 summaryFinancialLoss$ParkName<-unique(dat[dat$UnitCode %in% summaryFinancialLoss$UnitCode,]$ParkName)
 
 summaryFinancialLoss<-summaryFinancialLoss[,c(6,4,5)]
-summaryFinancialLoss$IncomeDif<-currency(summaryFinancialLoss$IncomeDif)
+summaryFinancialLoss$IncomeDif<-formattable::currency(summaryFinancialLoss$IncomeDif)
 names(summaryFinancialLoss)<-(c("National Park","Income Loss Estimate (1980-2018)", "Percent Income Loss (1980-2018)"))
 summaryFinancialLoss$`Percent Income Loss (1980-2018)`<-paste0(summaryFinancialLoss$`Percent Income Loss (1980-2018)`,'%')
 formattable::formattable(summaryFinancialLoss)
@@ -132,23 +132,26 @@ sum(summaryFinancialLoss$IncomeDif)
   
 
 
[…]

### 2020-05-28 Rcode/PlotsForPaper.R

@@ -245,3 +245,37 @@ gA <- ggplotGrob(p1)
 gB <- ggplotGrob(p2)
 grid::grid.newpage()
 grid::grid.draw(rbind(gA, gB))
+
+
+predDat%>%filter(UnitCode == "GLAC")%>%
+  ggplot(.,aes(x=date))+
+  geom_ribbon(aes(ymin=PredNoSmoke5CI, ymax=PredNoSmoke95CI),fill="#1f78b4",alpha=0.2)+
+  geom_ribbon(aes(ymin=PredNoSmoke25CI, ymax=PredNoSmoke75CI),fill="#1f78b4",alpha=0.5)+
+  geom_line(aes(y=RecreationVisits,color="darkgrey"),size=1.25)+
+  geom_point(aes(y=RecreationVisits,color="black"),size=2)+
+  geom_line(aes(y=PredNoSmoke50CI,color="#1f78b4"),linetype="longdash",size=1.25)+theme_classic()+
+  mytheme + scale_y_continuous(name="Recreational Visits", labels = scales::comma)+
+  scale_x_continuous(name="Date",breaks=c(1980,1990,2000,2010,2018))+
+  scale_color_identity(name = "",
+                       breaks = c("black", "#1f78b4"),
+                       labels = c("Observed", "Median Estimate Minimum Smoke"),
+                       guide = "legend")+ggtitle("Glacier National Park")+theme(legend.position="top")+
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.line.x = element_blank())
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2020-05-28 Markdown Documents/Manuscript.Rmd
Removed pdf template. Not dealing with it for now

@@
     df_print: paged
-  word_document: default
   pdf_document:
@@
     latex_engine: pdflatex
-    template: ../svm_ms_template.tex
+  word_document: default
 fontfamily: mathpazo
@@
 geometry: margin=1in
-keywords: pandoc, r markdown, knitr
+keywords: Wildfire, Smoke, Recreation, Park Science
 bibliography: ../References.bib
@@
 The national park system alone received approximately 318 million recreational visitors and generated over $40 billion in economic output to local economies in 2018 [@thomas_2018_2018].
-
-<!-- regular html comment -->
 
@@
 Our predicted fees correlate well with our observed fees that we have
-```{r echo = FALSE, cache=TRUE}
+```{r echo = FALSE, cache=TRUE,message=FALSE,warning=FALSE}
 xx<-na.omit(FEdat)
@@
                        colour_x = "black", 
-                       colour_xend = "#b2df8a",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey")+
+                       colour_xend = "#41ab5d",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey")+
   geom_rect( aes(xmin=100500000000, xmax=109500000000, ymin=-Inf, ymax=Inf), fill="grey") +
@@
   annotate("text", x = 62000000000, y = "Yosemite NP", label = "Observed fee revenue",vjust=-1,hjust=1,fontface=2)+
-  annotate("text", x = 80000000000, y = "Yosemite NP", label = "Predicted fee revenue no smoke",vjust=-1,color="#b2df8a",fontface=2)+
+  annotate("text", x = 80000000000, y = "Yosemite NP", label = "Predicted fee revenue no smoke",vjust=-1,color="#41ab5d",fontface=2)+
   labs(title = "Total Lost Revenue",
@@
 Discussion here
+
+# 5. Supplemental

### 2020-05-28 Markdown Documents/Manuscript.Rmd
moved a bunch of stuff to supplemental

@@
 
-We then used this to predict what the fees would have been for every year we didn't have
-```{r cache=TRUE, include=FALSE}
-fillData<-FEdat[,1:2]
-
-FEdat$PredictedFee<-apply(posterior_predict(feefit,fillData),2,mean) #assuming this is for a new column
-
-```
-
-Our predicted fees correlate well with our observed fees that we have
-```{r echo = FALSE, cache=TRUE,message=FALSE,warning=FALSE}
-xx<-na.omit(FEdat)
- mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-                  plot.title = element_text( size=18, color="black",face="bold"),
-                  axis.title.x = element_text( size=18),
-                  axis.title.y = element_text( size=18),
-                  axis.text=(element_text(color="black", size=14)),
-                  legend.title = element_text(colour="black", size=18),
-                  legend.text = element_text( size = 14))
-xx%>%
-  ggplot(.,aes(x=SingleVisitorEntryFee,y=PredictedFee))+
-  geom_jitter(size=2, alpha=0.3)+
-  geom_smooth(method="lm", se=FALSE, colour="black",size=2,linetype="longdash")+
-  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(xx$PredictedFee,xx$SingleVisitorEntryFee),digits=2)))+
-  ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+
-  theme(plot.title = element_text(hjust = 0.5))+
-  scale_y_continuous(name="Predicted Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))+
-  scale_x_continuous(name="Observed Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))
-```
-
 ### 2.1.3. Smoke Data
@@
 ### 2.2.2. Model Validation
-
-```{r echo=FALSE, message=FALSE,cache=TRUE}
-knitr::include_graphics(path.expand("~/SmokeProject/Train_Test.png"))
-```
 
@@
 
-## 3.3. Model Validation
+# 4. Discussion
+
+Discussion here
+
+# S. Supplemental
+
+## S.1. Visitation Fee Back Predicting Accuracy
+We then used this to predict what the fees would have been for every year we didn't have
+```{r cache=TRUE, include=FALSE}
+
+FEdat<-read.csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
+FEdat<-FEdat %>% dplyr::select(Park,Year,SingleVisitorEntryFee) #keep only these columns
+
+d<-na.omit(FEdat) #get rid of rows with any NA
+
+fillData<-FEdat[,1:2]
+
+FEdat$PredictedFee<-apply(posterior_predict(feefit,fillData),2,mean) #assuming this is for a new column
+
+```
+
+Our predicted fees correlate well with our observed fees that we have
+```{r echo = FALSE, cache=TRUE,message=FALSE,warning=FALSE}
+xx<-na.omit(FEdat)
+ mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+                  plot.title = element_text( size=18, color="black",face="bold"),
+                  axis.title.x = element_text( size=18),
+                  axis.title.y = element_text( size=18),
+                  axis.text=(element_text(color="black", size=14)),
+                  legend.title = element_text(colour="black", size=18),
+                  legend.text = element_text( size = 14))
+xx%>%
+  ggplot(.,aes(x=SingleVisitorEntryFee,y=PredictedFee))+
+  geom_jitter(size=2, alpha=0.3)+
+  geom_smooth(method="lm", se=FALSE, colour="black",size=2,linetype="longdash")+
+  geom_text(aes(x=5,y=15),size=10,label=paste("cor =",round(cor(xx$PredictedFee,xx$SingleVisitorEntryFee),digits=2)))+
+  ggtitle("Predicted vs Observed Single Visitor Entry Fee")+theme_classic()+mytheme+
+  theme(plot.title = element_text(hjust = 0.5))+
+  scale_y_continuous(name="Predicted Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))+
+  scale_x_continuous(name="Observed Single Visitor Entry Fee",labels = scales::dollar_format(prefix = "$"))
+```
+
+## S.2. Model Validation Process
+
+```{r echo=FALSE, message=FALSE,cache=TRUE}
+knitr::include_graphics(path.expand("~/SmokeProject/Train_Test.png"))
+```
+
+## S.3. Model Validation Results
 To assess the validity of our model, we used a k-fold cross validation technique.
@@
 ```
-
-# 4. Discussion
-
-Discussion here
-
-# 5. Supplemental

### 2020-06-01 Markdown Documents/Manuscript.Rmd
testing.
fixed x axis labels on fee revenue predictions vs observed plot. They were off by an order of magnitude

@@
 
-Alex to complete this section
+Will start this weekend.
 
@@
 ```{r echo=FALSE, cache=TRUE, warning=FALSE, message=FALSE, fig.height = 8, fig.width = 13}
-financialData<-merge(dat[,c(2,5)],financialData,by="UnitCode")
-financialData%>%filter(UnitCode %in% fitsig$Par)%>%group_by(ParkName)%>%
+financialData%>%filter(UnitCode %in% fitsig$Par)%>%group_by(Park)%>%
   summarise(sumincome=sum(incomeObserved),sumincomePred=sum(incomePred))%>%ungroup()%>%
@@
   ggplot(., 
-         aes(y = reorder(ParkName,diff),
+         aes(y = reorder(Park,diff),
              x = sumincome,
@@
                        colour_xend = "#41ab5d",dot_guide = T,dot_guide_size = NULL,dot_guide_colour = "lightgrey")+
-  geom_rect( aes(xmin=100500000000, xmax=109500000000, ymin=-Inf, ymax=Inf), fill="grey") +
-  geom_text( aes(label=paste0("-",round((diff/sumincome)*100,digits=1), "%"), y=ParkName, x=105000000000), fontface=2, size=4)+
-  annotate("text", x = 105000000000, y = "Glacier NP", label = "Difference",vjust=-1.5,fontface=2)+
+  geom_rect( aes(xmin=860000000, xmax=940000000, ymin=-Inf, ymax=Inf), fill="grey") +
+  geom_text( aes(label=paste0("-",round((diff/sumincome)*100,digits=1), "%"), y=Park, x=900000000), fontface=2, size=4)+
+  annotate("text", x = 900000000, y = "Glacier NP", label = "Difference",vjust=-1.5,fontface=2)+
   scale_x_continuous( labels = scales::dollar_format(prefix = "$")) +
-  annotate("text", x = 62000000000, y = "Yosemite NP", label = "Observed fee revenue",vjust=-1,hjust=1,fontface=2)+
-  annotate("text", x = 80000000000, y = "Yosemite NP", label = "Predicted fee revenue no smoke",vjust=-1,color="#41ab5d",fontface=2)+
+  annotate("text", x = 453357090, y = "Yosemite NP", label = "Observed fee revenue",vjust=-1,fontface=2)+
+  annotate("text", x = 682294835, y = "Yosemite NP", label = "Predicted fee revenue no smoke",vjust=-1,color="#41ab5d",fontface=2)+
   labs(title = "Total Lost Revenue",
@@
        y = "")+theme_classic()+mytheme+theme(plot.subtitle = element_text(face = "bold"))
-
 ```

### 2020-06-01 Rcode/HESBoelWorkingScript.R
fixed some labeling stuff on barbel plots

@@ -55,6 +55,7 @@ dat$PredNoSmoke25CI<-dat$trendvis+dat$PredNoSmoke25CI
 
 #write.csv(dat,file="Data/FullDataWithMinSmokePredictions.csv")
 dat<-read.csv("Data/FullDataWithMinSmokePredictions.csv")
+dat$date<-zoo::as.yearmon(dat$date)
 mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                 plot.title = element_text( size=18, color="black",face="bold"),
                 axis.title.x = element_text( size=18),
@@ -76,10 +77,9 @@ ggplot(.,aes(x=date))+
                        guide = "legend")+ggtitle("Rocky Mountain National Park")
 
 
-
   ggplot(dat,aes(x=date))+
   geom_line(aes(y=RecreationVisits),color="green")+
-  geom_line(aes(y=PredNoSmoke),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")+facet_wrap(~UnitCode,scales="free")
+  geom_line(aes(y=PredNoSmoke50CI),color="blue")+geom_smooth(aes(y=RecreationVisits),se=F,method="lm",color="red")+facet_wrap(~UnitCode,scales="free")
 
   p <- plot(fit, regex_pars = "stdsmoke UnitCode",
             prob = 0.5, prob_outer = 0.9)
@@ -120,6 +120,12 @@ summaryFinancialLoss<-financialData %>% group_by(UnitCode)%>%summarise(totrec=su
 summaryFinancialLoss$ParkName<-unique(dat[dat$UnitCode %in% summaryFinancialLoss$UnitCode,]$ParkName)
 
 summaryFinancialLoss<-summaryFinancialLoss[,c(6,4,5)]
+
+#summary stats
+sum(dat$RecreationVisits)/sum(dat$PredNoSmoke50CI)
+sum(summaryFinancialLoss$IncomeDif)
+##
+
 summaryFinancialLoss$IncomeDif<-formattable::currency(summaryFinancialLoss$IncomeDif)
 names(summaryFinancialLoss)<-(c("National Park","Income Loss Estimate (1980-2018)", "Percent Income Loss (1980-2018)"))
 summaryFinancialLoss$`Percent Income Loss (1980-2018)`<-paste0(summaryFinancialLoss$`Percent Income Loss (1980-2018)`,'%')
[…]

### 2020-06-01 Workflow Guidelines.Rmd
made a workflow guidelines document

---
title: "Workflow Guidelines"
author: "Matt Clark"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

## General workflow

1. Pull from Github. The Master should be linked to your R studio project so that you can Pull by hitting the blue arrow.

2. Make edits and add content committing frequently! Commits should be accompanied by detailed commit messages.

3. If working on a coding problem or specific figure, this should be done *outside* of the manuscript document. Once the figure/code chunk is complete, the code can be added to the manuscript document. This limits the probability of breaking the manuscript document for others while troubleshooting.

4. Knit to html and check that changes have rendered correctly.

5. Push commits. This can be done less frequently than specific commits. *NOTE*: Push only the manuscript.rmd document. Do not push the html document or the cache files.

## Specifics

### References
References are stored and pulled from the 'References.bib' file which is called in the 'Manuscript.rmd' document. Reference format is BibTeX. Citations can be accessed in this format from any citation manager (e.g. Zotero) or directly from Google Scholar. When new citations are added, they should be appended (copy and paste) to the bottom of the 'References.bib' file.

Here's an example:

"@article{richardson2012hidden,
  title={The hidden cost of wildfires: Economic valuation of health effects of wildfire smoke exposure in Southern California},
  author={Richardson, Leslie A and Champ, Patricia A and Loomis, John B},
  journal={Journal of Forest Economics},
  volume={18},
  number={1},
  pages={14--35},
  year={2012},
  publisher={Elsevier}
}"

To call this citation in the text of the document, paste the @article section surrounded by brackets e.g. [@richardson2012hidden]. To can multiple references, add them all into the brackets broken up by semicolons e.g. [@richardson2012hidden;@exampleRef2020]. *NOTE*: References will not appear in the literature cited section unless they are called in the in-text citations.

### Cache
We are caching all of the markdown code chunks to speed up html rendering (so it doesn't rerun the R code every time). This is good locally, but we don't want this stuff on the github. When creating an R code chunk, include the argument `cache = TRUE`. Do not commit or Push the cached files produced.

### Text comments
We want to be able to flag text and make comments about it. To do this, we will use double asterisk notation and bold our comments. Here's an example:

The United States national park system serves as a global model for how natural resource protection can provide numerous social and ecological services while generating economic returns on investment. In recent years parks have increasingly relied on visitation fees to fund conservation projects as well as basic infrastructure. In contrast with congressional funding, visitation is subject to fluctuation given climatic and environmental variability. While many environmental variables are relatively predictable and stable over time, wildfire smoke events have increased dramatically via climate change, particularly in the American West. In this project, we assess how recreational visitation to national parks has systematically varied with wildfire smoke. ** Over the last 38 years, we estimate that wildfire smoke has led to a 12% overall visitation decrease to National Parks in the Pacific West and Inter-Mountain Regions ** **Comment text**. When accounting for visitation fees, this loss may have cost the National Park Service as much as $580 million in total. Wildfire frequency and severity are expected to increase considerably in these regions over the next 50 years. Assuming current trends in smoke, park visitation, and park fees continue, we expect to see significant financial losses to the National Park Service via wildfire smoke in the coming years.

### Small text edits
When collaborators find small text errors (e.g. typos), they should make corrections without commenting. The commit message will serve as a comment about what changes were made in this scenario.

### File paths
When uploading data or model objects, relative file paths will be used so that all co-authors can pull/push the document without breaking it for themselves or others.

example: `load("../ModelObjects/feefit.rda")`

### R code
We do not want any R code or the accompanying messages/warnings to appear in the manuscript. R chunks should therefore include the following arguments:
`{r warning=FALSE, include=FALSE, cache=TRUE, message=FALSE}`

Time/memory intensive analyses should be included in the markdown document for reproducibility purposes, but should be evaluated *outside* of the Manuscript document and loaded in in a subsequent code chunk. For example:

`{r eval=FALSE, include=FALSE, cache=TRUE, message=FALSE}
fit<-stan_lmer(VisDiff~stdsmoke|UnitCode,iter=8000)
save(fit, file="residuals.rda")`

`{r  include=FALSE, cache=TRUE, message=FALSE}
load("../ModelObjects/residuals.rda")`

### 2020-06-02 Markdown Documents/Manuscript.Rmd
added some text to methods

@@
 abstract: The United States national park system serves as a global model for how natural resource protection can provide numerous social and ecological services while generating economic returns on investment.
-In recent years parks have increasingly relied on visitation fees to fund consevation projects as well as basic infrastructure.
+In recent years parks have increasingly relied on visitation fees to fund conservation projects as well as basic infrastructure.
 In contrast with congressional funding, visitation is subject to fluctuation given climatic and environmental variability.
-While many environmental variables are relatively predictable and stable over time, wildfire smoke events have increased drmatically via climate change, particurlarly in the American West.
+While many environmental variables are relatively predictable and stable over time, wildfire smoke events have increased dramatically via climate change, particularly in the American West.
 In this project, we assess how recreational visitation to national parks has systematically varied with wildfire smoke.
@@
 library(formattable)
-load("~/SmokeProject/ModelObjects/feefit.rda")
-load("~/SmokeProject/ModelObjects/residuals.rda")
-load("~/SmokeProject/ModelObjects/CVResidualMod.rds")
+load("../ModelObjects/feefit.rda")
+load("../ModelObjects/residuals.rda")
+load("../ModelObjects/CVResidualMod.rds")
 ```
@@
 
+used the way back machine
+
 We started with an incomplete dataset. Laura will write this part with information on collecting this data. 
 ```{r include=FALSE, cache=TRUE}
-FEdat<-read.csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
+FEdat<-read.csv("../Data/LauraStoddardEntryFeeCsv.csv")
 FEdat<-FEdat %>% dplyr::select(Park,Year,SingleVisitorEntryFee) #keep only these columns
@@
 library(rstan)
-Data<-read_csv("~/SmokeProject/Data/MergedDataComplete.csv")[,-1]
+Data<-read_csv("../Data/MergedDataComplete.csv")[,-1]
 Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
@@
 
-dat<-read.csv("~/SmokeProject/Data/FullDataWithMinSmokePredictions.csv")
+dat<-read.csv("../Data/FullDataWithMinSmokePredictions.csv")
 dat$date<-zoo::as.yearmon(dat$date)
@@
 #predicted vis fees 
-FEdat<-read.csv("~/SmokeProject/Data/ParkFees.csv")
+FEdat<-read.csv("../Data/ParkFees.csv")
 
@@
 
-FEdat<-read.csv("~/SmokeProject/Data/LauraStoddardEntryFeeCsv.csv")
+FEdat<-read.csv("../Data/LauraStoddardEntryFeeCsv.csv")
 FEdat<-FEdat %>% dplyr::select(Park,Year,SingleVisitorEntryFee) #keep only these columns
@@
 ```{r echo=FALSE, message=FALSE,cache=TRUE}
-knitr::include_graphics(path.expand("~/SmokeProject/Train_Test.png"))
+knitr::include_graphics(path.expand("../Train_Test.png"))
 ```
@@
 #save(fitTrain,file="CVResidualMod.rds")
-load("~/SmokeProject/ModelObjects/CVResidualMod.rds")
+load("../ModelObjects/CVResidualMod.rds")
 
@@
 #save(fitTrain2,file="CVResidualMod2.rds")
-load("~/SmokeProject/ModelObjects/CVResidualMod2.rds")
+load("../ModelObjects/CVResidualMod2.rds")
 
@@
 #save(fitTrain3,file="CVResidualMod3.rds")
-load("~/SmokeProject/ModelObjects/CVResidualMod3.rds")
+load("../ModelObjects/CVResidualMod3.rds")
 
@@
 #save(fitTrain4,file="CVResidualMod4.rds")
-load("~/SmokeProject/ModelObjects/CVResidualMod4.rds")
+load("../ModelObjects/CVResidualMod4.rds")
 
@@
 #save(fitTrain5,file="CVResidualMod5.rds")
-load("~/SmokeProject/ModelObjects/CVResidualMod5.rds")
+load("../ModelObjects/CVResidualMod5.rds")
 
@@
 #save(fitTrain6,file="CVResidualMod6.rds")
-load("~/SmokeProject/ModelObjects/CVResidualMod6.rds")
+load("../ModelObjects/CVResidualMod6.rds")
 
@@
 #save(fitTrain7,file="CVResidualMod7.rds")
-load("~/SmokeProject/ModelObjects/CVResidualMod7.rds")
+load("../ModelObjects/CVResidualMod7.rds")
 

### 2020-06-02 Markdown Documents/Manuscript.Rmd
added a space
Practice Commit
small change

@@
 Given the variability in recreational fees between parks and over time, the secondary goal of this project is to produce order of magnitude estimates of the financial losses that wildfire smoke has caused park units during the same period.
-The intention of this research is not to perform an in-depth economic analysis of smoke effects on each park specifically, but rather to provide an approximation of how wildfire smoke affects population level recreational decision making and what the downstream economic impacts have been historically and what the implications are for the future.
+The intention of this research is not to perform an in-depth economic analysis of smoke effects on each parks specifically, but rather to provide an approximation of how wildfire smoke affects population level recreational decision making and what the downstream economic impacts have been historically and what the implications are for the future.
 Overall, this research aims to better equip protected area managers with the information necessary to best manage recreational resources in the face of increasing wildfire events and climatic variability generally.

### 2020-06-25 Rcode/Carbon_Analysis.Rmd
new analysis

---
title: 'MERRA-2: Carbon Analysis'
author: "Alex Killion"
date: "6/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-06-29 Markdown Documents/Manuscript.Rmd
Updated vis fee methods section

@@
 
-used the way back machine
+Historic visitation fee data is not readily available from the National Park Service.
+The recreational visitor fees however, have been listed on the individual park websites historically.
+We accessed these historic webpages and aggregated the single visitor entry fee data using the [Internet Archive](https://archive.org/web/).
+We were able to access historic webpages for the years 1997-2018.
+Archived web pages were unavailable for years prior to 1997.
+For years 1980-1996, we performed trendline extension via linear regression to estimate the recreational visitor fee for each park, each year per the following:
 
-We started with an incomplete dataset. Laura will write this part with information on collecting this data. 
 ```{r include=FALSE, cache=TRUE}
@@
 
-We removed the NA's from this dataset built a linear model which estimates the fees based on the date and the park. 
 ```{r, echo=FALSE, cache=TRUE}
@@
 \begin{aligned}
-y_{tp} &\sim \begin{cases}\operatorname{Pois}(\mu,\lambda)\\ \mu = B_{0}+ B_{1}*X_{1} +B_{2}*X_{2}
+y_{tp} &\sim \begin{cases}\operatorname{Pois}(\mu,\lambda)\\ \mu = B_{0}+ B_{t}*X_{t} +B_{p}*X_{p}
 \end{cases}
@@
 
-We operationalized this model in rstanarm
+which yields the visitation estimate \(y\) for a given park \(_{p}\) in a given year \(_{t}\). Model details and validation can be found in Supplemental (S1).
+
+To estimate the visitor revenue for each park, each year, we multiplied the observed or estimated park fee by the number of visitors recorded for each park.
+This is an oversimplification of the visitation fee system and likely overestimates the fees paid by recreational visitors as it does not account for multiple visitors who may share one vehicle.
+As a result of this discrepancy, we caution that backcasted income losses should be taken at an order of magnitude level only, not as specific dollar estimates
 

### 2020-07-01 Rcode/Carbon_Analysis.Rmd
python

@@
 ---
+```{r}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+# ------------------------------------------------------------------------------------------------ #
+# How to Access the LP DAAC Data Pool with R
+# The following R code example demonstrates how to configure a connection to download data from an
+# Earthdata Login enabled server, specifically the LP DAAC Data Pool.
+# ------------------------------------------------------------------------------------------------ #
+
+if ("sys" %in% rownames(installed.packages()) == FALSE) {install.packages("sys")}
+if ("getPass" %in% rownames(installed.packages()) == FALSE) { install.packages("getPass")}
+if ("httr" %in% rownames(installed.packages()) == FALSE) { install.packages("httr")}
+
+# Load necessary packages into R
+library(sys)
+library(getPass)
+library(httr)
+# ---------------------------------SET UP ENVIRONMENT--------------------------------------------- #
+# IMPORTANT: Update the line below if you want to download to a different directory (ex: "c:/data/")
+dl_dir <- Sys.getenv("HOME")                                 # Set dir to download files to
+setwd(dl_dir)                                                # Set the working dir to the dl_dir
+usr <- file.path(Sys.getenv("USERPROFILE"))                  # Retrieve home dir (for netrc file)
+if (usr == "") {usr = Sys.getenv("HOME")}                    # If no user profile exists, use home
+netrc <- file.path(usr,'.netrc', fsep = .Platform$file.sep)  # Path to netrc file
+
+# ------------------------------------CREATE .NETRC FILE------------------------------------------ #
+# If you already have a .netrc file with your Earthdata Login credentials stored in your home
+# directory, this portion will be skipped. Otherwise you will be prompted for your NASA Earthdata
+# Login Username/Password and a netrc file will be created to store your credentials (in home dir)
+if (file.exists(netrc) == FALSE || grepl("urs.earthdata.nasa.gov", readLines(netrc)) == FALSE) {
+  netrc_conn <- file(netrc)
+
+  # User will be prompted for NASA Earthdata Login Username and Password below
+  writeLines(c("machine urs.earthdata.nasa.gov",
+               sprintf("login %s", getPass(msg = "Enter NASA Earthdata Login Username \n (or create an account at urs.earthdata.nasa.gov) :")),
+               sprintf("password %s", getPass(msg = "Enter NASA Earthdata Login Password:"))), netrc_conn)
+  close(netrc_conn)
+  }
+# ---------------------------CONNECT TO DATA POOL AND DOWNLOAD FILES------------------------------ #
+# Below, define either a single link to a file for download, a list of links, or a text file
+# containing links to the desired files to download. For a text file, there should be 1 file link
+# listed per line. Here we show examples of each of the three ways to download files.
+# **IMPORTANT: be sure to update the links for the specific files you are interested in downloading.
+
+# 1. Single file (this is just an example link, replace with your desired file to download):
+files <- "https://goldsmr5.gesdisc.eosdis.nasa.gov:443/opendap/MERRA2/M2I3NVAER.5.12.4/1980/01/MERRA2_100.inst3_3d_aer_Nv.19800101.nc4"
+
+# 2. List of files (these are just example links, replace with your desired files to download:
+#files <- c("https://e4ftl01.cr.usgs.gov/MOLA/MYD09GA.006/2002.07.06/MYD09GA.A2002187.h10v04.006.2015149103018.hdf",
+#           "https://e4ftl01.cr.usgs.gov/MOLT/MOD11A1.006/2000.03.09/MOD11A1.A2000069.h00v08.006.2015057070313.hdf")
+
+# 3. Textfile containing links (just an example, replace with your text file location):
+#files <- readLines("C:/datapool_downloads/URL_file_list.txt", warn = FALSE)
+
+# Loop through all files
+for (i in 1:length(files)) {
+  filename <-  tail(strsplit(files[i], '/')[[1]], n = 1) # Keep original filename
+
+  # Write file to disk (authenticating with netrc) using the current directory/filename
+  response <- GET(files[i], write_disk(filename, overwrite = TRUE), progress(),
+                  config(netrc = TRUE, netrc_file = netrc), set_cookies("LC" = "cookies"))
+
+  # Check to see if file downloaded correctly
+  if (response$status_code == 200) {
+    print(sprintf("%s downloaded at %s", filename, dl_dir))
+  } else {
+  print(sprintf("%s not downloaded. Verify that your username and password are correct in %s", filename, netrc))
+  }
+}
+
 ```
-
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-07-01 Rcode/Carbon_Analysis.Rmd
netcdf

--- 2020-06-25 Rcode/Carbon_Analysis.Rmd
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r}
+https://goldsmr5.gesdisc.eosdis.nasa.gov/opendap/MERRA2/M2I3NVAER.5.12.4/1980/01/contents.html
+
+url = "https://goldsmr5.gesdisc.eosdis.nasa.gov/opendap/MERRA2/M2I3NVAER.5.12.4/1980/01/MERRA2_100.inst3_3d_aer_Nv.19800101.nc4"
+data = nc_open(url)
+
+#have to delete duplicate ".nc4" in url
+url2 = "https://goldsmr5.gesdisc.eosdis.nasa.gov/opendap/MERRA2/M2I3NVAER.5.12.4/2018/07/MERRA2_400.inst3_3d_aer_Nv.20180708.nc4?BCPHILIC[0:7][0:71][225:277][75:177],OCPHILIC[0:7][0:71][225:277][75:177],BCPHOBIC[0:7][0:71][225:277][75:177],OCPHOBIC[0:7][0:71][225:277][75:177],lat[225:277],time[0:7],lon[75:177],lev"
+data2 = nc_open(url2)
+
 ```
 
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
 ```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2020-07-15 Rcode/MERRA2_File_organization.R
file org code for alex

@@ -1,0 +1,38 @@
+#make fake file names
+
+MERRA2_300.inst3_3d_aer_Nv.20101218.nc4.nc4
+
+
+MER300<-vector()
+for (i in c("01","02","03","04","05","06","07","08","09","11","12")){
+  for (j in c(2001,2002,2003,2004,2005,2006,2007,2008,2009,2010)){
+    for (l in c("01","05","10","25")){
+      
+    y<-paste0("MERRA2_300.inst3_3d_aer_Nv.",j,i,l,".nc4.nc4")
+    MER300<-c(MER300,y)
+}}}
+
+MER200<-vector()
+for (i in c("01","02","03","04","05","06","07","08","09","11","12")){
+  for (j in c(1991,1992,1993,1994,1995,1996,1997,1998,1999,2000)){
+    for (l in c("01","05","10","25")){
+      
+      y<-paste0("MERRA2_300.inst3_3d_aer_Nv.",j,i,l,".nc4.nc4")
+      MER200<-c(MER200,y)
+    }}}
+
+MER<-c(MER200,MER300)
+
+
+
+####Now let's just make individual batches for each month/year combination
+
+for (i in c(1991,1992,1993,1994,1995,1996,1997,1998,1999,2000)){
+for (j in c("01","02","03","04","05","06","07","08","09","11","12")){
+    x<-(paste0(i,j))
+    assign(paste0("x",i,j),MER[grep(x,MER)])
+    
+}}
+
+x199101
+

### 2020-07-15 Rcode/MERRA2_File_organization.R
small merra_200 vs 300 change

@@ -17,7 +17,7 @@ for (i in c("01","02","03","04","05","06","07","08","09","11","12")){
   for (j in c(1991,1992,1993,1994,1995,1996,1997,1998,1999,2000)){
     for (l in c("01","05","10","25")){
       
-      y<-paste0("MERRA2_300.inst3_3d_aer_Nv.",j,i,l,".nc4.nc4")
+      y<-paste0("MERRA2_200.inst3_3d_aer_Nv.",j,i,l,".nc4.nc4")
       MER200<-c(MER200,y)
     }}}
 

### 2020-07-15 Rcode/MERRA2_File_organization.R
create file batches with names from folder

@@ -28,11 +28,28 @@ MER<-c(MER200,MER300)
 ####Now let's just make individual batches for each month/year combination
 
 for (i in c(1991,1992,1993,1994,1995,1996,1997,1998,1999,2000)){
-for (j in c("01","02","03","04","05","06","07","08","09","11","12")){
+for (j in c("01","02","03","04","05","06","07","08","09","10","11","12")){
     x<-(paste0(i,j))
     assign(paste0("x",i,j),MER[grep(x,MER)])
     
 }}
 
 x199101
+
+
+#Now do it with the files stored in a folder
+fileNames<-list.files("~/SmokeProject/MERRA_DATA_ORG_PRACT")
+for (i in c(1980,2004)){
+  for (j in c("01","10")){
+    x<-(paste0(i,j))
+    assign(paste0("x",i,j),fileNames[grep(x,fileNames)])
+    
+  }}
+
+
+x200410
+
+
+
+
 

### 2020-07-15 Rcode/MERRA2_File_organization.R
added comments

@@ -38,16 +38,16 @@ x199101
 
 
 #Now do it with the files stored in a folder
-fileNames<-list.files("~/SmokeProject/MERRA_DATA_ORG_PRACT")
-for (i in c(1980,2004)){
-  for (j in c("01","10")){
+fileNames<-list.files("~/SmokeProject/MERRA_DATA_ORG_PRACT") #this is a practice example of what the actual data foler will look like
+for (i in c(1980,2004)){ #this will have all the years whe we actually do it. The example folder only has 1980 and 2004
+  for (j in c("01","10")){ #same as above but for months
     x<-(paste0(i,j))
     assign(paste0("x",i,j),fileNames[grep(x,fileNames)])
     
   }}
 
 
-x200410
+x200410 #this will give you all the file names from the folder that are from oct 2004 regardless of the DAY
 
 
 

### 2020-07-16 Rcode/Merra2.Rmd
clean

---
title: "Merra2"
author: "Alex Killion"
date: "11/26/2019"
output: html_document
---

Packages
```{r}
library(ncdf4)
library(chron)
library(lattice)
library(RColorBrewer)

library(raster)
library(rgdal)
library(exactextractr)
library(readr)
library(plyr)
library(purrr)
```

*For MATRIX*
Open netCDF
```{r}
# set path and filename
ncpath <- "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/Merra2/"
ncname <- "MERRA2_100.tavgM_2d_aer_Nx.198001.SUB"
ncfname <- paste(ncpath, ncname, ".nc", sep="")
dname <- "DUSMASS25"

ncin <- nc_open(ncfname)

# get longitude and latitude and time
lon <- ncvar_get(ncin,"lon")
nlon <- dim(lon)
lat <- ncvar_get(ncin,"lat")
nlat <- dim(lat)
print(c(nlon,nlat))

time <- ncvar_get(ncin,"time")

tunits <- ncatt_get(ncin,"time","units")
nt <- dim(time) #dim() tells you number of timesteps
nt
```

Extract a Variable
```{r}
# get temperature
tmp_array <- ncvar_get(ncin,dname) #dname in 1st chunk needs to be correct variable name
dlname <- ncatt_get(ncin,dname,"long_name")
dunits <- ncatt_get(ncin,dname,"units")
fillvalue <- ncatt_get(ncin,dname,"_FillValue")
dim(tmp_array)

#close file
nc_close(ncfname)
```

Convert Time
```{r}
# convert time -- split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])
chron(time,origin=c(tmonth, tday, tyear))
```

Replace with NA
```{r}
# replace netCDF fill values with NA's
tmp_array[tmp_array==fillvalue$value] <- NA
```

Map
```{r}
# quick map
image(lon,lat,tmp_array, col=rev(brewer.pal(10,"RdBu")))
```

*For Raster*
Stack netCDF
```{r}
filenames <- list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/Merra2/",
                        pattern='*.nc',full.names=TRUE)
#if you want to stack
#smoke <- raster::stack(filenames, varname="DUSMASS25")

```

Import NPS Boundaries
```{r}
#Data From - https://public-nps.opendata.arcgis.com/datasets/national-park-service-park-unit-boundaries

nps<-readOGR("/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/NPS_poly/NPS_poly.shp")

#select only polygons designated as National Parks
nps<-nps[nps@data$UNIT_TYPE=="National Park",]

#remove National Parks outside our Regions
nps<-nps[nps@data$REGION=="PW"|nps@data$REGION=="IM",]
nps<-nps[nps@data$UNIT_CODE!="NPSA",] #Samoa
nps<-nps[nps@data$STATE!="HI",]

#reformat to use park codes later
nps_df<-as.data.frame(nps)
nps_df<-as.matrix(nps_df$UNIT_CODE)

park_codes<-as.data.frame(nps_df)
names(park_codes)<-"park"

```

New data mean in Parks
```{r}
require(exactextractr)

#exact_extract - https://cran.r-project.org/web/packages/exactextractr/readme/README.html * see for other summary stats
poly<-sf::st_as_sfc(nps) #converts to new type of poly

#Get Merra2 files

fileNames<-list.files("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT")

for (i in c(1980,2004)){ #this will have all the years whe we actually do it. The example folder only has 1980 and 2004
  for (j in c("01","10")){ #same as above but for months
    x<-(paste0(i,j))
  assign(paste0("x",i,j),fileNames[grep(x,fileNames)])
  }}

#fix
x198001<-paste0("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT/",x198001) #need full filepath - can you add in loop above??
x200410<-paste0("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT/",x200410)

list_all<- list(x198001,x200410) #will need to output all lists above in for loop into another list

#the "level" dimension refers to veritcal layers where 72=surface and 1=top of atmosphere
#for a single day

for (i in 1:length (list_all)) {
    for (j in 1:length (list_all[i])){
 bcphilic<-mean(brick(list_all[[i]][[j]], varname="BCPHILIC", level=72)) #one day - first bracket is month, 2nd is day, within a day is 8 layers by hour
bcphobic<-mean(brick(list_all[[i]][[j]], varname="BCPHOBIC", level=72))
ocphilic<-mean(brick(list_all[[i]][[j]], varname="OCPHILIC", level=72))
ocphobic<-mean(brick(list_all[[i]][[j]], varname="OCPHOBIC", level=72))

all<-bcphilic+bcphobic+ocphilic+ocphobic

#Make list of all objects - assign(paste0(“day”,j), <- exact_extract(all, poly, 'mean') -

mean<-exact_extract(raster, poly, 'mean') #extracts mean for each park. Uses cells that intersect polygons.

}

#take mean of the assign list above
#write csv
  n<-substr(list_all[[1]][[1]],94,100)
    name<-paste0("/Users/killion/",n)
    newname<-paste0(name, ".csv")
write.csv(mean, newname, row.names = FALSE) #save as csv

  }

#loads those .csv and makes into single df
list_dates = list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_mean/", pattern="*.csv", full.names=TRUE)

final_means<-list_dates %>%
  map(read_csv) %>%
  reduce(cbind)

#append park code
final_means<-data.frame(park=park_codes, final_means)

#save
write.csv(final_means, "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/interm_data/final_means.csv", row.names = FALSE)

```

Mean PM2.5 in Parks
```{r}

#exact_extract - https://cran.r-project.org/web/packages/exactextractr/readme/README.html * see for other summary stats
poly<-sf::st_as_sfc(nps) #converts to new type of poly

filenames <- list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/Merra2/",
                        pattern='*.nc',full.names=TRUE)

#loop to calculate polygons means and save vector
  for (i in 1:length (filenames)) {
    raster<-raster(filenames[i], varname="DUSMASS25")#convert smoke data to raster
    mean<-exact_extract(raster, poly, 'mean') #extracts mean for each park. Uses cells that intersect polygons.
   date<-str_sub(filenames[i], -13,-8) #strips down filename to just date
  mean<-as.data.frame(mean) #convert to df
  names(mean)<-date #rename column with date
  date_file<-paste0("/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_mean/",date)
   newname<-paste0(date_file, ".csv")
write.csv(mean, newname, row.names = FALSE) #save as csv
}

#loads those .csv and makes into single df
list_dates = list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_mean/", pattern="*.csv", full.names=TRUE)

final_means<-list_dates %>%
  map(read_csv) %>%
  reduce(cbind)

#append park code
final_means<-data.frame(park=park_codes, final_means)

#save
write.csv(final_means, "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/interm_data/final_means.csv", row.names = FALSE)

```

Min PM2.5 in Parks
```{r}

#exact_extract - https://cran.r-project.org/web/packages/exactextractr/readme/README.html * see for other summary stats
poly<-sf::st_as_sfc(nps) #converts to new type of poly

#loop to calculate polygons means and save vector
  for (i in 1:length (filenames)) {
    raster<-raster(filenames[i], varname="DUSMASS25")#convert smoke data to raster
    min<-exact_extract(raster, poly, 'min') #extracts mean for each park. Uses cells that intersect polygons.
   date<-str_sub(filenames[i], -13,-8) #strips down filename to just date
  min<-as.data.frame(min) #convert to df
  names(min)<-date #rename column with date
  date_file<-paste0("/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_min/",date)
   newname<-paste0(date_file, ".csv")
write.csv(min, newname, row.names = FALSE) #save as csv
}

#loads those .csv and makes into single df
list_dates = list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_min/", pattern="*.csv", full.names=TRUE)

final_mins<-list_dates %>%
  map(read_csv) %>%
  reduce(cbind)

#append park code
final_mins<-data.frame(park=park_codes, final_mins)

#save
write.csv(final_mins, "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/interm_data/final_mins.csv", row.names = FALSE)

```

Max PM2.5 in Parks
```{r}

#exact_extract - https://cran.r-project.org/web/packages/exactextractr/readme/README.html * see for other summary stats
poly<-sf::st_as_sfc(nps) #converts to new type of poly

#loop to calculate polygons means and save vector
  for (i in 1:length (filenames)) {
    raster<-raster(filenames[i], varname="DUSMASS25")#convert smoke data to raster
    max<-exact_extract(raster, poly, 'max') #extracts mean for each park. Uses cells that intersect polygons.
   date<-str_sub(filenames[i], -13,-8) #strips down filename to just date
  max<-as.data.frame(max) #convert to df
  names(max)<-date #rename column with date
  date_file<-paste0("/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_max/",date)
   newname<-paste0(date_file, ".csv")
write.csv(max, newname, row.names = FALSE) #save as csv
}

#loads those .csv and makes into single df
list_dates = list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_max/", pattern="*.csv", full.names=TRUE)

final_max<-list_dates %>%
  map(read_csv) %>%
  reduce(cbind)

#append park code
final_max<-data.frame(park=park_codes, final_max)

#save
write.csv(final_max, "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/interm_data/final_max.csv", row.names = FALSE)

```

### 2020-07-16 Rcode/MERRA2_File_organization.R
list of file names for each month

@@ -47,9 +47,20 @@ for (i in c(1980,2004)){ #this will have all the years whe we actually do it. Th
   }}
 
 
-x200410 #this will give you all the file names from the folder that are from oct 2004 regardless of the DAY
+###same thing, but make a list of all the file names
+years<-c(1980,2004)
+months<-c("01","10")
+comb<-expand.grid(years,months)
+empty_list<-list() #create an empty list to fill with the file names
 
+for (i in 1:nrow(comb)){
+  for(j in years){
+    for(k in months){
+x<-(paste0(j,k))
+empty_list[[i]]<-fileNames[grep(x,fileNames)]
+names(empty_list)[i]<-paste0("x",j,k)
 
+}}}
 
 
 

### 2020-07-16 Rcode/MERRA2_File_organization.R
added comments

@@ -48,17 +48,17 @@ for (i in c(1980,2004)){ #this will have all the years whe we actually do it. Th
 
 
 ###same thing, but make a list of all the file names
-years<-c(1980,2004)
-months<-c("01","10")
-comb<-expand.grid(years,months)
+years<-c(1980,2004)  #all the years
+months<-c("01","10") #all the months
+comb<-expand.grid(years,months) #just to get the number of unique combinations
 empty_list<-list() #create an empty list to fill with the file names
 
-for (i in 1:nrow(comb)){
+for (i in 1:nrow(comb)){ #creates the number of list objects
   for(j in years){
     for(k in months){
 x<-(paste0(j,k))
-empty_list[[i]]<-fileNames[grep(x,fileNames)]
-names(empty_list)[i]<-paste0("x",j,k)
+empty_list[[i]]<-fileNames[grep(x,fileNames)]  #fill the list with the file names
+names(empty_list)[i]<-paste0("x",j,k) #name the list objects
 
 }}}
 

### 2020-07-16 Rcode/MERRA2_File_organization.R
fixed replicate month issue

@@ -47,6 +47,8 @@ for (i in c(1980,2004)){ #this will have all the years whe we actually do it. Th
   }}
 
 
+
+
 ###same thing, but make a list of all the file names
 years<-c(1980,2004)  #all the years
 months<-c("01","10") #all the months
@@ -54,13 +56,11 @@ comb<-expand.grid(years,months) #just to get the number of unique combinations
 empty_list<-list() #create an empty list to fill with the file names
 
 for (i in 1:nrow(comb)){ #creates the number of list objects
-  for(j in years){
-    for(k in months){
-x<-(paste0(j,k))
-empty_list[[i]]<-fileNames[grep(x,fileNames)]  #fill the list with the file names
-names(empty_list)[i]<-paste0("x",j,k) #name the list objects
-
-}}}
-
-
+  year<-comb[i,]$Var1
+  month<-comb[i,]$Var2
+      x<-(paste0(year,month))
+      empty_list[[i]]<-fileNames[grep(x,fileNames)]  #fill the list with the file names
+      names(empty_list)[i]<-paste0("x",year,month) #name the list objects
+      
+    }
 

### 2020-07-16 Rcode/Merra2.Rmd
got it

@@
 
-fileNames<-list.files("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT")
+fileNames<-list.files("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT",full.names = TRUE)
 
@@
 
-#fix
-x198001<-paste0("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT/",x198001) #need full filepath - can you add in loop above??
-x200410<-paste0("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT/",x200410)
+#fix - may not have to do this now
+# x198001<-paste0("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT/",x198001) #need full filepath - can you add in loop above??
+# x200410<-paste0("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT/",x200410)
 
@@
 #for a single day
+empty_list<-list()
 
 for (i in 1:length (list_all)) {
-    for (j in 1:length (list_all[i])){
+    for (j in 1:length (list_all[[i]])){
  bcphilic<-mean(brick(list_all[[i]][[j]], varname="BCPHILIC", level=72)) #one day - first bracket is month, 2nd is day, within a day is 8 layers by hour
@@
 #Make list of all objects - assign(paste0(“day”,j), <- exact_extract(all, poly, 'mean') - 
-
-mean<-exact_extract(raster, poly, 'mean') #extracts mean for each park. Uses cells that intersect polygons.
+empty_list[[j]]<-exact_extract(all, poly, 'mean') #intersects boundary
 
     }
+  #take mean of the assign list above
 
-#take mean of the assign list above
+out<-as.data.frame(empty_list)
+out_mean<-as.data.frame(rowMeans(out))
+
 #write csv
-  n<-substr(list_all[[1]][[1]],94,100)
-    name<-paste0("/Users/killion/",n)
+  n<-substr(list_all[[i]][[j]],94,100) # !!!!!!!! you will have to change this based on the length of your filepaths
+  colnames(out_mean)[1] <- n #give values column name based on month
+   name<-paste0("/Users/killion/output_means/",n) #create empty folder to save to
     newname<-paste0(name, ".csv")
-write.csv(mean, newname, row.names = FALSE) #save as csv
+write.csv(out_mean, newname, row.names = FALSE) #save as csv
 
@@
 #loads those .csv and makes into single df
-list_dates = list.files(path="/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/loop_output_mean/", pattern="*.csv", full.names=TRUE)
+list_dates = list.files(path="/Users/killion/output_means/", pattern="*.csv", full.names=TRUE) #same folder you saved above output
 
@@
 #save
-write.csv(final_means, "/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/interm_data/final_means.csv", row.names = FALSE)
+write.csv(final_means, "/Users/killion/final_means.csv", row.names = FALSE)
 

### 2020-07-16 Rcode/For_Clark.Rmd
final

---
title: "For_Clark"
author: "Alex Killion"
date: "7/16/2020"
output: html_document
---

Packages
```{r}
library(ncdf4)
library(chron)
library(lattice)
library(RColorBrewer)

library(raster)
library(rgdal)
library(exactextractr)
library(readr)
library(plyr)
library(purrr)
```

Import NPS Boundaries 
```{r}
#Data From - https://public-nps.opendata.arcgis.com/datasets/national-park-service-park-unit-boundaries

nps<-readOGR("/Volumes/GoogleDrive/My Drive/Local/Projects/Smoke/Smoke-Signal/NPS_poly/NPS_poly.shp")

#select only polygons designated as National Parks
nps<-nps[nps@data$UNIT_TYPE=="National Park",]

#remove National Parks outside our Regions
nps<-nps[nps@data$REGION=="PW"|nps@data$REGION=="IM",]
nps<-nps[nps@data$UNIT_CODE!="NPSA",] #Samoa
nps<-nps[nps@data$STATE!="HI",]

#reformat to use park codes later
nps_df<-as.data.frame(nps)
nps_df<-as.matrix(nps_df$UNIT_CODE)

park_codes<-as.data.frame(nps_df)
names(park_codes)<-"park"

```

New data mean in Parks 
L71 will need to be unhashed when working with our sample data - it selects only the months which have data in folder
L93 values need to change based on the length of your new filepath
```{r}

poly<-sf::st_as_sfc(nps) #converts to new type of poly

fileNames<-list.files("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT",full.names = TRUE)

years<-c(1980,2004)  #all the years
months<-c("01","10") #all the months
comb<-expand.grid(years,months) #just to get the number of unique combinations
list_all<-list() #create an empty list to fill with the file names

for (i in 1:nrow(comb)){ #creates the number of list objects
  year<-comb[i,]$Var1
  month<-comb[i,]$Var2
      x<-(paste0(year,month))
      list_all[[i]]<-fileNames[grep(x,fileNames)]  #fill the list with the file names
      names(list_all)[i]<-paste0("x",year,month) #name the list objects
}

# list_all<- list_all[c(1,4)] You will need to use this with our sample data folder

empty_list<-list()

for (i in 1:length (list_all)) {
    for (j in 1:length (list_all[[i]])){
 bcphilic<-mean(brick(list_all[[i]][[j]], varname="BCPHILIC", level=72)) #one day - first bracket is month, 2nd is day, within a day is 8 layers by hour - the "level" dimension refers to veritcal layers where 72=surface and 1=top of atmosphere
bcphobic<-mean(brick(list_all[[i]][[j]], varname="BCPHOBIC", level=72))
ocphilic<-mean(brick(list_all[[i]][[j]], varname="OCPHILIC", level=72))
ocphobic<-mean(brick(list_all[[i]][[j]], varname="OCPHOBIC", level=72))

all<-bcphilic+bcphobic+ocphilic+ocphobic

empty_list[[j]]<-exact_extract(all, poly, 'mean') #intersects boundary

    }

out<-as.data.frame(empty_list)
out_mean<-as.data.frame(rowMeans(out)) #take mean of the assign list above

#write csv
  n<-substr(list_all[[i]][[j]],94,100) # !!!!!!!! you will have to change 94 and 100 based on the length of your filepaths - this selects just the year and month i.e. 20041008
  colnames(out_mean)[1] <- n #give values column name based on month
   name<-paste0("/Users/killion/output_means/",n) #create empty folder to save to
    newname<-paste0(name, ".csv")
write.csv(out_mean, newname, row.names = FALSE) #save as csv

  }

#loads those .csv and makes into single df
list_dates = list.files(path="/Users/killion/output_means/", pattern="*.csv", full.names=TRUE) #same folder you saved above output

final_means<-list_dates %>%
  map(read_csv) %>%
  reduce(cbind)

#append park code
final_means<-data.frame(park=park_codes, final_means)

#save
write.csv(final_means, "/Users/killion/final_means2.csv", row.names = FALSE)

```

### 2020-07-16 Rcode/Merra2.Rmd
final

@@
 ```{r}
-require(exactextractr)
 
-#exact_extract - https://cran.r-project.org/web/packages/exactextractr/readme/README.html * see for other summary stats
 poly<-sf::st_as_sfc(nps) #converts to new type of poly
-
-#Get Merra2 files
 
@@
 
-for (i in c(1980,2004)){ #this will have all the years whe we actually do it. The example folder only has 1980 and 2004
-  for (j in c("01","10")){ #same as above but for months
-    x<-(paste0(i,j))
-  assign(paste0("x",i,j),fileNames[grep(x,fileNames)])
-  }}
+years<-c(1980,2004)  #all the years
+months<-c("01","10") #all the months
+comb<-expand.grid(years,months) #just to get the number of unique combinations
+list_all<-list() #create an empty list to fill with the file names
 
-#fix - may not have to do this now
-# x198001<-paste0("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT/",x198001) #need full filepath - can you add in loop above??
-# x200410<-paste0("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT/",x200410)
+for (i in 1:nrow(comb)){ #creates the number of list objects
+  year<-comb[i,]$Var1
+  month<-comb[i,]$Var2
+      x<-(paste0(year,month))
+      list_all[[i]]<-fileNames[grep(x,fileNames)]  #fill the list with the file names
+      names(list_all)[i]<-paste0("x",year,month) #name the list objects
+}
 
-list_all<- list(x198001,x200410) #will need to output all lists above in for loop into another list
+# list_all<- list_all[c(1,4)] You will need to use this with our sample data folder
 
-#the "level" dimension refers to veritcal layers where 72=surface and 1=top of atmosphere
-#for a single day
 empty_list<-list()
@@
     for (j in 1:length (list_all[[i]])){
- bcphilic<-mean(brick(list_all[[i]][[j]], varname="BCPHILIC", level=72)) #one day - first bracket is month, 2nd is day, within a day is 8 layers by hour
+ bcphilic<-mean(brick(list_all[[i]][[j]], varname="BCPHILIC", level=72)) #one day - first bracket is month, 2nd is day, within a day is 8 layers by hour - the "level" dimension refers to veritcal layers where 72=surface and 1=top of atmosphere
 bcphobic<-mean(brick(list_all[[i]][[j]], varname="BCPHOBIC", level=72))
@@
 
-#Make list of all objects - assign(paste0(“day”,j), <- exact_extract(all, poly, 'mean') - 
 empty_list[[j]]<-exact_extract(all, poly, 'mean') #intersects boundary
@@
     }
-  #take mean of the assign list above
 
 out<-as.data.frame(empty_list)
-out_mean<-as.data.frame(rowMeans(out))
+out_mean<-as.data.frame(rowMeans(out)) #take mean of the assign list above
 
 #write csv
-  n<-substr(list_all[[i]][[j]],94,100) # !!!!!!!! you will have to change this based on the length of your filepaths
+  n<-substr(list_all[[i]][[j]],94,100) # !!!!!!!! you will have to change 94 and 100 based on the length of your filepaths - this selects just the year and month i.e. 20041008
   colnames(out_mean)[1] <- n #give values column name based on month
@@
 #save
-write.csv(final_means, "/Users/killion/final_means.csv", row.names = FALSE)
+write.csv(final_means, "/Users/killion/final_means2.csv", row.names = FALSE)
 

### 2020-07-16 Rcode/For_Clark.Rmd
update

@@
 New data mean in Parks 
-L71 will need to be unhashed when working with our sample data - it selects only the months which have data in folder
+L71 will need to be hashed out when working with full data - it selects only the months which have data in our sample folder
 L93 values need to change based on the length of your new filepath
@@
 
-# list_all<- list_all[c(1,4)] You will need to use this with our sample data folder
+list_all<- list_all[c(1,4)] #You don't need this for final run with all data
 

### 2020-07-16 Rcode/For_Clark.Rmd
update2

@@
 
-New data mean in Parks 
+Mean in Parks w/ New Data
 L71 will need to be hashed out when working with full data - it selects only the months which have data in our sample folder

### 2020-07-16 Rcode/SmokeAggregationPolys.r
code for aggregating MERRA2 smoke files

@@ -1,0 +1,118 @@
+library(ncdf4)
+library(chron)
+library(lattice)
+library(RColorBrewer)
+
+library(raster)
+library(rgdal)
+library(exactextractr)
+library(readr)
+library(plyr)
+library(purrr)
+
+
+#Data From - https://public-nps.opendata.arcgis.com/datasets/national-park-service-park-unit-boundaries
+
+
+nps<-readOGR("~/SmokeProject/NPS_poly/NPS_poly.shp")
+
+#select only polygons designated as National Parks
+nps<-nps[nps@data$UNIT_TYPE=="National Park",]
+
+#remove National Parks outside our Regions
+nps<-nps[nps@data$REGION=="PW"|nps@data$REGION=="IM",]
+nps<-nps[nps@data$UNIT_CODE!="NPSA",] #Samoa
+nps<-nps[nps@data$STATE!="HI",]
+
+#reformat to use park codes later
+nps_df<-as.data.frame(nps)
+nps_df<-as.matrix(nps_df$UNIT_CODE)
+
+park_codes<-as.data.frame(nps_df)
+names(park_codes)<-"park"
+
+
+#Mean in Parks w/ New Data
+#L71 will need to be hashed out when working with full data - it selects only the months which have data in our sample folder
+#L93 values need to change based on the length of your new filepath
+
+poly<-sf::st_as_sfc(nps) #converts to new type of poly
+
+
+fileNames<-list.files("~/SmokeProject/MERRA_DATA_ORG_PRACT",full.names = TRUE)
+
+years<-c(1980,2004)  #all the years
+months<-c("01","10") #all the months
+comb<-expand.grid(years,months) #just to get the number of unique combinations
+list_all<-list() #create an empty list to fill with the file names
+
+for (i in 1:nrow(comb)){ #creates the number of list objects
+  year<-comb[i,]$Var1
+  month<-comb[i,]$Var2
+  x<-(paste0(year,month))
+  list_all[[i]]<-fileNames[grep(x,fileNames)]  #fill the list with the file names
+  names(list_all)[i]<-paste0("x",year,month) #name the list objects
+}
+
+list_all<- list_all[c(1,4)] #You don't need this for final run with all data
+
+empty_list<-list()
+
+for (i in 1:length (list_all)) {
+  for (j in 1:length (list_all[[i]])){
[…]

### 2020-07-17 Rcode/SmokeAggregationPolys.r
code used to et the final means

@@ -39,10 +39,10 @@ names(park_codes)<-"park"
 poly<-sf::st_as_sfc(nps) #converts to new type of poly
 
 
-fileNames<-list.files("~/SmokeProject/MERRA_DATA_ORG_PRACT",full.names = TRUE)
+fileNames<-list.files("/Users/Matthewclark989/Downloads/MERRA_all",full.names = TRUE)
 
-years<-c(1980,2004)  #all the years
-months<-c("01","10") #all the months
+years<-c(1980:2019)  #all the years
+months<-c("01","02","03","04","05","06","07","08","09","10","11","12") #all the months
 comb<-expand.grid(years,months) #just to get the number of unique combinations
 list_all<-list() #create an empty list to fill with the file names
 
@@ -54,7 +54,7 @@ for (i in 1:nrow(comb)){ #creates the number of list objects
   names(list_all)[i]<-paste0("x",year,month) #name the list objects
 }
 
-list_all<- list_all[c(1,4)] #You don't need this for final run with all data
+
 
 empty_list<-list()
 
@@ -76,7 +76,7 @@ for (i in 1:length (list_all)) {
   
   
   #write csv
-  n<-substr(list_all[[i]][[j]],97,103) # !!!!!!!! you will have to change 94 and 100 based on the length of your filepaths - this selects just the year and month i.e. 20041008
+  n<-substr(list_all[[i]][[j]],71,78) # !!!!!!!! you will have to change 94 and 100 based on the length of your filepaths - this selects just the year and month i.e. 20041008
   colnames(out_mean)[1] <- n #give values column name based on month
   name<-paste0("/Users/Matthewclark989/Documents/output_means/",n) #create empty folder to save to
   newname<-paste0(name, ".csv")

### 2020-07-22 Rcode/For_Clark_MAX.Rmd
Max Code

--- 2020-07-16 Rcode/For_Clark.Rmd
@@
 
-New data mean in Parks 
+Max in Parks w/ New Data (average within day - get max day for each month)
 L71 will need to be hashed out when working with full data - it selects only the months which have data in our sample folder
@@
 ```{r}
-
+#where to get max function
+install.packages("remotes")
+remotes::install_github("ejanalysis/analyze.stuff", force=TRUE)
+library(analyze.stuff)
+library(exactextractr)
 poly<-sf::st_as_sfc(nps) #converts to new type of poly
@@
 out<-as.data.frame(empty_list)
-out_mean<-as.data.frame(rowMeans(out)) #take mean of the assign list above
+out_max<-as.data.frame(rowMaxs(out)) #take mean of the assign list above
 
@@
   n<-substr(list_all[[i]][[j]],94,100) # !!!!!!!! you will have to change 94 and 100 based on the length of your filepaths - this selects just the year and month i.e. 20041008
-  colnames(out_mean)[1] <- n #give values column name based on month
-   name<-paste0("/Users/killion/output_means/",n) #create empty folder to save to
+  colnames(out_max)[1] <- n #give values column name based on month
+   name<-paste0("/Users/killion/output_max/",n) #create empty folder to save to
     newname<-paste0(name, ".csv")
-write.csv(out_mean, newname, row.names = FALSE) #save as csv
+write.csv(out_max, newname, row.names = FALSE) #save as csv
 
@@
 #loads those .csv and makes into single df
-list_dates = list.files(path="/Users/killion/output_means/", pattern="*.csv", full.names=TRUE) #same folder you saved above output
+list_dates = list.files(path="/Users/killion/output_max/", pattern="*.csv", full.names=TRUE) #same folder you saved above output
 
-final_means<-list_dates %>%
+library(tidyverse)
+final_max<-list_dates %>%
   map(read_csv) %>%
@@
 #append park code
-final_means<-data.frame(park=park_codes, final_means)
+final_max<-data.frame(park=park_codes, final_max)
 
 #save
-write.csv(final_means, "/Users/killion/final_means2.csv", row.names = FALSE)
+write.csv(final_max, "/Users/killion/final_max.csv", row.names = FALSE)
 

### 2020-07-22 Rcode/DataMergeClean.R
merging the new vis and smoke data

@@ -1,0 +1,50 @@
+mean_s_dat<-read.csv("~/SmokeProject/Data/final_means2.csv")
+Visdat<-read.csv("~/SmokeProject/Data/VisitationData.csv")
+mean_s_dat<-tidyr::gather(mean_s_dat,key="Date",value="Smoke",X19800131:X20191231)
+
+mean_s_dat$Year<-as.integer(substr(mean_s_dat$Date,2,5))
+mean_s_dat$Month<-as.integer(substr(mean_s_dat$Date,6,7))
+mean_s_dat$UnitCode<-mean_s_dat$park
+mean_s_dat<-mean_s_dat[,3:6]
+dat<-merge(Visdat,mean_s_dat,by=c("Year","Month","UnitCode"))
+dat$RecreationVisits<-as.numeric(dat$RecreationVisits)
+dat$stdsmoke<-as.vector(scale(dat$Smoke))
+
+#add max smoke
+max_s_dat<-read.csv("~/SmokeProject/Data/final_max.csv")
+max_s_dat<-tidyr::gather(max_s_dat,key="Date",value="Smoke",X19800131:X20191231)
+max_s_dat$Year<-as.integer(substr(max_s_dat$Date,2,5))
+max_s_dat$Month<-as.integer(substr(max_s_dat$Date,6,7))
+max_s_dat$UnitCode<-max_s_dat$park
+max_s_dat$MAXSmoke<-max_s_dat$Smoke
+max_s_dat<-max_s_dat[,4:7]
+dat<-merge(dat,max_s_dat,by=c("Year","Month","UnitCode"))
+dat$stdsmokeMAX<-as.vector(scale(dat$MAXSmoke))
+
+#Create a season variable
+dat$Season<-ifelse(dat$Month %in% c(3:5),"Spring",
+                    ifelse(dat$Month %in% c(6:8),"Summer",
+                           ifelse(dat$Month %in% c(9:11),"Fall","Winter")))
+
+#make combined variables for park/season and park/month
+dat$CatColS<-paste0(dat$UnitCode,dat$Season)
+
+dat$CatColM<-paste0(dat$UnitCode,dat$Month)
+
+
+#add vis season
+x<-dat%>%group_by(UnitCode,Season)%>%
+  summarise(mv=mean(RecreationVisits))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(mv==max(mv))%>%mutate(cc=paste0(UnitCode,Season))
+
[…]

### 2020-07-22 Rcode/DiferencefromTrend.R

@@ -1,11 +1,13 @@
 library(tidyverse)
 library(lme4)
 library(rstan)
-Data<-read_csv("Data/MergedDataComplete.csv")[,-1]
-Data$date<-zoo::as.yearmon(paste0(Data$Month,Data$Year),"%m%Y")
-Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
-Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
-Data<-na.omit(Data)
+Data<-read_csv("Data/MergedDataCompleteFINAL.csv")[,-1]
+Data[order(Data$UnitCode, Data$Date),]
+Data$Date<-zoo::as.yearmon(paste(Data$Year, Data$Month), "%Y %m")
+Data<-Data[order(Data$UnitCode, Data$Date),]
+#Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
+#Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
+#Data<-na.omit(Data)
 #add smoke season
 #x<-Data%>%group_by(UnitCode,Season)%>%
  # summarise(ms=mean(Smoke))%>%ungroup()%>%
@@ -20,43 +22,52 @@ Data<-na.omit(Data)
 Data<-Data%>%filter( SeasType =="High")
 
 #filter by smoke affected regions
-Data<-Data%>%
-  filter(Region %in% c("Intermountain","Pacific West"))
+#Data<-Data%>%
+ # filter(Region %in% c("Intermountain","Pacific West"))
 
 
 #get vis and smoke trends
-fits <- lmList(Smoke ~ date | UnitCode, data=Data) 
-Data$trendsmoke<-predict(fits,date=allData$date)
+fits <- lmList(Smoke ~ Date | UnitCode, data=Data) 
+Data$trendsmoke<-predict(fits,date=Data$Date)
 
-fits <- lmList(RecreationVisits ~ date | UnitCode, data=Data) 
-Data$trendvis<-predict(fits,date=Data$date)
+fits <- lmList(RecreationVisits ~ Date | UnitCode, data=Data) 
+Data$trendvis<-predict(fits,date=Data$Date)
 
 Data$SmokeDiff<-scale(Data$Smoke-Data$trendsmoke)
[…]

### 2020-07-22 Rcode/VidDataCleaning.R

@@ -1,12 +1,0 @@
-library(tidyverse)
-VisDat<-read_csv("~/Smoke_Proj/Data/VisDat_NoHeader.csv")
-View(VisDat)
-
-#this data is massive.
-#Let's chop it down a bit
-names(VisDat)
-VisDat<-VisDat %>% 
-  select(ParkName,UnitCode,Year,Month,RecreationVisits)
-
-write.csv(VisDat, file="~/Smoke_Proj/Data/VisitationDataClean.csv")
-nrow(VisDat)

### 2020-07-23 Rcode/DiferencefromTrend.R
residual break point model

@@ -2,22 +2,21 @@ library(tidyverse)
 library(lme4)
 library(rstan)
 Data<-read_csv("Data/MergedDataCompleteFINAL.csv")[,-1]
-Data[order(Data$UnitCode, Data$Date),]
 Data$Date<-zoo::as.yearmon(paste(Data$Year, Data$Month), "%Y %m")
 Data<-Data[order(Data$UnitCode, Data$Date),]
 #Data<-Data%>%group_by(UnitCode,Month)%>%mutate(ARVis = lag(RecreationVisits) )
 #Data$VisChange<-(Data$RecreationVisits-Data$ARVis)/Data$ARVis
 #Data<-na.omit(Data)
 #add smoke season
-#x<-Data%>%group_by(UnitCode,Season)%>%
- # summarise(ms=mean(Smoke))%>%ungroup()%>%
-  #group_by(UnitCode)%>%filter(ms==max(ms))%>%mutate(cc=paste0(UnitCode,Season))
+x<-Data%>%group_by(UnitCode,Season)%>%
+  summarise(ms=mean(Smoke))%>%ungroup()%>%
+  group_by(UnitCode)%>%filter(ms==max(ms))%>%mutate(cc=paste0(UnitCode,Season))
 
 
-#Data$SmokeSeas<-ifelse(Data$CatColS %in% x$cc,
- #                      "High","NotHigh")
+Data$SmokeSeas<-ifelse(Data$CatColS %in% x$cc,
+                       "High","NotHigh")
 
-#Data<-Data%>%filter( SmokeSeas =="High")
+Data<-Data%>%filter( SmokeSeas =="High")
 #filter by high season
 Data<-Data%>%filter( SeasType =="High")
 
@@ -49,30 +48,54 @@ Data%>%filter(UnitCode=="YOSE")%>%
 
 
 
-Data%>%
-  ggplot(.,aes(x=Smoke,y=VisDiff))+
-  geom_point(aes(color=UnitCode))+
-  geom_smooth(aes(color=UnitCode),se=F,method = "lm")+theme_classic()
+
 
 Data%>%
   ggplot(.,aes(x=Smoke,y=VisDiff))+
   geom_point()+
-  geom_smooth(se=F)+theme_classic()+facet_wrap(~UnitCode,scales="free")
+  geom_smooth(se=F,method="lm")+theme_classic()+facet_wrap(~UnitCode,scales="free")+
+  ggtitle("mean smoke")
 
[…]

### 2020-07-23 Rcode/Residuals_Break_point.R

@@ -1,0 +1,39 @@
+library(tidyverse)
+library(rstan)
+dat<-read_csv("Data/MergedDataCompleteFINAL.csv")[,-1]
+dat$Date<-zoo::as.yearmon(paste(dat$Year, dat$Month), "%Y %m")
+dat<-dat[order(dat$UnitCode, dat$Date),]
+dat<-dat%>%filter( SeasType =="High")
+fits <- lmList(RecreationVisits ~ Date | UnitCode, data=dat) 
+dat$trendvis<-predict(fits,date=dat$Date)
+dat$VisDiff<-dat$RecreationVisits-dat$trendvis
+dat%>%filter(UnitCode=="YOSE")%>%
+  ggplot(.,aes(x=Date,y=RecreationVisits))+geom_point()+geom_smooth(se=F,method = "lm")+
+  geom_point(aes(x=Date,y=trendvis),color="green")+ggtitle("Yosemite NP")
+
+
+
+
+
+data_list <- list(
+  N = nrow(dat),
+  Nprk = length(unique(dat$UnitCode)),
+  vis = dat$VisDiff,
+  smoke = dat$stdsmoke,
+  pcode = as.numeric(as.factor(dat$UnitCode )))
+
+options(mc.cores=8)
+
+
+l<-stan( file="Residuals_Break_Point.stan" , data=data_list,chains=8,iter=5000,control(adapt_delta=0.99) ,warmup = 800 )
+
+print( l , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
+
+
+
+
+
+
+
+
+

### 2020-07-23 Rcode/Residuals_Break_point.R
trend vis as a vector

@@ -6,7 +6,7 @@ dat<-dat[order(dat$UnitCode, dat$Date),]
 dat<-dat%>%filter( SeasType =="High")
 fits <- lmList(RecreationVisits ~ Date | UnitCode, data=dat) 
 dat$trendvis<-predict(fits,date=dat$Date)
-dat$VisDiff<-dat$RecreationVisits-dat$trendvis
+dat$VisDiff<-as.vector(dat$RecreationVisits-dat$trendvis)
 dat%>%filter(UnitCode=="YOSE")%>%
   ggplot(.,aes(x=Date,y=RecreationVisits))+geom_point()+geom_smooth(se=F,method = "lm")+
   geom_point(aes(x=Date,y=trendvis),color="green")+ggtitle("Yosemite NP")

### 2020-07-23 Rcode/Residuals_Break_point.R
slight change to adapt_delta secification

@@ -25,7 +25,7 @@ data_list <- list(
 options(mc.cores=8)
 
 
-l<-stan( file="Residuals_Break_Point.stan" , data=data_list,chains=8,iter=5000,control(adapt_delta=0.99) ,warmup = 800 )
+l<-stan( file="Residuals_Break_Point.stan" , data=data_list,chains=8,iter=5000,control=list(adapt_delta=0.99) ,warmup = 800 )
 
 print( l , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
 

### 2020-07-23 Rcode/DiferencefromTrend.R
model for hes bowl comp

@@ -77,7 +77,7 @@ Data%>%
 Data%>%
   ggplot(.,aes(x=Smoke,y=RecreationVisits))+
   geom_point()+
-  geom_smooth(se=F,method="lm")+theme_classic()+facet_wrap(~UnitCode,scales="free")
+  geom_smooth(se=F)+theme_classic()+facet_wrap(~UnitCode,scales="free")
 
 
 Data%>%

### 2020-07-23 Rcode/Residuals_Break_point.R

@@ -29,7 +29,9 @@ l<-stan( file="Residuals_Break_Point.stan" , data=data_list,chains=8,iter=5000,c
 
 print( l , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
 
+options(mc.cores=3)
 
+zz<-stan( file="~/SmokeProject/StanCode/Residuals_FIXED_BP.stan" , data=data_list,chains=3 )
 
 
 

### 2020-12-04 Markdown Documents/Manuscript.Rmd

@@
 ### 2.1.1. Visitor Data
+
+Visitation data was obtained through the [National Park Service Visitor Use Statistics Portal](https://irma.nps.gov/STATS/).
+These data are generally collected via car counters, permit information, and concessionaire reporting.
+Methods of collection vary from park to park.
+Unit specific information can be found on the National Park Service Visitor Use Statistics Portal.
+Although collection methods are inconsistent across the sample, all data was taken as reported.
+We retrieved monthly data for all national park units in the Intermountain and Pacific West regions for years 1980 through 2019.
+To reduce issues regarding park seasonality, we then subset our data to only the three high season months associated with each park.
+For example, only Summer (June, July, Aug) data was used for Yellowstone National Park.
 

### 2020-12-06 Rcode/AQI_Alex.Rmd
AQI

---
title: "AQI"
author: "Alex Killion"
date: "12/4/2020"
output: html_document
---

```{r}
temp <- list.files("/Users/killion/Documents/GitHub/SmokeProject/Data/AQI", pattern="*.csv", full.names = TRUE) #county AQI by day 1980-2020
myfiles <- lapply(temp, read.csv)

all <- do.call(rbind,myfiles) #create on dataframe

library(dplyr)
library(lubridate)

all$Date <- ymd(all$Date) #reformat date

all$State.Code<-as.numeric(all$State.Code)
all$State.Code<-sprintf("%02d", all$State.Code)#convert state code to 2 digit to match .shp county file
all$County.Code<-sprintf("%03d", all$County.Code)#convert county code to 3 digit to match .shp county file
all$CODE<-paste0(all$State.Code, all$County.Code) #merge two codes to match to spatial data

median<-all %>% #get monthly means by county
  group_by(County=CODE, Year=year(Date), Month=month(Date)) %>%
  summarise(Monthly_Med=median(AQI))

library(sf)
library(sp)
library(rgeos)
#intersect parks with counties to get county ID for park
counties<-read_sf("/Users/killion/Documents/GitHub/SmokeProject/GIS/Counties/cb_2018_us_county_5m.shp")
parks<-read_sf("/Users/killion/Documents/GitHub/SmokeProject/GIS/NPS_poly/NPS_poly.shp")

parks_sp<-as_Spatial(parks) #convert filetype
parks2 <- gBuffer(parks_sp, byid=TRUE, width=0)#add buffer so polygons don't intersect self

parks3<-st_as_sf(parks2)#back convert file

counties<-st_transform(counties, crs=st_crs(parks)) #match projection

join<-st_join(parks3, counties, largest=TRUE) #add largest county id to each park

join_df<-as.data.frame(join)

#select only polygons designated as National Parks
join_df<-join_df[join_df$UNIT_TYPE=="National Park",]

#remove National Parks outside our Regions
join_df<-join_df[join_df$REGION=="PW"|join_df$REGION=="IM",]
join_df<-join_df[join_df$UNIT_CODE!="NPSA",] #Samoa
join_df<-join_df[join_df$STATE!="HI",]

join_df$County<-paste0(join_df$STATEFP, join_df$COUNTYFP) #create merged code to match AQI data

#keep only columns needed
parks_county<-join_df[,c(2,26)] #keep only park name and county/state ID

#join dataframes

final<-left_join(parks_county, median, by=c("County" = "County"))
write.csv(final, "/Users/killion/Documents/GitHub/SmokeProject/Data/AQI_Final_Data.csv")

```

### 2020-12-07 Rcode/RUN_ARmod.R
AR Code

@@ -50,15 +50,16 @@ data_list <- list(
   pcode = as.numeric(as.factor(dat$UnitCode )),
   smoke = stdize(dat$medSmoke))
 
-options(mc.cores=6)
+options(mc.cores=2)
 
 rstan::rstan_options(autowrite=TRUE)
 
 #run the mod
-mod<-rstan::stan( file="~/SmokeProject/StanCode/AR_Only_RNdm_sl2.stan " , 
-                        data=data_list,chains=6,iter=3000,
+
+mod<-rstan::stan( file="~/SmokeProject/StanCode/ARMod_Simple.stan " , 
+                        data=data_list,chains=1,iter=3000,
                         control=list(adapt_delta=0.95,max_treedepth = 10),
-                  refresh= max(3000/20, 1),save_warmup=T)
+                  refresh= max(300/20, 1),save_warmup=T)
 
 
 print( mod , probs=c( (1-0.89)/2 , 1-(1-0.89)/2 ) )
@@ -151,7 +152,8 @@ mod_BP_simple<-rstan::stan( file="~/SmokeProject/StanCode/ARMod_Simple.stan " ,
 dat%>%filter(UnitCode=="ROMO")%>%
   ggplot(., aes(x=stdize(medSmoke),y=RecreationVisits))+geom_point()+geom_smooth()
 
-
+dat%>%filter(UnitCode=="PINN")%>%
+  ggplot(., aes(x=stdize(Year),y=RecreationVisits))+geom_point()+geom_smooth()
 
 
 

### 2020-12-07 Rcode/For_Clark_Median.Rmd

--- 2020-07-16 Rcode/For_Clark.Rmd
@@
 library(purrr)
+library(matrixStats)
 ```
@@
 
+#mean of all pixels in park for each day
 empty_list[[j]]<-exact_extract(all, poly, 'mean') #intersects boundary
@@
 out<-as.data.frame(empty_list)
-out_mean<-as.data.frame(rowMeans(out)) #take mean of the assign list above
+out_median<-as.data.frame(rowMedians(as.matrix(out))) #median of mean day value
 
@@
   n<-substr(list_all[[i]][[j]],94,100) # !!!!!!!! you will have to change 94 and 100 based on the length of your filepaths - this selects just the year and month i.e. 20041008
-  colnames(out_mean)[1] <- n #give values column name based on month
+  colnames(out_median)[1] <- n #give values column name based on month
    name<-paste0("/Users/killion/output_means/",n) #create empty folder to save to
     newname<-paste0(name, ".csv")
-write.csv(out_mean, newname, row.names = FALSE) #save as csv
+write.csv(out_median, newname, row.names = FALSE) #save as csv
 
@@
 
-final_means<-list_dates %>%
+final_medians<-list_dates %>%
   map(read_csv) %>%
@@
 #append park code
-final_means<-data.frame(park=park_codes, final_means)
+final_medians<-data.frame(park=park_codes, final_medians)
 
 #save
-write.csv(final_means, "/Users/killion/final_means2.csv", row.names = FALSE)
+write.csv(final_medians, "/Users/killion/final_medians2.csv", row.names = FALSE)
 

### 2020-12-07 Rcode/Validation_code.Rmd

--- 2020-07-16 Rcode/For_Clark.Rmd
@@
 ---
-title: "For_Clark"
+title: "Untitled"
 author: "Alex Killion"
-date: "7/16/2020"
+date: "7/20/2020"
 output: html_document
 ---
-
-Packages
-```{r}
-library(ncdf4)
-library(chron)
-library(lattice)
-library(RColorBrewer)
-
-library(raster)
-library(rgdal)
-library(exactextractr)
-library(readr)
-library(plyr)
-library(purrr)
-```
 
@@
 
-New data mean in Parks 
-L71 will need to be hashed out when working with full data - it selects only the months which have data in our sample folder
-L93 values need to change based on the length of your new filepath
 ```{r}
@@
 
-fileNames<-list.files("/Users/killion/Documents/GitHub/SmokeProject/MERRA_DATA_ORG_PRACT",full.names = TRUE)
-
-years<-c(1980,2004)  #all the years
-months<-c("01","10") #all the months
-comb<-expand.grid(years,months) #just to get the number of unique combinations
-list_all<-list() #create an empty list to fill with the file names
-
-for (i in 1:nrow(comb)){ #creates the number of list objects
-  year<-comb[i,]$Var1
-  month<-comb[i,]$Var2
-      x<-(paste0(year,month))
-      list_all[[i]]<-fileNames[grep(x,fileNames)]  #fill the list with the file names
-      names(list_all)[i]<-paste0("x",year,month) #name the list objects
-}
-
-list_all<- list_all[c(1,4)] #You don't need this for final run with all data
+list_all<- list.files("/Users/killion/Downloads/Random_Month",full.names = TRUE)
 
@@
 out<-as.data.frame(empty_list)
-out_mean<-as.data.frame(rowMeans(out)) #take mean of the assign list above
+out_mean_random<-as.data.frame(rowMeans(out, na.rm=TRUE)) #take mean of the assign list above
 

### 2021-03-11 MarkdownDocuments/Manuscript_Real.Rmd
manuscript formatted well

---
title: Estimating the Financial Impact of Wildfire Smoke on US National parks
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: University of Michigan
  name: Alexander Killion
- affiliation: Boise State University
  name: Matthew A. Williamson
- affiliation: Boise State University
  name: Vicken Hillis
- affiliation: Boise State University
  name: Laura? 
date: "`r format(Sys.time(), '%B %d, %Y')`"
keywords: Wildfire, Smoke, Recreation, Park Science
bibliography: ../SmokePaper.bib
abstract: abstract text
fontsize: 12pt
link-citations: yes
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex

---

# Introduction
Humanity must urgently adapt to our rapidly changing and increasingly hazardous climate [@IPCC]. Avoiding the worst-case projections of climate change will require adaptations at all levels of human society, from multinational governments to individuals [CITE?]. Governments and individuals are aware of this, but still are failing to act [@steg_limiting_2018]. Knowledge of adaptive strategies and their benefits is only weakly associated with actual adaptive behavior [@van_valkengoed_meta-analyses_2019]. While governments tend to implement adaptation policies in the wake of extreme events, less is known about how individuals are changing their day to day behaviors to adapt to their changing climates [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].

A particularly salient environmental effect of climate change in the United States is the increase in wildfire frequency and severity over the last 50 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016]. This trend is expected to worsen over the next 50 years as climate change accounts for approximately 50% of the increase in burned area [@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015]. Along with an increase in wildfires in the West, is a projected increase in wildfire smoke [@schroeder_validation_2008]. Fire prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].

Wildfire smoke is a mixture of many gasses aerated by biomass combustion across a landscape [@naeher_woodsmoke_2007]. While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter (PM2.5) is particularly harmful to human health [@deryugina_mortality_2019]. PM2.5 Causes between 260,000 - 600,000 global deaths annually and significantly increasing community mortality even under acute exposure [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013].

Wildfires now account for 50% of all PM2.5 in some western regions compared to less than 20% in 2010 [@burke_changing_2021]. As any level of ambient wildfire smoke increases community morbidity, public health agencies increasingly call for individuals to limit outdoor recreational behavior when wildfire smoke is present [@cascio_wildland_2018; @reid_colleen_e_critical_2016; @oneill_summary_2013].

Recent literature reviews on recreational behavior have found a dearth of research describing how individuals respond to low air quality, as well as research which gives predictions for how climate change will affect the recreation landscape [@otoole_climate_2019;@zajchowski_air_2019-1]. Additionally, a survey of federal land managers indicated that decision makers in areas such as national parks feel that they have a shortage of information describing how air quality affects the recreational behavior of their visitors [@zajchowski_air_2019]. It is of particular importance to understand how individuals respond to wildfire smoke because of the highly variable nature of wildfires, their relationship to climate change, and the great distances that smoke can travel during and after wildfire events [@dombeck_wildfire_2004; @collins_regional_2006].

This article uses ambient smoke and visitation data collected from national parks in the Western and Intermountain United States to answer whether individuals are adapting their recreation behavior in response to increasing wildfire smoke. The case study presented here aims to provide insights as to the overall trajectory of realized climate adaptation on the individual level.

\newpage

# References

### 2021-05-10 Rcode/PaperPlotsFinal.R
updates smoke scatterplot timeline

@@ -46,10 +46,18 @@ fit<-lm(formula = stdsmokemed ~ medSmoke, data = dat)
 stdd<-function(x){
   predict.lm(fit,newdata=data.frame(medSmoke=x))
 }
+
+HS<-dat%>%filter(SeasType=="High")%>%filter(stdsmokemed>5)%>%filter(UnitCode!="LAVO")%>%filter(UnitCode!="KICA")
+HSL<-dat%>%filter(SeasType=="High")%>%filter(stdsmokemed>5)%>%filter(UnitCode=="LAVO")
+HSK<-dat%>%filter(SeasType=="High")%>%filter(stdsmokemed>5)%>%filter(UnitCode=="KICA")
+
 smokePlt<-dat%>%filter(SeasType=="High")%>%
 ggplot(.,aes(x=Date,y=medSmoke,by=UnitCode))+
-  geom_point(alpha=0.4,size=4)+
-  theme_classic()+mytheme+ylab("Median Smoke Value")+
+  geom_point(alpha=0.4,size=3)+
+    geom_text(data=HS,aes(x=Date-2.1,y=medSmoke,label=paste(ParkName,paste0(Month,"/",Year))),size=3,fontface="bold")+
+    geom_text(data=HSK,aes(x=Date-2.7,y=medSmoke,label=paste(ParkName,paste0(Month,"/",Year))),size=3,fontface="bold")+
+    geom_text(data=HSL,aes(x=Date-2.9,y=medSmoke,label=paste(ParkName,paste0(Month,"/",Year))),size=3,fontface="bold")+
+    theme_classic()+mytheme+ylab("Median Smoke Value")+
 scale_y_continuous(sec.axis =sec_axis(trans = stdd ,name="Standardized"))
 
 #make bp example data

### 2021-05-11 MarkdownDocuments/Manuscript_Real.Rmd
tons of updates

@@
 ---
-title: Estimating the Financial Impact of Wildfire Smoke on US National parks
+title: Breakpoint Models Show No Evidence of Recreational Adaptation to Increasing Wildfire Smoke in the American West
 author:
@@
 keywords: Wildfire, Smoke, Recreation, Park Science
-bibliography: ../SmokePaper.bib
+bibliography: [../SmokePaper.bib, ../OtherRefs.bib]
 abstract: abstract text
@@
     template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
+header-includes:
+- \usepackage{pdfpages}
+- \usepackage[font={footnotesize}]{caption}
 
 ---
+```{r packages, include=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
+library(knitr)
+library(rmarkdown)
+library(bookdown)
+library(formattable)
+library(kableExtra)
+library(dplyr)
+library(magrittr)
+library(prettydoc)
+library(htmltools)
+library(knitcitations)
+library(devtools)
+library(tidyverse)
+library(scales)
+library(extrafont)
+```
+
+```{r setup, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
+# Chunk options: see http://yihui.name/knitr/options/ ###
+
+## Text results
+#opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)
+
+## Code decoration
+#opts_chunk$set(tidy = TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), highlight = TRUE)
+
+## Caching code
+knitr::opts_chunk$set(cache = 2, cache.path = "Cache/")
+
+## Plots
+#opts_chunk$set(fig.path = "Figures_MS/", dev=c('pdf', 'png'), dpi = 300)
+
+## Locate figures as close as possible to requested position (=code)
+knitr::opts_chunk$set(fig.pos = 'H')
+
+setwd("C:/Users/Matthewclark989/Documents/SmokeProject")
+```
+
+```{r plotzsrc, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
+source("~/SmokeProject/Rcode/PaperPlotsFinal.R")
+#source("~/SmokeProject/Rcode/PosteriorPlots.r")
+
+```
 
 # Introduction
-Humanity must urgently adapt to our rapidly changing and increasingly hazardous climate [@IPCC].
-Avoiding the worst-case projections of climate change will require adaptations at all levels of human society, from multinational governments to individuals [CITE?].
+Humanity must urgently adapt to our increasingly variable and hazardous climate [@IPCC].
+Avoiding the worst-case projections of climate change will require adaptations at all levels of human society, from multinational governments to individuals [@IPCC].
 Governments and individuals are aware of this, but still are failing to act [@steg_limiting_2018].
-Knowledge of adaptive strategies and their benefits is only weakly associated with actual adaptive behavior [@van_valkengoed_meta-analyses_2019].
-While governments tend to implement adaptation policies in the wake of extreme events, less is known about how individuals are changing their day to day behaviors to adapt to their changing climates [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].
+Knowledge of adaptive strategies and their benefits does not directly lead to adaptive behavior [@van_valkengoed_meta-analyses_2019].
+While governments tend to implement climate adaptation policies in the wake of extreme events, less is known about how individuals are reacting and possibly adapting to a changing climate [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].
 
-A particularly salient environmental effect of climate change in the United States is the increase in wildfire frequency and severity over the last 50 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016].
-This trend is expected to worsen over the next 50 years as climate change accounts for approximately 50% of the increase in burned area [@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
-Along with an increase in wildfires in the West, is a projected increase in wildfire smoke [@schroeder_validation_2008].
+A particularly salient environmental effect of climate change in the United States is the increasing wildfire frequency and severity over the last 40 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016].
+Because climate change accounts for approximately half of the increase in burned area during this time, fires are expected to worsen over the next 40 years as the climate continues to warm.
+[@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
+Accompanying the increase in wildfires in the West, is a projected increase in wildfire smoke [@schroeder_validation_2008].
 Fire prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].
 
-Wildfire smoke is a mixture of many gasses aerated by biomass combustion across a landscape [@naeher_woodsmoke_2007].
+![image alt text](../Figures/smkplt.png)
+
+Wildfire smoke is a complex mixture of gasses aerated by biomass combustion across the landscape [@naeher_woodsmoke_2007].
 While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter (PM2.5) is particularly harmful to human health [@deryugina_mortality_2019].
@@
 
-Wildfires now account for 50% of all PM2.5 in some western regions compared to less than 20% in 2010 [@burke_changing_2021].
-As any level of ambient wildfire smoke increases community morbidity, public health agencies increasingly call for individuals to limit outdoor recreational behavior when wildfire smoke is present [@cascio_wildland_2018; @reid_colleen_e_critical_2016; @oneill_summary_2013].
+Wildfires now account for 50% of all atmospheric PM2.5 in some Western regions compared to less than 20% in 2010 [@burke_changing_2021].
+As any level of ambient PM2.5 increases community morbidity, public health agencies increasingly call for individuals to limit outdoor recreational behavior when wildfire smoke is present [@cascio_wildland_2018; @reid_colleen_e_critical_2016; @oneill_summary_2013].
 
@@
 Additionally, a survey of federal land managers indicated that decision makers in areas such as national parks feel that they have a shortage of information describing how air quality affects the recreational behavior of their visitors [@zajchowski_air_2019].
-It is of particular importance to understand how individuals respond to wildfire smoke because of the highly variable nature of wildfires, their relationship to climate change, and the great distances that smoke can travel during and after wildfire events [@dombeck_wildfire_2004; @collins_regional_2006].
 
+It is of particular importance to understand how individuals are responding to wildfire smoke because it may provide a test case for how individuals will adapt to climate change generally.
 This article uses ambient smoke and visitation data collected from national parks in the Western and Intermountain United States to answer whether individuals are adapting their recreation behavior in response to increasing wildfire smoke.
+We use a hierarchical break-point modeling approach to test for linear and threshold effects of smoke on national park visitation.
 The case study presented here aims to provide insights as to the overall trajectory of realized climate adaptation on the individual level.
+
+# Methods
+## Data Collection
+
+### Visitation Data
+Visitation data was obtained through the [National Park Service Visitor Use Statistics Portal](https://irma.nps.gov/STATS/).
+These data are generally collected via car counters, permit information, and concessionaire reporting.
+Methods of collection vary from park to park.
+Unit specific information can be found on the National Park Service Visitor Use Statistics Portal.
+Although collection methods are inconsistent across the sample, all data was taken as reported.
+We retrieved monthly data for all national park units in the Intermountain and Pacific West regions for years 1980 through 2019.
+To reduce issues regarding park seasonality, we then subset our data to only the three month high season associated with each park.
+For example, only Summer (June, July, Aug) data was used for Yellowstone National Park.
+
+### SmokeData
+Alex to fill in this part
+
+## Analyses
+
+### Modeling
+Threshold effects are common in ecology and human behavioral sciences [@muradian2001ecological; @huggett2005concept; @macy1991chains].
+Toms & Lesperance [-@toms2003piecewise] show that breakpoint models are an effective tool to identify these thresholds and gather inference about the effects of predictors before and after in natural systems.
+Here we develop a breakpoint model to identify if a threshold exists for visitation response to ambient smoke and if so, what the effect of smoke is on visitation to national parks in the American West.
+
+![image alt text](../Figures/bpplot.png)
+
+$$f\]n =\textbackslash begin \{ cases \}a \&\textbackslash text \{ if \$n =0 \$\} \textbackslash \textbackslash r\textbackslash cdot f\] \{n -1\} \&\textbackslash text \{ else \}\textbackslash end \{ cases \$$
+
+### Validation
+Text about model checking
+
+# Results
+
+![image alt text](../Figures/slpz1.png)
+
+![image alt text](../Figures/slpz2.png)
+
+# Discussion
+We probably need national level policy to protect people [@rai_policy_2020]. There is huge variation in the way states react and plan for climate change [@hu_evaluating_2018].
+
+because of the highly variable nature of wildfires, their relationship to climate change, and the great distances that smoke can travel during and after wildfire events [@dombeck_wildfire_2004; @collins_regional_2006].
 

### 2021-05-11 MarkdownDocuments/Manuscript.Rmd
got the equations rendered

--- 2020-12-04 Markdown Documents/Manuscript.Rmd
@@
 fontsize: 11pt
+template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex
+
 geometry: margin=1in

### 2021-05-21 MarkdownDocuments/Manuscript_Real.Rmd
equations looking pretty good!

@@
 - \usepackage[font={footnotesize}]{caption}
+- \usepackage{bm}
+- \usepackage{amsmath}
 
@@
 Accompanying the increase in wildfires in the West, is a projected increase in wildfire smoke [@schroeder_validation_2008].
-Fire prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].
+Fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].
 
@@
 Wildfire smoke is a complex mixture of gasses aerated by biomass combustion across the landscape [@naeher_woodsmoke_2007].
-While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter (PM2.5) is particularly harmful to human health [@deryugina_mortality_2019].
-PM2.5 Causes between 260,000 - 600,000 global deaths annually and significantly increasing community mortality even under acute exposure [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013].
+While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter ($\text{PM}_{2.5}$) is particularly harmful to human health [@deryugina_mortality_2019].
+$\text{PM}_{2.5}$ Causes between 260,000 - 600,000 global deaths annually and significantly increasing community mortality even under acute exposure [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013].
 
-Wildfires now account for 50% of all atmospheric PM2.5 in some Western regions compared to less than 20% in 2010 [@burke_changing_2021].
-As any level of ambient PM2.5 increases community morbidity, public health agencies increasingly call for individuals to limit outdoor recreational behavior when wildfire smoke is present [@cascio_wildland_2018; @reid_colleen_e_critical_2016; @oneill_summary_2013].
+Wildfires now account for 50% of all atmospheric $\text{PM}_{2.5}$ in some Western regions compared to less than 20% in 2010 [@burke_changing_2021].
+As any level of ambient $\text{PM}_{2.5}$ increases community morbidity, public health agencies increasingly call for individuals to limit outdoor recreational behavior when wildfire smoke is present [@cascio_wildland_2018; @reid_colleen_e_critical_2016; @oneill_summary_2013].
 
@@
 It is of particular importance to understand how individuals are responding to wildfire smoke because it may provide a test case for how individuals will adapt to climate change generally.
-This article uses ambient smoke and visitation data collected from national parks in the Western and Intermountain United States to answer whether individuals are adapting their recreation behavior in response to increasing wildfire smoke.
-We use a hierarchical break-point modeling approach to test for linear and threshold effects of smoke on national park visitation.
+This article uses ambient smoke and visitation data collected from national parks in the Western and Intermountain United States from 1980-2019 to answer whether individuals are adapting their recreation behavior in response to increasing wildfire smoke.
+We hypothesize that the effect of ambient smoke on national park visitation may be nonlinear, only showing a measurable impact in the upper ranges of the observed smoke values.
+Threshold effects such as this are common in ecology and human behavioral sciences [@muradian2001ecological; @huggett2005concept; @macy1991chains].
+Toms & Lesperance [-@toms2003piecewise] show that breakpoint models are an effective tool to account for these thresholds and gather inference about the effects of predictors before and after in natural systems.
+Here we develop a series of three hierarchical breakpoint models to determine if threshold effects are present in national park visitation response to ambient smoke and if so, what the effect of smoke is on visitation to national parks in the American West.
 The case study presented here aims to provide insights as to the overall trajectory of realized climate adaptation on the individual level.
@@
 ### Modeling
-Threshold effects are common in ecology and human behavioral sciences [@muradian2001ecological; @huggett2005concept; @macy1991chains].
-Toms & Lesperance [-@toms2003piecewise] show that breakpoint models are an effective tool to identify these thresholds and gather inference about the effects of predictors before and after in natural systems.
-Here we develop a breakpoint model to identify if a threshold exists for visitation response to ambient smoke and if so, what the effect of smoke is on visitation to national parks in the American West.
+We first account for underlying change in national park visitation from 1980-2019 by developing a temporal autoregressive model which we validate and build upon to answer our main research question.
+This baseline autoregressive model uses the visitation data from the three previous years to predict the visitation to a given park in a given month.
+This formulation of an autoregressive model is commonly referred to as an AR(*k*) model, with this particular instance holding that *k* = 3 [@standev2021stancore].
+We chose a *k* value of 3 as the added within-sample predictive value of increasing *k* diminished considerably for any greater value.
+
+We formulated this baseline autoregressive model using a bayesian heierarchical framework.
+This approach is appropriate here as it allows us to account between park variation in visitation trends while acknowledging the interconnectedness of these trends through partial pooling [@statrethinkingbook].
+For each park ($j$) we estimate a parameter value ($\bm{\nu1...k}$) representing the trend in visitation from one, two, and three years prior ($t-1,2,...k$), to the month we are estimating ($i$).
+Finally, we estimate both global ($\alpha0$) and park specific ($\bm{\beta0}_{j}$) intercepts.
+We modeled these data using a negative binomial distribution to allow us to estimate an overdispersion paremeter ($\phi$) rather than assuming the dispersion is equal to the mean.
+
+$$ y_{ij} \sim NegBinom(\mu_{ij},\phi) $$
+
+\begin{equation}
+\text{log}(\mu_{ij}) = 
+\alpha0+\bm{\beta0}_{j}+\bm{\nu1}_{j}*\text{vis}_{i,t-1}+ \\
+\bm{\nu2}_{j}*\text{vis}_{i,t-2}+\bm{\nu3}_{j}*\text{vis}_{i,t-3}  
+\end{equation}
+
+$$\bm{\beta0}_{j} \sim Norm(0,\sigma_{j})$$
+
+teeeext
 
@@
 
-$$f\]n =\textbackslash begin \{ cases \}a \&\textbackslash text \{ if \$n =0 \$\} \textbackslash \textbackslash r\textbackslash cdot f\] \{n -1\} \&\textbackslash text \{ else \}\textbackslash end \{ cases \$$
+$$ y_{ij} \sim NegBinom(\mu_{ij},\phi) $$
+
+\begin{equation}
+\text{log}(\mu_{ij}) = 
+\begin{cases} 
+\alpha0+\bm{\beta0}_{j}+\bm{\beta1}_{j}*\text{smoke}_{i}+\bm{\nu1}_{j}*\text{vis}_{i,t-1}+ \\
+\bm{\nu2}_{j}*\text{vis}_{i,t-2}+\bm{\nu3}_{j}*\text{vis}_{i,t-3} & \text{if } \text{smoke}_{i} < \text{BP} \\ \\
+\beta0+\bm{\beta0}_{j}+\bm{\beta1}_{j}*\text{BP}+(\text{smoke}_{i}-\text{BP})*\bm{\beta2}_{j}+  \\ \bm{\nu1}_{j}*\text{vis}_{i,t-1}+\bm{\nu2}_{j}*\text{vis}_{i,t-2}+\bm{\nu3}_{j}*\text{vis}_{i,t-3}        & \text{if } \text{smoke}_{i} \geq \text{BP}
+\end{cases}
+\end{equation}
+
+$$\bm{\beta0}_{j} \sim Norm(0,\sigma_{j})$$
 

### 2021-05-22 MarkdownDocuments/Manuscript_Real.Rmd
updated plots in code and manuscript

@@
 
-![image alt text](../Figures/smkplt.png)
+![image alt text. Righ-hand y axis shows the standardized values of smoke ](../Figures/smkplt.png)
 
@@
 ### Visitation Data
-Visitation data was obtained through the [National Park Service Visitor Use Statistics Portal](https://irma.nps.gov/STATS/).
+Visitation data were obtained on the monthly scale through the [National Park Service Visitor Use Statistics Portal](https://irma.nps.gov/STATS/).
 These data are generally collected via car counters, permit information, and concessionaire reporting.
@@
 
-We formulated this baseline autoregressive model using a bayesian heierarchical framework.
+We formulated this baseline autoregressive model using a Bayesian hierarchical framework.
 This approach is appropriate here as it allows us to account between park variation in visitation trends while acknowledging the interconnectedness of these trends through partial pooling [@statrethinkingbook].
 For each park ($j$) we estimate a parameter value ($\bm{\nu1...k}$) representing the trend in visitation from one, two, and three years prior ($t-1,2,...k$), to the month we are estimating ($i$).
-Finally, we estimate both global ($\alpha0$) and park specific ($\bm{\beta0}_{j}$) intercepts.
-We modeled these data using a negative binomial distribution to allow us to estimate an overdispersion paremeter ($\phi$) rather than assuming the dispersion is equal to the mean.
+Finally, we estimate both global ($\alpha0$) and park specific ($\bm{\beta0}_{j}$) intercepts to yield equation 1 below.
+We modeled these data using a negative binomial distribution to allow us to estimate an overdispersion parameter ($\phi$) rather than assuming the dispersion is equal to the mean.
 
@@
 
-$$\bm{\beta0}_{j} \sim Norm(0,\sigma_{j})$$
+$$\bm{\beta0}_{j} \sim Norm(0,\sigma)$$
 
-teeeext
+After validating that the model above captures the baseline trends in national park visitation from 1980-2019, we built upon it to estimate the if a critical threshold for visitation response to wildfire smoke exists in these data and if so, what the impact of wildfire smoke is on visitation post-threshold.
+To do this, we developed a series of three breakpoint models, allowing for three increasingly extreme breakpoints (i.e.
+critical thresholds).
+We standardized these smoke data, and thus the thresholds as well, by dividing by two standard deviations following the recommendation made by Gelman [-@gelman2008scaling].
+A threshold of 1.0 can then be interpreted as two standard deviations above the mean value of smoke in the data.
+A visual comparison of the observed actual smoke values and standardized values can be seen in Fig.
+1.
 
-![image alt text](../Figures/bpplot.png)
+As shown conceptually in Fig.
+2, we estimated the effect of smoke on visitation before ($\bm{\beta1}$) and after ($\bm{\beta2}$) each breakpoint (BP).
+Just as with the autoregressive terms, we estimated unique paramater values for each park ($j$) in a heierarchical framework.
+The complete model used for each of the three breakpoint values is then as seen in equation 2.
+
+![image alt text](../Figures/bpplot2.png)
 
@@
 
-$$\bm{\beta0}_{j} \sim Norm(0,\sigma_{j})$$
+$$\bm{\beta0}_{j} \sim Norm(0,\sigma)$$
+
+Following the recommendation given by Gelman et al.
+[-@gelman2008weakly] for producing stable, conservative estimates, we specified long-tailed regularizing priors for all autoregressive terms, the global intercept value, and the overdispersion parameter.
+We used normally distributed regularizing priors for estimating the variance parameter ($\sigma$) and the pre and post-breakpoint smoke effects to allow for more efficient sampling while generative within-sample predictions at high smoke values.
 
 ### Validation
-Text about model checking
+Due to the considerable computational time (about 1 day) needed to run the program described above, we validated the results of both the autoregressive-only and breakpoint models using within-sample posterior predictive checks.
+All models were first inspected to ensure adequate mixing of Markov chains and $\hat{R}$ values below 1.05 for all parameter estimates.
+
+![image alt text](../Figures/ARVal.png)
 
 # Results
+While some between-park variation exists, our overall study shows no evidence for recreational visitor response to wildfire smoke in national parks in the American West.
 

### 2021-05-22 Rcode/FinalFullAnalysisCode.R

@@ -173,6 +173,6 @@ plot(y,preds)
 cor(y,preds)
 
 
-
-
-
+median(z$phi)
+median(z$Intercept)
+median(z$AR_term[30])

### 2021-05-22 Rcode/PaperPlotsFinal.R

@@ -1,5 +1,5 @@
 library(tidyverse)
-dat<-read_csv("Data/MergedDataCompleteFINAL.csv")[,-1]
+dat<-read_csv("~/SmokeProject/Data/MergedDataCompleteFINAL.csv")[,-1]
 #filter to only high season data
 #dat<-dat%>%filter(SeasType =="High")
 
@@ -96,8 +96,8 @@ bpdat<-gather(bpdat,key="bp",value="dep",dep0.5,dep1,dep1.5)
     scale_color_manual(values=cols2)+
     scale_y_continuous(breaks=c(0),labels="Intercept", limits = c(-0.5,0.3))+
     scale_x_continuous(breaks=c(0,0.5, 1.0, 1.5))+
-    geom_text(x=0,y=0.2,label="Slope 1",color="black",fontface="bold",size=6)+
-    geom_text(x=2.5,y=0.1,label="Slope 2",color="black",fontface="bold",size=6)+
+    geom_text(x=0,y=0.2,label="Slope~(beta)~1",color="black",fontface="bold",size=6,parse=TRUE)+
+    geom_text(x=2.5,y=0.1,label="Slope~(beta)~2",color="black",fontface="bold",size=6,parse=TRUE)+
     annotate("segment", x = 0, xend = 0, y = 0.175, yend = 0.12, colour = "black", size=1.5, arrow=arrow())+
     annotate("segment", x = 2.5, xend = 2.4, y = 0.075, yend = 0.02, colour = "black", size=1.5, arrow=arrow())+
     labs(color = "Breakpoint")+xlab("Smoke (standardized)")+ylab("Visitation")+
@@ -107,7 +107,30 @@ bpdat<-gather(bpdat,key="bp",value="dep",dep0.5,dep1,dep1.5)
     theme_classic()+mytheme+theme(legend.position = "none") 
   
   
+  #validation plot
+  load("~/SmokeProject/ModelObjects/modAROnly.rda")
+  y<-data_list$count #for pp checking in shinystan
+  #shinystan::launch_shinystan(mod)
+  z<-extract(mod)
+  preds<-apply(z$count_pred,2,median)
+  plot(y,preds)
+  cor(y,preds)
   
+  dat$PredVis<-preds
+  
+  dat$Date <- as.Date(paste(dat$Year, dat$Month, sep="-"), "%Y-%M")
+  
+  dat%>%filter(UnitCode =="GLAC")%>%
[…]

### 2021-05-22 Rcode/PosteriorPlots.r

@@ -2,9 +2,9 @@ library(tidyverse)
 library(rstan)
 library(ggthemes)
 library(ggridges)
-load("ModelObjects/modSmokeSet0_5.rda")#modSmoke2
-load("ModelObjects/modSmokeSet1.rda")#modSmoke3
-load("ModelObjects/modSmokeSet1_5.rda")#modSmoke4
+load("~/SmokeProject/ModelObjects/modSmokeSet0_5.rda")#modSmoke2
+load("~/SmokeProject/ModelObjects/modSmokeSet1.rda")#modSmoke3
+load("~/SmokeProject/ModelObjects/modSmokeSet1_5.rda")#modSmoke4
 
 #0.5
 z<-extract(modSmoke2)
@@ -49,17 +49,23 @@ slp1plot<-slpz%>%
   group_by(UnitCode,Model)%>%
   summarise(lower=quantile(Estimate,.25),
             upper=quantile(Estimate,.75),
+            top = quantile(Estimate,.95),
+            bottom = quantile(Estimate,.05),
             mid=quantile(Estimate,.5))%>%
   mutate(half=ifelse(UnitCode %in% 
                        c("JOTR","KICA","LAVO","MEVE","MORA","NOCA","OLYM","PEFO","PINN","REDW","ROMO","SAGU","SEQU","YELL","YOSE","ZION"),"2nd","1st"))%>%
   mutate(UnitCodeRev=fct_rev(UnitCode))%>%
-  ggplot(., aes(y = mid ,x=UnitCodeRev,ymin=lower,ymax=upper,color=Model))+ggtitle("A")+
+  ggplot(., aes(y = mid ,x=UnitCodeRev,ymin=lower,ymax=upper,color=Model))+ggtitle("")+#maybe put A/B here for panel lable
+  geom_linerange( mapping=aes(x=UnitCodeRev, ymin=bottom, ymax=top,color=Model), size=0.25,position = position_dodge(width = 0.5),alpha=0.8) +##change deets
   geom_pointrange(position = position_dodge(width = 0.5),size=1,alpha=0.8)+xlab("NPS Unit Code")+
-  scale_color_manual(name="Breakpoint (SD)",labels=c("1.5","1.0","1.5") ,values=cols2)+facet_wrap(~half,scales="free")+ylab("Standardized Effect Size")+
+  scale_color_manual(name="Breakpoint",labels=c("0.5","1.0","1.5") ,values=cols2)+facet_wrap(~half,scales="free")+ylab("Standardized Effect Size")+
[…]

### 2021-05-24 MarkdownDocuments/Manuscript_Real.Rmd
plots 'n writing

@@
 Due to the considerable computational time (about 1 day) needed to run the program described above, we validated the results of both the autoregressive-only and breakpoint models using within-sample posterior predictive checks.
-All models were first inspected to ensure adequate mixing of Markov chains and $\hat{R}$ values below 1.05 for all parameter estimates.
+All models were first inspected to ensure adequate mixing of Markov chains, $\hat{R}$ values below 1.05 for all parameter estimates, and the absence of divergent transitions after warm up.
+
+We examined the within-sample predictive capacity of the autoregressive only model to be sure we adequately accounted for baseline changes in visitation before building on this model to test for the effects of smoke on visitation.
+This baseline model allowed us to account for approximately 40% of all variation in the data, with the bulk of the error occurring when actual visitation was very high or very low (Fig 3).
 
@@
 While some between-park variation exists, our overall study shows no evidence for recreational visitor response to wildfire smoke in national parks in the American West.
+Figures 4 and 5 show the parameter estimates for the pre and post breakpoint smoke effects respectively.
+In all three breakpoint models, only two parks have 90% credible intervals that do not overlap zero for slope 1, both of which have opposite signs (Fig.
+4).
+This trend is paralleled for slope 2 in the two lower breakpoint models and all 90% credibility intervals overlap 0 for slope 2 in the model with the most extreme breakpoint (Fig.
+5).
+
+Intuitively, the within-sample $R^2$ increases only slightly when accounting for smoke effects, from 0.40 to 0.41 compared to the autoregressive only model.
+This suggests that ambiant smoke in national parks in the American West is not driving any measurable changes in visitation even at sustained and dangerous levels.
+We then conclude that visitors are not altering recreation behavior in response to smoke and therefore no critical threshold for adaptation is detectable in these data.
 
@@
 # Discussion
+We demonstrate that given the data available to us, we detect no visitor adaptation to increasing wildfire smoke in national parks in the American West.
+This result is troubling on both immediatly and in the long-term.
+We discuss earlier the public health dangers of wildfire smoke, and in particular $\text{PM}_{2.5}$.
+
 We probably need national level policy to protect people [@rai_policy_2020]. There is huge variation in the way states react and plan for climate change [@hu_evaluating_2018].

### 2021-05-24 Rcode/PaperPlotsFinal.R

@@ -132,5 +132,23 @@ bpdat<-gather(bpdat,key="bp",value="dep",dep0.5,dep1,dep1.5)
   
   
   
+  y<-data_list$count #for pp checking in shinystan
+  #shinystan::launch_shinystan(mod)
+  z<-extract(modSmoke2)
+  preds<-apply(z$count_pred,2,median)
+  plot(y,preds)
+  cor(y,preds)
   
+  dat$PredVis<-preds
+  
+  dat$Date <- as.Date(paste(dat$Year, dat$Month, sep="-"), "%Y-%M")
+  
+  dat%>%filter(UnitCode =="GLAC")%>%
+    ggplot(.,aes(x=Date))+geom_point(aes(y=RecreationVisits),color="blue")+geom_point(aes(y=PredVis),color="red")
+  
+  ggplot(dat,aes(x=RecreationVisits,y=PredVis))+
+    geom_abline(intercept = 0, slope = 1,linetype="dashed",color="#525252",size=1.5 )+
+    geom_point(alpha=0.5)+theme_classic()+mytheme+
+    ylab("Predicted Visitation")+xlab("Actual Visitation")+
+    geom_text(x=5000,y=20000,color="black",label=paste0("R^2 == ", round(cor(y,preds)^2,digits = 2)),parse=TRUE)
   

### 2021-05-24 Rcode/PosteriorPlots.r

@@ -58,7 +58,7 @@ slp1plot<-slpz%>%
   ggplot(., aes(y = mid ,x=UnitCodeRev,ymin=lower,ymax=upper,color=Model))+ggtitle("")+#maybe put A/B here for panel lable
   geom_linerange( mapping=aes(x=UnitCodeRev, ymin=bottom, ymax=top,color=Model), size=0.25,position = position_dodge(width = 0.5),alpha=0.8) +##change deets
   geom_pointrange(position = position_dodge(width = 0.5),size=1,alpha=0.8)+xlab("NPS Unit Code")+
-  scale_color_manual(name="Breakpoint",labels=c("0.5","1.0","1.5") ,values=cols2)+facet_wrap(~half,scales="free")+ylab("Standardized Effect Size")+
+  scale_color_manual(name="Breakpoint",labels=c("1.5","1.0","0.5") ,values=cols2)+facet_wrap(~half,scales="free")+ylab("Standardized Effect Size")+
   coord_flip()+geom_hline(yintercept=0,linetype="dashed",color="darkgrey" )+
   theme_classic()+mytheme+theme(strip.background = element_blank(),
                                              strip.text.x = element_blank())+
@@ -106,7 +106,7 @@ slp2plot<-slpz2%>%
   ggplot(., aes(y = mid ,x=UnitCodeRev,ymin=lower,ymax=upper,color=Model))+ggtitle("")+
   geom_linerange( mapping=aes(x=UnitCodeRev, ymin=bottom, ymax=top,color=Model), size=0.25,position = position_dodge(width = 0.5),alpha=0.8) +##change deets
   geom_pointrange(position = position_dodge(width = 0.5),size=1,alpha=0.8)+xlab("NPS Unit Code")+
-  scale_color_manual(name="Breakpoint",labels=c("0.5","1.0","1.5") ,values=cols2)+facet_wrap(~half,scales="free")+ylab("Standardized Effect Size")+
+  scale_color_manual(name="Breakpoint",labels=c("1.5","1.0","0.5") ,values=cols2)+facet_wrap(~half,scales="free")+ylab("Standardized Effect Size")+
   coord_flip()+geom_hline(yintercept=0,linetype="dashed",color="darkgrey" )+
   theme_classic()+mytheme+theme(strip.background = element_blank(),
                                 strip.text.x = element_blank())+

### 2021-05-24 MarkdownDocuments/Manuscript_Real.Rmd
text edits

@@
 This formulation of an autoregressive model is commonly referred to as an AR(*k*) model, with this particular instance holding that *k* = 3 [@standev2021stancore].
-We chose a *k* value of 3 as the added within-sample predictive value of increasing *k* diminished considerably for any greater value.
+We chose a *k* of 3 as the added within-sample predictive value of increasing *k* diminished considerably for any greater value.
 
@@
 
-After validating that the model above captures the baseline trends in national park visitation from 1980-2019, we built upon it to estimate the if a critical threshold for visitation response to wildfire smoke exists in these data and if so, what the impact of wildfire smoke is on visitation post-threshold.
+After validating that the model above captures the baseline trends in national park visitation from 1980-2019, we built upon it to estimate if a critical threshold for visitation response to wildfire smoke exists in these data and if so, what the impact of wildfire smoke is on visitation post-threshold.
 To do this, we developed a series of three breakpoint models, allowing for three increasingly extreme breakpoints (i.e.
@@
 2, we estimated the effect of smoke on visitation before ($\bm{\beta1}$) and after ($\bm{\beta2}$) each breakpoint (BP).
-Just as with the autoregressive terms, we estimated unique paramater values for each park ($j$) in a heierarchical framework.
+Just as with the autoregressive terms, we estimated unique parameter values for each park ($j$) in a hierarchical framework.
 The complete model used for each of the three breakpoint values is then as seen in equation 2.
@@
 Intuitively, the within-sample $R^2$ increases only slightly when accounting for smoke effects, from 0.40 to 0.41 compared to the autoregressive only model.
-This suggests that ambiant smoke in national parks in the American West is not driving any measurable changes in visitation even at sustained and dangerous levels.
+This suggests that ambient smoke in national parks in the American West is not driving any measurable changes in visitation even at sustained and dangerous levels.
 We then conclude that visitors are not altering recreation behavior in response to smoke and therefore no critical threshold for adaptation is detectable in these data.
@@
 We demonstrate that given the data available to us, we detect no visitor adaptation to increasing wildfire smoke in national parks in the American West.
-This result is troubling on both immediatly and in the long-term.
+This result is troubling both immediately and in the long-term.
 We discuss earlier the public health dangers of wildfire smoke, and in particular $\text{PM}_{2.5}$.
+These results suggest that..
 
@@
 
+A key limitation of this study is the lack of availability of fine scale visitation data to US national parks.
+Daily visitation data would allow us to identify a more complete picture of if and how visitors respond to smoke events day to day.
+Current data at the monthly scale allows us to investigate the big picture trends in visitor behavior, but future research would benefit greatly from daily count data.
+
+It is possible that individuals postpone visits to national parks by days or weeks when smoke is present, behavior which our study current study would be unable to capture.
+It is also possibly that this trend would not hold for recreation on all public lands.
+US national parks draw a great number non-local visitors, many of whom are coming from other states or countries and are likely visiting for the first time [@schuett2010visits].
+We speculate that these visitors are less likely to change their plans due to wildfire smoke than individuals recreating locally or repeatedly in one location.
+We propose that this unique feature of national parks makes their visitors particularly less adaptive to changing climatic conditions.
+Another path for future research might investigate if the trend shown here holds for other public lands with different visitor profiles.
+
 \newpage

### 2021-07-06 MarkdownDocuments/Manuscript_Real.Rmd
added abstract took another pass on intro

@@
 bibliography: [../SmokePaper.bib, ../OtherRefs.bib]
-abstract: abstract text
+abstract: Humans are going to have to adapt to our increasingly variable and hazardous climate.
+Rapidly increasing wildfire smoke in the American West represents a serious risk to human health and also a pertinent test-case for individual climate adaptation generally.
+Here we aggregate data on ambient smoke and recreational visitation in 32 U.S.
+national parks from 1980 - 2019 to test for observable adaptation to increasing wildfire smoke.
+We develop a series of breakpoint models to allow for threshold effects in recreational response.
+These models demonstrate that even when dangerous levels of smoke are present, individuals are failing to limit outdoor recreation.
+Future research may further our understanding of individual-level climate adaptation by focusing on alternative hazards or finer time series.
+
 fontsize: 12pt
@@
 # Introduction
-Humanity must urgently adapt to our increasingly variable and hazardous climate [@IPCC].
-Avoiding the worst-case projections of climate change will require adaptations at all levels of human society, from multinational governments to individuals [@IPCC].
-Governments and individuals are aware of this, but still are failing to act [@steg_limiting_2018].
-Knowledge of adaptive strategies and their benefits does not directly lead to adaptive behavior [@van_valkengoed_meta-analyses_2019].
-While governments tend to implement climate adaptation policies in the wake of extreme events, less is known about how individuals are reacting and possibly adapting to a changing climate [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].
+Climate change has already begun to dramatically change the social-ecological landscapes in which humans have learned to operate [@IPCC].
+All levels of human society must urgently adapt to these new realities if we are to continue to thrive over the coming millenia [@IPCC].
+As a society, we are aware of this, but still are failing to respond adequately [@steg_limiting_2018].
+Even when adaptive strategies and their benefits are known, we do not observe adaptive behaviors [@van_valkengoed_meta-analyses_2019].
+While research has demonstrated that extreme events tend to spur climate adaptation at the government level, less is known about how individuals are reacting and possibly adapting to a changing climate [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].
 
 A particularly salient environmental effect of climate change in the United States is the increasing wildfire frequency and severity over the last 40 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016].
-Because climate change accounts for approximately half of the increase in burned area during this time, fires are expected to worsen over the next 40 years as the climate continues to warm.
-[@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
-Accompanying the increase in wildfires in the West, is a projected increase in wildfire smoke [@schroeder_validation_2008].
+As climate change accounts for approximately half of the increase in burned area during this time, fires are expected to worsen over the next 40 years as the climate continues to warm [@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
+Accompanying the increase in wildfires in the American West, is a projected increase in wildfire smoke [@schroeder_validation_2008].
 Fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].
@@
 While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter ($\text{PM}_{2.5}$) is particularly harmful to human health [@deryugina_mortality_2019].
-$\text{PM}_{2.5}$ Causes between 260,000 - 600,000 global deaths annually and significantly increasing community mortality even under acute exposure [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013].
+$\text{PM}_{2.5}$ causes between 260,000 - 600,000 global deaths annually and significantly increases community mortality even under acute exposure [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013].
 
-Wildfires now account for 50% of all atmospheric $\text{PM}_{2.5}$ in some Western regions compared to less than 20% in 2010 [@burke_changing_2021].
+In some regions of the American West, wildfires now account for 50% of all atmospheric $\text{PM}_{2.5}$, compared to less than 20% in 2010 [@burke_changing_2021].
 As any level of ambient $\text{PM}_{2.5}$ increases community morbidity, public health agencies increasingly call for individuals to limit outdoor recreational behavior when wildfire smoke is present [@cascio_wildland_2018; @reid_colleen_e_critical_2016; @oneill_summary_2013].

### 2021-07-07 MarkdownDocuments/Manuscript_Real.Rmd
Finished first draft of the text. Next up is fig captions

@@
 bibliography: [../SmokePaper.bib, ../OtherRefs.bib]
-abstract: Humans are going to have to adapt to our increasingly variable and hazardous climate.
-Rapidly increasing wildfire smoke in the American West represents a serious risk to human health and also a pertinent test-case for individual climate adaptation generally.
+abstract: Humans must adapt to our increasingly variable and hazardous climate.
+Worsening wildfire smoke events in the American West represent a serious risk to human health and also a pertinent test-case for individual climate adaptation generally.
 Here we aggregate data on ambient smoke and recreational visitation in 32 U.S.
-national parks from 1980 - 2019 to test for observable adaptation to increasing wildfire smoke.
+national parks from 1980 - 2019 to test for observable adaptation to worsening wildfire smoke.
 We develop a series of breakpoint models to allow for threshold effects in recreational response.
@@
 ```{r plotzsrc, include=FALSE, cache=TRUE, message = FALSE, warning=FALSE}
-source("~/SmokeProject/Rcode/PaperPlotsFinal.R")
+#source("~/SmokeProject/Rcode/PaperPlotsFinal.R")
 #source("~/SmokeProject/Rcode/PosteriorPlots.r")
@@
 As a society, we are aware of this, but still are failing to respond adequately [@steg_limiting_2018].
-Even when adaptive strategies and their benefits are known, we do not observe adaptive behaviors [@van_valkengoed_meta-analyses_2019].
-While research has demonstrated that extreme events tend to spur climate adaptation at the government level, less is known about how individuals are reacting and possibly adapting to a changing climate [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].
+Even when adaptive strategies and their benefits are known, we do not demonstrate adaptive behaviors [@van_valkengoed_meta-analyses_2019].
+While research has shown that extreme events tend to spur climate adaptation at the government level, less is known about how individuals are reacting and possibly adapting to a changing climate [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].
 
@@
 # Methods
-## Data Collection
+## Data collection
 
-### Visitation Data
+### Visitation data
 Visitation data were obtained on the monthly scale through the [National Park Service Visitor Use Statistics Portal](https://irma.nps.gov/STATS/).
@@
 
-### SmokeData
+### Smoke data
 Alex to fill in this part
@@
 # Discussion
-We demonstrate that given the data available to us, we detect no visitor adaptation to increasing wildfire smoke in national parks in the American West.
+Given the data available to us, we detect no visitor adaptation to increasing wildfire smoke in national parks in the American West.
 This result is troubling both immediately and in the long-term.
-We discuss earlier the public health dangers of wildfire smoke, and in particular $\text{PM}_{2.5}$.
-These results suggest that..
+As discussed above, $\text{PM}_{2.5}$ significantly increases community morbidity even under acute exposure.
+In concert with showing no overall trends in behavioral adaptation, these data provide specific instances of historic smoke events where visitation did not deviate from normal.
 
-We probably need national level policy to protect people [@rai_policy_2020]. There is huge variation in the way states react and plan for climate change [@hu_evaluating_2018].
+The highly variable nature of wildfires, their relationship to climate change, and the great distances that smoke can travel during and after wildfire events make it difficult for individuals to plan around smoke events [@dombeck_wildfire_2004; @collins_regional_2006].
+In conjunction, U.S.
+national parks draw a great number of non-local visitors, many of whom are coming from other states or countries and are likely visiting for the first time [@schuett2010visits].
+We speculate that these visitors are less likely to change their plans due to wildfire smoke than individuals recreating locally or repeatedly in one location.
+We propose that these unique features of national parks and wildfire smoke make visitors particularly unable to adapt to changing climatic conditions.
 
-because of the highly variable nature of wildfires, their relationship to climate change, and the great distances that smoke can travel during and after wildfire events [@dombeck_wildfire_2004; @collins_regional_2006].
+Based on our findings here, we put forward that a regional or national level policy limiting visitation during dangerous smoke events may be necessary to protect would-be visitors to U.S.
+national parks [@rai_policy_2020].
+Presently, there is considerable variation in the way states react and plan for climate change and associated hazards [@hu_evaluating_2018].
+
+## Limitations and fututre research
 
@@
 Current data at the monthly scale allows us to investigate the big picture trends in visitor behavior, but future research would benefit greatly from daily count data.
+It is possible that individuals postpone visits to national parks by days or weeks when smoke is present, behavior which our current study would be unable to capture.
 
-It is possible that individuals postpone visits to national parks by days or weeks when smoke is present, behavior which our study current study would be unable to capture.
-It is also possibly that this trend would not hold for recreation on all public lands.
-US national parks draw a great number non-local visitors, many of whom are coming from other states or countries and are likely visiting for the first time [@schuett2010visits].
-We speculate that these visitors are less likely to change their plans due to wildfire smoke than individuals recreating locally or repeatedly in one location.
-We propose that this unique feature of national parks makes their visitors particularly less adaptive to changing climatic conditions.
-Another path for future research might investigate if the trend shown here holds for other public lands with different visitor profiles.
+It is also possible that this trend would not be observed for recreation on all public lands or for all climate change-induced hazards.
+Another path for future research might investigate if the trend shown here holds for other public lands with different visitor profiles and for other hazards.
+
+# Conclusion 
+The results presented here indicate that individuals are not modifying their behavior to adapt to worsening wildfire smoke events.
+It is unclear if this finding is representative of individual climate adaptation generally or is unique to this system.
+Regardless of the generality of our findings, this study has specific implications for management of U.S.
+national parks.
+Future research may confirm or negate the trend shown here in individual climate adaptation broadly.
+This will have considerable implications for the future of all communities living with climate-induced hazards.
 

### 2021-07-07 MarkdownDocuments/Manuscript_Real.Rmd
fig captions drafted

@@
 Climate change has already begun to dramatically change the social-ecological landscapes in which humans have learned to operate [@IPCC].
-All levels of human society must urgently adapt to these new realities if we are to continue to thrive over the coming millenia [@IPCC].
+All levels of human society must urgently adapt to these new realities if we are to continue to thrive over the coming decades[@IPCC].
 As a society, we are aware of this, but still are failing to respond adequately [@steg_limiting_2018].
@@
 Fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].
+referred to
 
-![image alt text. Righ-hand y axis shows the standardized values of smoke ](../Figures/smkplt.png)
+![Smoke observations in each of the 32 national parks included in this study from 1980-2019.
+Smoke events greater than 5 on our standardized scale (10 standard deviations above the mean) are labeled.
+The right-hand y-axis shows the standardized values of smoke referred to when describing the chosen critical thresholds.
+Smoke values on the left-hand y-axis are....ALEX..](../Figures/smkplt.png)
 
@@
 
-![image alt text](../Figures/bpplot2.png)
+![Conceptual diagram showing the design of the three breakpoint models used in this analysis.
+Note that slope 1 is not necessarily 0 in the design of these models.
+All three models also do not necessarily have different intercepts.
+The lines showing the three different breakpoints are separated along the y-axis only for visual clarity.](../Figures/bpplot2.png)
 
@@
 
-![image alt text](../Figures/ARVal.png)
+![Scatterplot showing the within-sample predictive accuracy of the autoregressive baseline model used in this study. This model accounts for 40% of all variation in the data.](../Figures/ARVal.png)
 
@@
 
-![image alt text](../Figures/slpz1.png)
+![Posterior parameter estimates for slope 1 for each of the 32 national parks included in this study for each of the three breakpoint models.
+Points represent the median prediction.
+Thick lines show the 50% credibility intervals and thin lines show the 90% intervals.](../Figures/slpz1.png)
 
-![image alt text](../Figures/slpz2.png)
+![Posterior parameter estimates for slope 2 for each of the 32 national parks included in this study for each of the three breakpoint models.
+Points represent the median prediction.
+Thick lines show the 50% credibility intervals and thin lines show the 90% intervals.](../Figures/slpz2.png)
 

### 2021-08-10 MarkdownDocuments/Manuscript_Real.Rmd
manuscript edits and a figure for the supplemental

@@
 abstract: Humans must adapt to our increasingly variable and hazardous climate.
-Worsening wildfire smoke events in the American West represent a serious risk to human health and also a pertinent test-case for individual climate adaptation generally.
+Worsening wildfire smoke events in the American West represent a serious risk to human health and also a pertinent test case for individual climate adaptation generally.
 Here we aggregate data on ambient smoke and recreational visitation in 32 U.S.
@@
 Climate change has already begun to dramatically change the social-ecological landscapes in which humans have learned to operate [@IPCC].
-All levels of human society must urgently adapt to these new realities if we are to continue to thrive over the coming decades[@IPCC].
+All levels of human society must urgently adapt to these new realities if we are to continue to thrive over the coming decades [@IPCC].
 As a society, we are aware of this, but still are failing to respond adequately [@steg_limiting_2018].
@@
 
-A particularly salient environmental effect of climate change in the United States is the increasing wildfire frequency and severity over the last 40 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016].
+A particularly salient environmental outcome of climate change in the United States is the increasing wildfire frequency and severity over the last 40 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016].
 As climate change accounts for approximately half of the increase in burned area during this time, fires are expected to worsen over the next 40 years as the climate continues to warm [@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
@@
 Fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].
-referred to
 
@@
 Wildfire smoke is a complex mixture of gasses aerated by biomass combustion across the landscape [@naeher_woodsmoke_2007].
-While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter ($\text{PM}_{2.5}$) is particularly harmful to human health [@deryugina_mortality_2019].
+While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter ($\text{PM}_{2.5}$) are particularly harmful to human health [@deryugina_mortality_2019].
 $\text{PM}_{2.5}$ causes between 260,000 - 600,000 global deaths annually and significantly increases community mortality even under acute exposure [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013].

### 2021-08-10 SmokeCorPlot.R

@@ -1,0 +1,24 @@
+library(tidyverse)
+BS_Column<-read_csv("~/SmokeProject/Data/BS_Data/BS_column_mean.csv")
+BS_Surface<-read_csv("~/SmokeProject/Data/BS_Data/BS_surface_mean.csv")
+BS_Col_long<-BS_Column%>%gather(key="time",value="smokeCol",m_198001:m_202009)
+BS_Surf_long<-BS_Surface%>%gather(key="time",value="smokeSurf",m_198001:m_202009)
+
+smokecor<-cor(BS_Col_long$smokeCol,BS_Surf_long$smokeSurf)
+
+BSD<-cbind(BS_Surf_long,BS_Col_long[,3])
+
+mytheme<- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+                plot.title = element_text( size=18, color="black",face="bold"),
+                axis.title.x = element_text( size=18),
+                axis.title.y = element_text( size=18),
+                axis.text=(element_text(color="black", size=14)),
+                legend.title = element_text(colour="black", size=18),
+                legend.text = element_text( size = 14))
+
+ggplot(BSD,aes(x=smokeCol*1e5,y=smokeSurf*1e8))+
+  geom_abline(intercept = 0, slope = 1,linetype="dashed",color="#525252",size=1.5 )+
+  geom_point(alpha=0.5)+theme_classic()+mytheme+
+  ylab("Surface Monthly Mean e-8")+xlab("Column Monthly Mean e-5")+
+  geom_text(x=0.5,y=1.5,color="black",label=paste0("cor = ", round(smokecor,digits = 2)),size=8)
+

### 2021-12-19 MarkdownDocuments/Manuscript_Real.Rmd
file path chng

@@
 ---
-title: Breakpoint Models Show No Evidence of Recreational Adaptation to Increasing Wildfire Smoke in the American West
+title: Breakpoint Models Show No Evidence of Recreational Response to Increasing Wildfire Smoke in the American West
 author:
@@
   name: Vicken Hillis
-- affiliation: Boise State University
-  name: Laura? 
+
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 bibliography: [../SmokePaper.bib, ../OtherRefs.bib]
-abstract: Humans must adapt to our increasingly variable and hazardous climate.
-Worsening wildfire smoke events in the American West represent a serious risk to human health and also a pertinent test case for individual climate adaptation generally.
-Here we aggregate data on ambient smoke and recreational visitation in 32 U.S.
-national parks from 1980 - 2019 to test for observable adaptation to worsening wildfire smoke.
-We develop a series of breakpoint models to allow for threshold effects in recreational response.
-These models demonstrate that even when dangerous levels of smoke are present, individuals are failing to limit outdoor recreation.
-Future research may further our understanding of individual-level climate adaptation by focusing on alternative hazards or finer time series.
+abstract: Ambient wildfire smoke in the American West has worsened considerably in recent decades, while the number of individuals recreating outdoors has simultaneously surged.
+Wildfire smoke poses a serious risk to human health, especially during long periods of exposure and during exercise.
+Here we aggregate data on surface level ($\text{PM}_{2.5}$) and recreational visitation in 32 U.S.
+national parks from 1980 - 2019 to examine how visitors respond to ambient wildfire smoke.
+We hypothesize that visitor response may exhibit a threshold effect where ambient smoke reduces visitation after a critical level, but not before.
+We develop a series of breakpoint models to test this hypothesis.
+Overall, these models show little to no effect of ambient smoke on visitation to the 32 parks tested, even when allowing for critical thresholds at the extreme upper ranges of the smoke data.
+This finding has implications for the management of recreation areas, public health, and climate change adaptation broadly.
 
@@
 
-setwd("C:/Users/Matthewclark989/Documents/SmokeProject")
+setwd("C:/Users/Matt/Documents/SmokeProject")
 ```
@@
 # Introduction
-Climate change has already begun to dramatically change the social-ecological landscapes in which humans have learned to operate [@IPCC].
-All levels of human society must urgently adapt to these new realities if we are to continue to thrive over the coming decades [@IPCC].
-As a society, we are aware of this, but still are failing to respond adequately [@steg_limiting_2018].
-Even when adaptive strategies and their benefits are known, we do not demonstrate adaptive behaviors [@van_valkengoed_meta-analyses_2019].
-While research has shown that extreme events tend to spur climate adaptation at the government level, less is known about how individuals are reacting and possibly adapting to a changing climate [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].
-
-A particularly salient environmental outcome of climate change in the United States is the increasing wildfire frequency and severity over the last 40 years, primarily in the American West [@kitzberger_contingent_2007; @westerling_increasing_2016].
-As climate change accounts for approximately half of the increase in burned area during this time, fires are expected to worsen over the next 40 years as the climate continues to warm [@westerling_increasing_2016; @abatzoglou_impact_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
+Wildfires in the American West have increased dramatically in both frequency and size over the last 40 years [@kitzberger_contingent_2007; @westerling_increasing_2016].
+Approximately half of this increase in burned area can be attributed to anthropogenic climate change [@abatzoglou_impact_2016].
+Thus, wildfires in western forests are expected to continue to worsen over the next 40 years as the climate continues to warm [@westerling_increasing_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
 Accompanying the increase in wildfires in the American West, is a projected increase in wildfire smoke [@schroeder_validation_2008].
 Fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].
+
+Wildfire smoke is a complex mixture of gasses aerated by biomass combustion across the landscape [@naeher_woodsmoke_2007].
+While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter ($\text{PM}_{2.5}$) are particularly harmful to human health [@deryugina_mortality_2019].
+$\text{PM}_{2.5}$ causes between 260,000 - 600,000 global deaths annually and significantly increases community mortality even under acute exposure, especially in areas experiencing outbreaks of COVID-19 infection [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013; @zhou2021excess].
 
@@
 
-Wildfire smoke is a complex mixture of gasses aerated by biomass combustion across the landscape [@naeher_woodsmoke_2007].
-While many of these compounds have been shown to cause health problems in humans, organic carbon and black carbon particulate matter less than 2.5 $\mu$m in diameter ($\text{PM}_{2.5}$) are particularly harmful to human health [@deryugina_mortality_2019].
-$\text{PM}_{2.5}$ causes between 260,000 - 600,000 global deaths annually and significantly increases community mortality even under acute exposure [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013].
-
 In some regions of the American West, wildfires now account for 50% of all atmospheric $\text{PM}_{2.5}$, compared to less than 20% in 2010 [@burke_changing_2021].
@@
 
-Recent literature reviews on recreational behavior have found a dearth of research describing how individuals respond to low air quality, as well as research which gives predictions for how climate change will affect the recreation landscape [@otoole_climate_2019;@zajchowski_air_2019-1].
-Additionally, a survey of federal land managers indicated that decision makers in areas such as national parks feel that they have a shortage of information describing how air quality affects the recreational behavior of their visitors [@zajchowski_air_2019].
+While calls to limit outdoor recreation during wildfire smoke events have become ubiquitous in the summertime American West, a survey of federal land managers indicated that they feel they have a shortage of information describing how air quality actually affects the recreational behavior of their visitors [@zajchowski_air_2019].
+Reviews of scientific literature come to a similar conclusion, pointing to a dearth of research describing how individuals respond to low air quality (although see [@keiser2018air]) and how climate change is affecting the recreation landscape [@otoole_climate_2019; @zajchowski_air_2019-1].
 
-It is of particular importance to understand how individuals are responding to wildfire smoke because it may provide a test case for how individuals will adapt to climate change generally.
-This article uses ambient smoke and visitation data collected from national parks in the Western and Intermountain United States from 1980-2019 to answer whether individuals are adapting their recreation behavior in response to increasing wildfire smoke.
+This article uses surface level $\text{PM}_{2.5}$ and visitation data collected from national parks in the Western and Intermountain United States from 1980-2019 to answer whether individuals alter their recreation behavior in response to ambient wildfire smoke.
 We hypothesize that the effect of ambient smoke on national park visitation may be nonlinear, only showing a measurable impact in the upper ranges of the observed smoke values.
 Threshold effects such as this are common in ecology and human behavioral sciences [@muradian2001ecological; @huggett2005concept; @macy1991chains].
-Toms & Lesperance [-@toms2003piecewise] show that breakpoint models are an effective tool to account for these thresholds and gather inference about the effects of predictors before and after in natural systems.
-Here we develop a series of three hierarchical breakpoint models to determine if threshold effects are present in national park visitation response to ambient smoke and if so, what the effect of smoke is on visitation to national parks in the American West.
-The case study presented here aims to provide insights as to the overall trajectory of realized climate adaptation on the individual level.
+Toms & Lesperance [-@toms2003piecewise] show that breakpoint models are an effective tool to account for these thresholds and make inferences about the effect of a predictor before and after a particular threshold in natural systems.
+Here we develop a series of three hierarchical breakpoint models to determine if a threshold effect is present in national park visitation response to ambient smoke and if so, what the effect of smoke is on visitation post-threshold.
 
@@
 ### Visitation data
-Visitation data were obtained on the monthly scale through the [National Park Service Visitor Use Statistics Portal](https://irma.nps.gov/STATS/).
+We obtained monthly visitation data through the [National Park Service Visitor Use Statistics Portal](https://irma.nps.gov/STATS/).
 These data are generally collected via car counters, permit information, and concessionaire reporting.
@@
 We retrieved monthly data for all national park units in the Intermountain and Pacific West regions for years 1980 through 2019.
-To reduce issues regarding park seasonality, we then subset our data to only the three month high season associated with each park.
-For example, only Summer (June, July, Aug) data was used for Yellowstone National Park.
+
+The 32 national parks included in the two regions in our sample have different high seasons for visitation.
+Wildfire smoke occurs year-round, but primarily in the Summer months.
+In some seasons for some parks there is a natural visitation drop driven by temperature alone that coincides with highest levels of wildfire smoke.
+To ensure that we estimate visitation changes driven by smoke rather than seasonality, we subset our data to only the three month high season associated with each park.
+For example, only Summer (June, July, Aug.) data was used for Yellowstone National Park and only Winter data (Dec., Jan., Feb.) was used for Death Valley National Park.
 
@@
 ### Modeling
-We first account for underlying change in national park visitation from 1980-2019 by developing a temporal autoregressive model which we validate and build upon to answer our main research question.
-This baseline autoregressive model uses the visitation data from the three previous years to predict the visitation to a given park in a given month.
+Visitation to US national parks has increased sharply in recent years [@bergstrom2020does].
+We first account for this underlying change in national park visitation from 1980-2019 by developing a hierarchical, temporal autoregressive model which we assess for within-sample predictive accuracy and build upon.
+This baseline autoregressive model (equation 1) predicts the visitation to a given park in a given month using the recorded visitation in that park and month over the previous three years.
 This formulation of an autoregressive model is commonly referred to as an AR(*k*) model, with this particular instance holding that *k* = 3 [@standev2021stancore].
@@
 We formulated this baseline autoregressive model using a Bayesian hierarchical framework.
-This approach is appropriate here as it allows us to account between park variation in visitation trends while acknowledging the interconnectedness of these trends through partial pooling [@statrethinkingbook].
+This approach is appropriate here as it allows us to account for between-park variation in visitation trends while acknowledging the interconnectedness of these trends through partial pooling [@statrethinkingbook].
 For each park ($j$) we estimate a parameter value ($\bm{\nu1...k}$) representing the trend in visitation from one, two, and three years prior ($t-1,2,...k$), to the month we are estimating ($i$).
@@
 We modeled these data using a negative binomial distribution to allow us to estimate an overdispersion parameter ($\phi$) rather than assuming the dispersion is equal to the mean.
+We fit this model using 10 chains run for 8,000 iterations.
 
@@
 \text{log}(\mu_{ij}) = 
-\alpha0+\bm{\beta0}_{j}+\bm{\nu1}_{j}*\text{vis}_{i,t-1}+ \\
-\bm{\nu2}_{j}*\text{vis}_{i,t-2}+\bm{\nu3}_{j}*\text{vis}_{i,t-3}  
+\alpha0+\bm{\beta0}_{j}+\bm{\nu1}_{j}\cdot\text{vis}_{i,t-1}+ \\
+\bm{\nu2}_{j}\cdot\text{vis}_{i,t-2}+\bm{\nu3}_{j}\cdot\text{vis}_{i,t-3}  
 \end{equation}
@@
 
-After validating that the model above captures the baseline trends in national park visitation from 1980-2019, we built upon it to estimate if a critical threshold for visitation response to wildfire smoke exists in these data and if so, what the impact of wildfire smoke is on visitation post-threshold.
+This baseline autoregressive model showed adequate diagnostic statistics, exhibiting ample mixing of Markov chains, $\hat{R}$ values equal to 1 for all parameter estimates, and the absence of divergent transitions after warm up.
+We examined the within-sample predictive capacity of this autoregressive-only model to be sure we sufficiently accounted for baseline changes in visitation before building on this model to test for the effects of smoke on visitation.
+This baseline model allowed us to account for approximately 40% of all variation in the data, with the bulk of the error occurring when actual visitation was very high or very low (Fig 2).
+
+![Scatterplot showing the within-sample predictive accuracy of the autoregressive baseline model used in this study. This model accounts for 40% of all variation in the data.](../Figures/ARVal.png)
+
+After checking that the model above captures the baseline trends in national park visitation from 1980-2019, we built upon it to estimate if a critical threshold for visitation response to wildfire smoke exists in these data and if so, what the impact of wildfire smoke is on visitation post-threshold.
 To do this, we developed a series of three breakpoint models, allowing for three increasingly extreme breakpoints (i.e.
@@
 A threshold of 1.0 can then be interpreted as two standard deviations above the mean value of smoke in the data.
-A visual comparison of the observed actual smoke values and standardized values can be seen in Fig.
+A visual comparison of the observed smoke values plotted on both the standardized and unstandardized scales can be seen in Fig.
 1.
 
-As shown conceptually in Fig.
-2, we estimated the effect of smoke on visitation before ($\bm{\beta1}$) and after ($\bm{\beta2}$) each breakpoint (BP).
+We estimated the effect of smoke on visitation before ($\bm{\beta1}$) and after ($\bm{\beta2}$) each breakpoint (BP).
 Just as with the autoregressive terms, we estimated unique parameter values for each park ($j$) in a hierarchical framework.
 The complete model used for each of the three breakpoint values is then as seen in equation 2.
-
-![Conceptual diagram showing the design of the three breakpoint models used in this analysis.
-Note that slope 1 is not necessarily 0 in the design of these models.
-All three models also do not necessarily have different intercepts.
-The lines showing the three different breakpoints are separated along the y-axis only for visual clarity.](../Figures/bpplot2.png)
+Each model was run using 10 chains for 8,000 iterations.
 
@@
 \begin{cases} 
-\alpha0+\bm{\beta0}_{j}+\bm{\beta1}_{j}*\text{smoke}_{i}+\bm{\nu1}_{j}*\text{vis}_{i,t-1}+ \\
-\bm{\nu2}_{j}*\text{vis}_{i,t-2}+\bm{\nu3}_{j}*\text{vis}_{i,t-3} & \text{if } \text{smoke}_{i} < \text{BP} \\ \\
-\beta0+\bm{\beta0}_{j}+\bm{\beta1}_{j}*\text{BP}+(\text{smoke}_{i}-\text{BP})*\bm{\beta2}_{j}+  \\ \bm{\nu1}_{j}*\text{vis}_{i,t-1}+\bm{\nu2}_{j}*\text{vis}_{i,t-2}+\bm{\nu3}_{j}*\text{vis}_{i,t-3}        & \text{if } \text{smoke}_{i} \geq \text{BP}
+\alpha0+\bm{\beta0}_{j}+\bm{\beta1}_{j}\cdot\text{smoke}_{i}+\bm{\nu1}_{j}\cdot\text{vis}_{i,t-1}+ \\
+\bm{\nu2}_{j}\cdot\text{vis}_{i,t-2}+\bm{\nu3}_{j}\cdot\text{vis}_{i,t-3} & \text{if } \text{smoke}_{i} < \text{BP} \\ \\
+\beta0+\bm{\beta0}_{j}+\bm{\beta1}_{j}\cdot\text{BP}+(\text{smoke}_{i}-\text{BP})\cdot\bm{\beta2}_{j}+  \\ \bm{\nu1}_{j}\cdot\text{vis}_{i,t-1}+\bm{\nu2}_{j}\cdot\text{vis}_{i,t-2}+\bm{\nu3}_{j}\cdot\text{vis}_{i,t-3}        & \text{if } \text{smoke}_{i} \geq \text{BP}
 \end{cases}
@@
 [-@gelman2008weakly] for producing stable, conservative estimates, we specified long-tailed regularizing priors for all autoregressive terms, the global intercept value, and the overdispersion parameter.
-We used normally distributed regularizing priors for estimating the variance parameter ($\sigma$) and the pre and post-breakpoint smoke effects to allow for more efficient sampling while generative within-sample predictions at high smoke values.
+We used normally distributed regularizing priors for estimating the standard deviation parameter ($\sigma$) and the pre and post-breakpoint smoke effects to allow for more efficient sampling while generating within-sample predictions at high smoke values.
 
-### Validation
-Due to the considerable computational time (about 1 day) needed to run the program described above, we validated the results of both the autoregressive-only and breakpoint models using within-sample posterior predictive checks.
-All models were first inspected to ensure adequate mixing of Markov chains, $\hat{R}$ values below 1.05 for all parameter estimates, and the absence of divergent transitions after warm up.
+As the goal of this paper is inference, we test just the suite of breakpoint models described above, which were developed a priori to evaluate our hypothesis that smoke would exhibit a non-linear threshold effect on park visitation [@tredennick2021practical].
+We do not test our models against other candidate models using information criterion metrics or cross validation, as the objective of such tests is prediction rather than inference and we would not expect our models to be predictively valid strictly speaking [@akaike1998information; @tredennick2021practical].
 
-We examined the within-sample predictive capacity of the autoregressive only model to be sure we adequately accounted for baseline changes in visitation before building on this model to test for the effects of smoke on visitation.
-This baseline model allowed us to account for approximately 40% of all variation in the data, with the bulk of the error occurring when actual visitation was very high or very low (Fig 3).
-
-![Scatterplot showing the within-sample predictive accuracy of the autoregressive baseline model used in this study. This model accounts for 40% of all variation in the data.](../Figures/ARVal.png)
+```{r, echo=FALSE, eval=FALSE}
+![Conceptual diagram showing the design of the three breakpoint models used in this analysis.
+Note that slope 1 is not necessarily 0 in the design of these models.
+All three models also do not necessarily have different intercepts.
+The lines showing the three different breakpoints are separated along the y-axis only for visual clarity.](../Figures/bpplot2.png)
+```
 
 # Results
+All three breakpoint models showed adequate model diagnostics, including well-mixed markov chains, $\hat{R}$ values less than 1.05 for all estimated parameters, and the absence of divergent transitions after warm up [@gelman2020bayesian].
+We then take the coefficient estimates produced from our models as reliable for inference regarding our a priori hypothesis.
+
 While some between-park variation exists, our overall study shows no evidence for recreational visitor response to wildfire smoke in national parks in the American West.
@@
 In all three breakpoint models, only two parks have 90% credible intervals that do not overlap zero for slope 1, both of which have opposite signs (Fig.
+3).
+This trend is paralleled for slope 2 in the two lower breakpoint models and all 90% credibility intervals overlap 0 for slope 2 in the model with the most extreme breakpoint (Fig.
 4).
-This trend is paralleled for slope 2 in the two lower breakpoint models and all 90% credibility intervals overlap 0 for slope 2 in the model with the most extreme breakpoint (Fig.
-5).
 
@@
 This suggests that ambient smoke in national parks in the American West is not driving any measurable changes in visitation even at sustained and dangerous levels.
-We then conclude that visitors are not altering recreation behavior in response to smoke and therefore no critical threshold for adaptation is detectable in these data.
+We thus conclude that visitors are not altering recreation behavior in response to smoke and therefore no critical threshold for adaptation is detectable in these data.
 
@@
 
+Visually examining the marginal effect of ambient smoke on park recreational visits provides a similar intuition as above (Fig.
+5).
+We see little difference in the effect of smoke before and after the hypothesized thresholds.
+In addition, we descriptively show high levels of visitation even under very high levels of ambient smoke.
+In Redwood National Park for example, we see visitation within the normal range even during the August 2018 wildfire events, which produced smoke levels over 30 times the total sample standard deviation.
+
+![Marginal effect of ambient smoke on monthly park visitation for four of our study parks with the highest median smoke values from 1980 to 2019.
+Parks displayed are Glacier National Park, Grand Teton National Park, Great Sand Dunes National Park & Preserve, and Redwood National Park.
+For simplicity, we show just the model with the lowest breakpoint value (0.5 on our standardized scale).
+Blue shading represents the full range of posterior predictions for visitation given the mean value for each autoregressive term in the model.
+Intensity of shading shows the credibility interval for each prediction, with the middle black line showing the median prediction at each value of smoke.](../Figures/MarginalEffects.png)
+
 # Discussion
-Given the data available to us, we detect no visitor adaptation to increasing wildfire smoke in national parks in the American West.
+Given the data available, we detect no visitor adaptation to increasing wildfire smoke in national parks in the American West.
 This result is troubling both immediately and in the long-term.
@@
 
-Based on our findings here, we put forward that a regional or national level policy limiting visitation during dangerous smoke events may be necessary to protect would-be visitors to U.S.
+Based on our findings here, we suggest that a regional or national level policy limiting visitation during dangerous smoke events may be necessary to protect would-be visitors to U.S.
 national parks [@rai_policy_2020].
@@
 
-A key limitation of this study is the lack of availability of fine scale visitation data to US national parks.
+A key limitation of this study is the lack of fine scale visitation data to US national parks.
 Daily visitation data would allow us to identify a more complete picture of if and how visitors respond to smoke events day to day.
@@
 
+# Supplemental material
+![We should correlate the column values to just the surface layer values to show they are correlated.
+There doesn’t seem to be a set standard on which to use and I have seen both argued for.
+Surface is clearly what matters to visitors, but I guess the algorithm isn’t super accurate for the lowest layer.
+Including the entire column captures everything, but layers higher up in the column are susceptible to move away from the pixel location from wind.
+](../Figures/SmokeCor.png)
+
 \newpage

### 2022-04-20 MarkdownDocuments/Manuscript_Real.Rmd
Hopefully final commit. Cleaning up uncomitted stuff

@@
 ---
-title: Breakpoint Models Show No Evidence of Recreational Response to Increasing Wildfire Smoke in the American West
+title: Breakpoint Models Show No Evidence of Thresholds in Recreational Response to Increasing Wildfire Smoke in the American West
 author:
@@
   name: Alexander Killion
-- affiliation: Boise State University
+- affiliation: Boise State University - Human Environment Systems
   name: Matthew A. Williamson
-- affiliation: Boise State University
+- affiliation: Boise State University - Human Environment Systems
   name: Vicken Hillis
@@
 Wildfire smoke poses a serious risk to human health, especially during long periods of exposure and during exercise.
-Here we aggregate data on surface level ($\text{PM}_{2.5}$) and recreational visitation in 32 U.S.
-national parks from 1980 - 2019 to examine how visitors respond to ambient wildfire smoke.
+Here we aggregate data on black carbon, a major component of wildfire smoke, and recreational visitation in 32 U.S.
+national parks from 1980 - 2019 to examine how visitors respond to wildfire smoke.
 We hypothesize that visitor response may exhibit a threshold effect where ambient smoke reduces visitation after a critical level, but not before.
@@
 Overall, these models show little to no effect of ambient smoke on visitation to the 32 parks tested, even when allowing for critical thresholds at the extreme upper ranges of the smoke data.
+This suggests that wildfire smoke does not significantly alter behavior of park attendance.
 This finding has implications for the management of recreation areas, public health, and climate change adaptation broadly.
@@
 
-# Introduction
+# 1 Introduction
 Wildfires in the American West have increased dramatically in both frequency and size over the last 40 years [@kitzberger_contingent_2007; @westerling_increasing_2016].
 Approximately half of this increase in burned area can be attributed to anthropogenic climate change [@abatzoglou_impact_2016].
-Thus, wildfires in western forests are expected to continue to worsen over the next 40 years as the climate continues to warm [@westerling_increasing_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015].
-Accompanying the increase in wildfires in the American West, is a projected increase in wildfire smoke [@schroeder_validation_2008].
-Fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014].
+Thus, wildfires in western forests are expected to worsen over the next 40 years as the climate continues to warm [@westerling_increasing_2016; @yue_ensemble_2013; @wuebbles_climate_2017; @harvey_human-caused_2016; @peterson_new_2003; @barbero_climate_2015; @balch_warming_2022].
+Accompanying the increase in wildfires in the American West, is a projected increase in wildfire smoke; fire-prone areas may experience a two-fold increase in wildfire smoke by 2100 [@hurteau_projected_2014; @schroeder_validation_2008].
 
@@
 $\text{PM}_{2.5}$ causes between 260,000 - 600,000 global deaths annually and significantly increases community mortality even under acute exposure, especially in areas experiencing outbreaks of COVID-19 infection [@johnston_fay_h_estimated_2012; @naeher_woodsmoke_2007; @kloog_long-_2013; @zhou2021excess].
+While both $\text{PM}_{2.5}$ and wildfire smoke generally are made up of organic material (i.e.
+organic carbon) and black carbon, black carbon is the most visible component [@yu_black_2019].
+Black carbon is particularly prominent in visible wildfire smoke, as it has high light absorption properties that contribute to the dark appearance of plumes [@petzold_recommendations_2013].
 
-![Smoke observations in each of the 32 national parks included in this study from 1980-2019.
-Smoke events greater than 5 on our standardized scale (10 standard deviations above the mean) are labeled.
+![Monthly smoke observations in each of the 32 national parks included in this study from 1980-2019.
+Monthly medians greater than 5 on our standardized scale (10 standard deviations above the mean) are labeled.
 The right-hand y-axis shows the standardized values of smoke referred to when describing the chosen critical thresholds.
-Smoke values on the left-hand y-axis are....ALEX..](../Figures/smkplt.png)
+Smoke values on the left-hand y-axis are shown in kilograms per time-averaged two-dimensional meter.](../Figures/smkplt.png)
 
@@
 While calls to limit outdoor recreation during wildfire smoke events have become ubiquitous in the summertime American West, a survey of federal land managers indicated that they feel they have a shortage of information describing how air quality actually affects the recreational behavior of their visitors [@zajchowski_air_2019].
-Reviews of scientific literature come to a similar conclusion, pointing to a dearth of research describing how individuals respond to low air quality (although see [@keiser2018air]) and how climate change is affecting the recreation landscape [@otoole_climate_2019; @zajchowski_air_2019-1].
+Reviews of scientific literature come to a similar conclusion, pointing to a dearth of research describing how individuals respond to low air quality (although see Keiser, Lade, and Rudik [-@keiser2018air]) and how climate change is affecting the recreation landscape [@otoole_climate_2019; @zajchowski_air_2019-1].
 
-This article uses surface level $\text{PM}_{2.5}$ and visitation data collected from national parks in the Western and Intermountain United States from 1980-2019 to answer whether individuals alter their recreation behavior in response to ambient wildfire smoke.
+This article uses estimates of black carbon and visitation data collected from national parks in the American West from 1980-2019 to answer whether individuals alter their recreation behavior in response to ambient wildfire smoke.
 We hypothesize that the effect of ambient smoke on national park visitation may be nonlinear, only showing a measurable impact in the upper ranges of the observed smoke values.
@@
 
-# Methods
-## Data collection
+# 2 Methods
+## 2.1 Data collection
 
-### Visitation data
-We obtained monthly visitation data through the [National Park Service Visitor Use Statistics Portal](https://irma.nps.gov/STATS/).
+### 2.1.1 Visitation data
+We obtained monthly visitation data through the [National Park Service (NPS) Visitor Use Statistics Portal](https://irma.nps.gov/STATS/).
 These data are generally collected via car counters, permit information, and concessionaire reporting.
@@
 
+While wildfires have seen a dramatic increase across the United States over the last four decades, changes in wildfire smoke have been most concentrated in the westernmost regions of the country [@burke_changing_2021].
+In the contiguous United States, the western regions are the only areas with a significant smoke burden where the majority of the smoke they see originates from wildfires within their own borders [@brey_connecting_2018].
+Therefore, as to not confound our analysis with unpredictable smoke events which originate from distant wildfires, we limit our sample to the national parks in the American West (i.e.
+Intermountain and Pacific West NPS regions).
+
 The 32 national parks included in the two regions in our sample have different high seasons for visitation.
 Wildfire smoke occurs year-round, but primarily in the Summer months.
-In some seasons for some parks there is a natural visitation drop driven by temperature alone that coincides with highest levels of wildfire smoke.
+In some seasons, for some parks, there is a natural visitation drop driven by temperature alone that coincides with highest levels of wildfire smoke.
 To ensure that we estimate visitation changes driven by smoke rather than seasonality, we subset our data to only the three month high season associated with each park.
@@
 
-### Smoke data
-Alex to fill in this part
+### 2.1.2 Smoke data
+To detect times of high concentrations of black carbon associated with smoke, we used data from the second Modern-Era Retrospective analysis for Research and Applications [@gelaro_modern-era_2017].
+MERRA-2 is a NASA atmospheric reanalysis that begins in 1980 and replaces the original MERRA reanalysis using an upgraded version of the Goddard Earth Observing System Model, Version 5 data assimilation system [@rienecker_merra_2011].
+MERRA-2 provides mean monthly measurements of black carbon starting in 1980 at a spatial resolution of 0.625° x 0.5°.
+We used the monthly black carbon column mass density measurements (kg/$\text{m}^{-2}$) and calculated the mean of those pixels that intersected national park boundaries to represent wildfire smoke at the park level [@merra_tavgu_2d_lnd_nx_2015].
+We also correlated the column black carbon values to the amount of black carbon only at the surface layer (i.e., the measurements closest to the ground; S1).
 
-## Analyses
+## 2.2 Analyses
 
-### Modeling
 Visitation to US national parks has increased sharply in recent years [@bergstrom2020does].
 We first account for this underlying change in national park visitation from 1980-2019 by developing a hierarchical, temporal autoregressive model which we assess for within-sample predictive accuracy and build upon.
-This baseline autoregressive model (equation 1) predicts the visitation to a given park in a given month using the recorded visitation in that park and month over the previous three years.
-This formulation of an autoregressive model is commonly referred to as an AR(*k*) model, with this particular instance holding that *k* = 3 [@standev2021stancore].
-We chose a *k* of 3 as the added within-sample predictive value of increasing *k* diminished considerably for any greater value.
+This baseline autoregressive model (equation 1) predicts the visitation to a given park in a given month based on the recorded visitation in that park and month in the previous year.
+This formulation of an autoregressive model is commonly referred to as an AR(*k*) model, where *k* is the number of previous time periods used for prediction [@standev2021stancore].
+In this particular instance we chose *k* = 1 to reduce issues of collinearity and identifiability among autoregressive predictors [@pesaran_bayesian_2019].
 
 We formulated this baseline autoregressive model using a Bayesian hierarchical framework.
-This approach is appropriate here as it allows us to account for between-park variation in visitation trends while acknowledging the interconnectedness of these trends through partial pooling [@statrethinkingbook].
-For each park ($j$) we estimate a parameter value ($\bm{\nu1...k}$) representing the trend in visitation from one, two, and three years prior ($t-1,2,...k$), to the month we are estimating ($i$).
+This approach accounts for between-park variation in visitation trends while acknowledging the interconnectedness of these trends through partial pooling [@statrethinkingbook].
+For each park ($j$) we estimate a parameter value ($\bm{\nu}$) representing the trend in visitation from one year to the next, in order to predict the visitation for each month in our data set ($i$).
 Finally, we estimate both global ($\alpha0$) and park specific ($\bm{\beta0}_{j}$) intercepts to yield equation 1 below.
 We modeled these data using a negative binomial distribution to allow us to estimate an overdispersion parameter ($\phi$) rather than assuming the dispersion is equal to the mean.
-We fit this model using 10 chains run for 8,000 iterations.
+We fit this model using 7 Markov chains run for 5,000 iterations.
 
@@
 \text{log}(\mu_{ij}) = 
-\alpha0+\bm{\beta0}_{j}+\bm{\nu1}_{j}\cdot\text{vis}_{i,t-1}+ \\
-\bm{\nu2}_{j}\cdot\text{vis}_{i,t-2}+\bm{\nu3}_{j}\cdot\text{vis}_{i,t-3}  
+\alpha0+\bm{\beta0}_{j}+\bm{\nu}_{j}\cdot\text{vis}_{i,t-1} \\
 \end{equation}
@@
 We examined the within-sample predictive capacity of this autoregressive-only model to be sure we sufficiently accounted for baseline changes in visitation before building on this model to test for the effects of smoke on visitation.
-This baseline model allowed us to account for approximately 40% of all variation in the data, with the bulk of the error occurring when actual visitation was very high or very low (Fig 2).
+This baseline model allowed us to account for approximately 37% of all variation in the data, with the bulk of the error occurring when actual visitation was very high or very low (Fig 2).
 
-![Scatterplot showing the within-sample predictive accuracy of the autoregressive baseline model used in this study. This model accounts for 40% of all variation in the data.](../Figures/ARVal.png)
+![Scatterplot showing the within-sample predictive accuracy of the autoregressive baseline model used in this study.
+This model accounts for 37% of all variation in the data.
+Black dashed line shows an exact 1:1 relationship between the predicted visitation and actual visitation axes.
+](../Figures/ARVal.png)
 
-After checking that the model above captures the baseline trends in national park visitation from 1980-2019, we built upon it to estimate if a critical threshold for visitation response to wildfire smoke exists in these data and if so, what the impact of wildfire smoke is on visitation post-threshold.
+After confirming that the model above captures the baseline trends in national park visitation from 1980-2019, we built upon it to estimate if a critical threshold for visitation response to wildfire smoke exists in these data and if so, what the impact of wildfire smoke is on visitation post-threshold.
 To do this, we developed a series of three breakpoint models, allowing for three increasingly extreme breakpoints (i.e.
@@
 We standardized these smoke data, and thus the thresholds as well, by dividing by two standard deviations following the recommendation made by Gelman [-@gelman2008scaling].
-A threshold of 1.0 can then be interpreted as two standard deviations above the mean value of smoke in the data.
-A visual comparison of the observed smoke values plotted on both the standardized and unstandardized scales can be seen in Fig.
-1.
+The three thresholds tested were 0.0, 0.5, and 1.0, which can thus be interpreted as the mean smoke value, one standard deviation above the mean, and two standard deviations above the mean respectively.
+These increasingly extreme thresholds contain the highest 45%, 8%, and 3% of observed smoke values out of the 3,744 data points in our sample.
+A visual comparison of the observed smoke values plotted on both the standardized and unstandardized scales can be seen in figure 1.
 
 We estimated the effect of smoke on visitation before ($\bm{\beta1}$) and after ($\bm{\beta2}$) each breakpoint (BP).
-Just as with the autoregressive terms, we estimated unique parameter values for each park ($j$) in a hierarchical framework.
+Just as with the autoregressive term ($\bm{\nu}$), we estimated unique parameter values for each park ($j$) in a hierarchical framework.
 The complete model used for each of the three breakpoint values is then as seen in equation 2.
-Each model was run using 10 chains for 8,000 iterations.
+Each model was run using 7 Markov chains for 5,000 iterations.
 
@@
 \begin{cases} 
-\alpha0+\bm{\beta0}_{j}+\bm{\beta1}_{j}\cdot\text{smoke}_{i}+\bm{\nu1}_{j}\cdot\text{vis}_{i,t-1}+ \\
-\bm{\nu2}_{j}\cdot\text{vis}_{i,t-2}+\bm{\nu3}_{j}\cdot\text{vis}_{i,t-3} & \text{if } \text{smoke}_{i} < \text{BP} \\ \\
-\beta0+\bm{\beta0}_{j}+\bm{\beta1}_{j}\cdot\text{BP}+(\text{smoke}_{i}-\text{BP})\cdot\bm{\beta2}_{j}+  \\ \bm{\nu1}_{j}\cdot\text{vis}_{i,t-1}+\bm{\nu2}_{j}\cdot\text{vis}_{i,t-2}+\bm{\nu3}_{j}\cdot\text{vis}_{i,t-3}        & \text{if } \text{smoke}_{i} \geq \text{BP}
+\alpha0+\bm{\beta0}_{j}+\bm{\beta1}_{j}\cdot\text{smoke}_{i}+\bm{\nu}_{j}\cdot\text{vis}_{i,t-1} 
+ & \text{if } \text{smoke}_{i} < \text{BP} \\ \\
+\beta0+\bm{\beta0}_{j}+\bm{\beta1}_{j}\cdot\text{BP}+(\text{smoke}_{i}-\text{BP})\cdot\bm{\beta2}_{j}+  \\ \bm{\nu}_{j}\cdot\text{vis}_{i,t-1}\ & \text{if } \text{smoke}_{i} \geq \text{BP}
 \end{cases}
@@
 Following the recommendation given by Gelman et al.
-[-@gelman2008weakly] for producing stable, conservative estimates, we specified long-tailed regularizing priors for all autoregressive terms, the global intercept value, and the overdispersion parameter.
+[-@gelman2008weakly] for producing stable, conservative estimates, we specified long-tailed regularizing priors for the autoregressive term, the global intercept value, and the overdispersion parameter.
 We used normally distributed regularizing priors for estimating the standard deviation parameter ($\sigma$) and the pre and post-breakpoint smoke effects to allow for more efficient sampling while generating within-sample predictions at high smoke values.
 
-As the goal of this paper is inference, we test just the suite of breakpoint models described above, which were developed a priori to evaluate our hypothesis that smoke would exhibit a non-linear threshold effect on park visitation [@tredennick2021practical].
+As the goal of this paper is inference, we test just the suite of breakpoint models described above, which were developed *a priori* to evaluate our hypothesis that smoke would exhibit a non-linear threshold effect on park visitation [@tredennick2021practical].
 We do not test our models against other candidate models using information criterion metrics or cross validation, as the objective of such tests is prediction rather than inference and we would not expect our models to be predictively valid strictly speaking [@akaike1998information; @tredennick2021practical].
 
-```{r, echo=FALSE, eval=FALSE}
-![Conceptual diagram showing the design of the three breakpoint models used in this analysis.
-Note that slope 1 is not necessarily 0 in the design of these models.
-All three models also do not necessarily have different intercepts.
-The lines showing the three different breakpoints are separated along the y-axis only for visual clarity.](../Figures/bpplot2.png)
-```
+# 3 Results
+All three breakpoint models showed adequate model diagnostics, including well-mixed Markov chains, $\hat{R}$ values of 1 for all estimated parameters, and the absence of divergent transitions after warm up [@gelman2020bayesian].
+We then take the coefficient estimates produced from our models as reliable for inference regarding our *a priori* hypothesis.
 
-# Results
-All three breakpoint models showed adequate model diagnostics, including well-mixed markov chains, $\hat{R}$ values less than 1.05 for all estimated parameters, and the absence of divergent transitions after warm up [@gelman2020bayesian].
-We then take the coefficient estimates produced from our models as reliable for inference regarding our a priori hypothesis.
-
-While some between-park variation exists, our overall study shows no evidence for recreational visitor response to wildfire smoke in national parks in the American West.
-Figures 4 and 5 show the parameter estimates for the pre and post breakpoint smoke effects respectively.
+While some between-park variation exists, our overall study shows no evidence for threshold effects in recreational visitor response to wildfire smoke in national parks in the American West.
+Figures 3 and 4 show the parameter estimates for the pre and post breakpoint smoke effects respectively.
 In all three breakpoint models, only two parks have 90% credible intervals that do not overlap zero for slope 1, both of which have opposite signs (Fig.
 3).
-This trend is paralleled for slope 2 in the two lower breakpoint models and all 90% credibility intervals overlap 0 for slope 2 in the model with the most extreme breakpoint (Fig.
+The post-breakpoint, slope 2, parameter estimates exclusively overlap zero at a 90% credibility interval (Fig.
 4).
-
-Intuitively, the within-sample $R^2$ increases only slightly when accounting for smoke effects, from 0.40 to 0.41 compared to the autoregressive only model.
-This suggests that ambient smoke in national parks in the American West is not driving any measurable changes in visitation even at sustained and dangerous levels.
-We thus conclude that visitors are not altering recreation behavior in response to smoke and therefore no critical threshold for adaptation is detectable in these data.
+The uncertainty in these estimates increases dramatically as we use more extreme breakpoint values and the amount of data post-breakpoint decreases.
+We note however, that even parks with relatively certain estimates in the more extreme breakpoint models (0.5 & 1) still exclusively overlap 0, demonstrating no impact of ambient smoke on park visitation even at very dangerous levels.
 
@@
 Parks displayed are Glacier National Park, Grand Teton National Park, Great Sand Dunes National Park & Preserve, and Redwood National Park.
-For simplicity, we show just the model with the lowest breakpoint value (0.5 on our standardized scale).
+For simplicity, we show just the model with the middle breakpoint value (0.5 on our standardized scale).
 Blue shading represents the full range of posterior predictions for visitation given the mean value for each autoregressive term in the model.
@@
 
-# Discussion
-Given the data available, we detect no visitor adaptation to increasing wildfire smoke in national parks in the American West.
-This result is troubling both immediately and in the long-term.
-As discussed above, $\text{PM}_{2.5}$ significantly increases community morbidity even under acute exposure.
-In concert with showing no overall trends in behavioral adaptation, these data provide specific instances of historic smoke events where visitation did not deviate from normal.
+Given the clear lack of threshold effects in visitation response to ambient wildfire smoke, we conducted an exploratory analysis on the overall effect of smoke on visitation in our study parks (i.e.
+without a breakpoint).
+We formulate this exactly as in the pre-breakpoint slope in equation 2.
+We again ran this model for 5,000 iterations using 7 Markov chains, which exhibited adequate mixing, a lack of divergent transitions after warmup, and $\hat{R}$ values of 1 for all estimated parameters.
 
-The highly variable nature of wildfires, their relationship to climate change, and the great distances that smoke can travel during and after wildfire events make it difficult for individuals to plan around smoke events [@dombeck_wildfire_2004; @collins_regional_2006].
+The parameter estimates for the overall effect of ambient smoke on visitation response in all 32 parks in our sample are considerably more narrow than the estimates produced via the breakpoint models (Fig.
+6).
+Still, the 90% credibility intervals estimating the smoke effect for each park exclusively overlap 0, indicating no overall effect, further supporting our findings above.
+Intuitively, the within-sample $R^2$ remains equal to 0.37 after accounting for the overall smoke effects, showing no improvement compared to the autoregressive only model.
+This suggests that ambient smoke in national parks in the American West is not driving any measurable changes in visitation even at sustained and dangerous levels.
+We thus conclude that visitors are not altering recreation behavior in response to smoke and therefore no overall effect or critical threshold for adaptation is detectable in these data.
+
+![Posterior parameter estimates for the overall effect of smoke on visitation for each of the 32 national parks included in this study.
+Points represent the median prediction.
+Thick lines show the 50% credibility intervals and thin lines show the 90% intervals.](../Figures/NoBPPosterior.png)
+
+# 4 Discussion
+
+We did not detect visitor adaptation to increasing wildfire smoke in national parks in the American West.
+This result is troubling both specifically for visitor health in U.S.
+national parks and for climate change adaptation broadly.
+As discussed above, wildfire smoke significantly increases community morbidity even under acute exposure.
+In concert with showing no overall trends in behavioral adaptation, these data provide specific instances of historic smoke events where visitation did not deviate from normal (Fig.
+5).
+
+The highly variable nature of wildfires, their relationship to climate change, and the great distances that smoke can travel during and after wildfire events make it difficult for individuals to plan around smoke events [@dombeck_wildfire_2004; @collins_regional_2006; @brey_connecting_2018].
 In conjunction, U.S.
@@
 
-## Limitations and fututre research
+Considering our findings more generally, humans must adapt quickly to the new realities of our increasingly variable climate if we are to continue to thrive over the coming decades [@masson-delmotte_ipcc_2021].
+Climate change is already dramatically altering the social-ecological landscapes in which humans have learned to operate [@folke_our_2021].
+This research contributes to a broader body of work showing that despite increasing awareness, as a society we still fail to respond adequately [@steg_limiting_2018].
+Even when adaptive strategies and their benefits are known, we do not translate strategies into action [@van_valkengoed_meta-analyses_2019].
+While research has shown that extreme events tend to spur climate adaptation on the part of governments, individual response is much less consistent [@bierbaum_comprehensive_2013; @berrang-ford_tracking_2019; @berrang-ford_are_2011].
+Therefore, as we speculate that the system of recreational visitation to U.S.
+national parks may be particularly problematic, we suppose that the trends observed in this study may be characteristic of individual response to many changing environmental conditions.
+
+## 4.1 Limitations and future research
 
@@
 It is also possible that this trend would not be observed for recreation on all public lands or for all climate change-induced hazards.
-Another path for future research might investigate if the trend shown here holds for other public lands with different visitor profiles and for other hazards.
+Other paths for future research may therefore be to investigate if the trend shown here holds for other public lands with different visitor profiles and for other hazards.
 
-# Conclusion 
+# 5 Conclusion 
 The results presented here indicate that individuals are not modifying their behavior to adapt to worsening wildfire smoke events.

### 2022-04-20 Rcode/FinalFullAnalysisCode.R

@@ -12,17 +12,17 @@ x<-dat[385:1920,]$AR_Vis==dat[1:1536,]$RecreationVisits
 which(x != TRUE)
 
 #create a 2 year AR lag
-dat$AR_Vis2<-lag(dat$RecreationVisits,n=768)
-x<-dat[769:1920,]$AR_Vis2==dat[1:1152,]$RecreationVisits
-which(x != TRUE)
+#dat$AR_Vis2<-lag(dat$RecreationVisits,n=768)
+#x<-dat[769:1920,]$AR_Vis2==dat[1:1152,]$RecreationVisits
+#which(x != TRUE)
 
 #create a 3 year AR lag
-dat$AR_Vis3<-lag(dat$RecreationVisits,n=1152)
-x<-dat[1153:1920,]$AR_Vis3==dat[1:768,]$RecreationVisits
-which(x != TRUE)
+#dat$AR_Vis3<-lag(dat$RecreationVisits,n=1152)
+#x<-dat[1153:1920,]$AR_Vis3==dat[1:768,]$RecreationVisits
+#which(x != TRUE)
 
 #chaeck a random row to make sure it's all good
-dat[dat$UnitCode == "YOSE" & dat$Year ==2012 & dat$Month==7,]$RecreationVisits == dat[dat$UnitCode == "YOSE" & dat$Year ==2015 & dat$Month==7,]$AR_Vis3
+#dat[dat$UnitCode == "YOSE" & dat$Year ==2012 & dat$Month==7,]$RecreationVisits == dat[dat$UnitCode == "YOSE" & dat$Year ==2015 & dat$Month==7,]$AR_Vis3
 
 #get ridof the 1st 3 years of data
 dat<-na.omit(dat)
@@ -41,20 +41,20 @@ data_list <- list(
   count = dat$RecreationVisits,
   #smoke = stdize(dat$medSmoke),
   arVis = stdize(dat$AR_Vis),
-  arVis2 = stdize(dat$AR_Vis2),
-  arVis3 = stdize(dat$AR_Vis3),
+  #arVis2 = stdize(dat$AR_Vis2),
+  #arVis3 = stdize(dat$AR_Vis3),
   pcode = as.numeric(as.factor(dat$UnitCode ))
   )
 
-options(mc.cores=10)
+options(mc.cores=7)#10
 
 rstan::rstan_options(autowrite=TRUE)
 
 mod<-rstan::stan( file="~/SmokeProject/StanCode/AROnly.stan " , 
[…]

### 2022-04-20 Rcode/PaperPlotsFinal.R

@@ -55,12 +55,13 @@ HSK<-dat%>%mutate(ParkName= gsub(" NP","",dat$ParkName))%>%filter(SeasType=="Hig
 smokePlt<-dat%>%
   filter(SeasType=="High")%>%
 ggplot(.,aes(x=Date,y=medSmoke,by=UnitCode))+
-  geom_point(alpha=0.4,size=3)+
-    geom_text(data=HS,aes(x=Date-2.1,y=medSmoke,label=paste(ParkName,paste0(Month,"/",Year))),size=3.5,fontface="bold")+
-    geom_text(data=HSK,aes(x=Date-2.7,y=medSmoke,label=paste(ParkName,paste0(Month,"/",Year))),size=3.5,fontface="bold")+
-    geom_text(data=HSL,aes(x=Date-2.9,y=medSmoke,label=paste(ParkName,paste0(Month,"/",Year))),size=3.5,fontface="bold")+
-    theme_classic()+mytheme+ylab("Median Smoke Value")+
-scale_y_continuous(sec.axis =sec_axis(trans = stdd ,name="Standardized"))
+  geom_point(alpha=0.4,size=2.5)+
+    geom_text(data=HS,aes(x=Date-2.4,y=medSmoke+1e-9,label=paste(ParkName,paste0(Month,"/",Year))),size=2.8,fontface="bold")+
+    geom_text(data=HSK,aes(x=Date-3,y=medSmoke,label=paste(ParkName,paste0(Month,"/",Year))),size=2.8,fontface="bold")+
+    geom_text(data=HSL,aes(x=Date-3.2,y=medSmoke,label=paste(ParkName,paste0(Month,"/",Year))),size=2.8,fontface="bold")+
+    theme_classic()+mytheme+#ylab("Monthly Median Black Carbon Density ")+
+    labs(y = bquote('Median Black Carbon'~(Kg/M^-2)), x = "Month")+
+scale_y_continuous(sec.axis =sec_axis(trans = stdd ,name="Standardized (2 SD)"))
 
 #make bp example data
 ind<-seq(-1,3,0.01)
@@ -116,14 +117,14 @@ bpdat<-gather(bpdat,key="bp",value="dep",dep0.5,dep1,dep1.5)
   which(x != TRUE)
   
   #create a 2 year AR lag
-  dat$AR_Vis2<-lag(dat$RecreationVisits,n=768)
[…]

### 2022-04-20 Rcode/PosteriorPlots.r

@@ -6,26 +6,30 @@ load("~/SmokeProject/ModelObjects/modSmokeSet0_5.rda")#modSmoke2
 load("~/SmokeProject/ModelObjects/modSmokeSet1.rda")#modSmoke3
 load("~/SmokeProject/ModelObjects/modSmokeSet1_5.rda")#modSmoke4
 
+dat<-read_csv("Data/MergedDataCompleteFINAL.csv")[,-1]
+dat<-dat%>%mutate(NameShrt=str_remove(.$ParkName,c( " NP")))%>%
+  mutate(NameShrt=str_remove(.$NameShrt,c( " & PRES")))
+dat$UnitCode<-dat$NameShrt
 #0.5
 z<-rstan::extract(modSmoke2)
 sl20.5<-as.data.frame(z$slope1)
 names(sl20.5)<-unique(as.factor(dat$UnitCode ))
 sl20.5$Model<-rep("bp0.5",nrow(sl20.5))
-sl20.5<-gather(sl20.5,key="UnitCode",value="Estimate",ARCH:ZION)
+sl20.5<-gather(sl20.5,key="UnitCode",value="Estimate",Arches:Zion)
 
 #1.0
-z<-extract(modSmoke3)
+z<-rstan::extract(modSmoke3)
 sl21.0<-as.data.frame(z$slope1)
 names(sl21.0)<-unique(as.factor(dat$UnitCode ))
 sl21.0$Model<-rep("bp1.0",nrow(sl21.0))
-sl21.0<-gather(sl21.0,key="UnitCode",value="Estimate",ARCH:ZION)
+sl21.0<-gather(sl21.0,key="UnitCode",value="Estimate",Arches:Zion)
 
 #1.5
-z<-extract(modSmoke4)
+z<-rstan::extract(modSmoke4)
 sl21.5<-as.data.frame(z$slope1)
 names(sl21.5)<-unique(as.factor(dat$UnitCode ))
 sl21.5$Model<-rep("bp1.5",nrow(sl21.5))
-sl21.5<-gather(sl21.5,key="UnitCode",value="Estimate",ARCH:ZION)
+sl21.5<-gather(sl21.5,key="UnitCode",value="Estimate",Arches:Zion)
 
 slpz<-rbind(sl20.5,sl21.0,sl21.5)
 
@@ -35,15 +39,18 @@ cols2<-c("bp0.5" = "#37a7abff",
          "bp1.0" = "#36679eff",
          "bp1.5" = "#3d3369ff")
 
[…]
